### 2021-05-14 paper/paper.tex
paper: start of paper

\documentclass[a4paper, 10pt]{article}

\usepackage[colorlinks]{hyperref}
\usepackage{microtype}

\title{All watched over by machines of loving grace}
\author{Dominic P. Mulligan}
\date{Systems Research Group, Arm Research, Cambridge}

\begin{document}

\maketitle

\begin{abstract}
To enforce system-wide security policies, operating systems implement a self-contained \emph{kernel} which serves as the root of all trust within the computer system.
The kernel is the sole component able to directly interact with hardware, and provides services to executing user-space software, for example the abstraction of a hierarchical filesystem.
User-space code resides outside of the kernel and interacts with it through a defined \emph{ABI}.

To enforce logical soundness, many interactive theorem proving systems implement a self-contained \emph{kernel} which serves as the root of all trust within the system.
The kernel is the sole system component able to authenticate a proof as legitimate.
Untrusted code resides outside of the kernel and interacts with it through a defined \emph{API}.

In this paper, we explore the consequences of structuring proof-checking software as \emph{supervisory software}.
We develop a proof-checking kernel for Gordon's HOL---an expressive typed higher-order logic---that is exposed to untrusted code through a defined ABI.
Kernel objects are protected---and important invariants are maintained---through machine-oriented notions of separation, akin to ``privilege rings'' or ``exception levels'' on typical commodity hardware, rather than programming language module boundaries.

We find that this structuring has some interesting consequences.
The typical proof-checker's \emph{meta-language}---usually some derivative of Milner's ML, privileged amongst all programming languages---is absent: any programming language able to produce code that can interact with the kernel's binary interface can be used to guide proof search, with no risk of logical unsoundness.
Moreover, the proof-checking kernel may expose functionality to untrusted code that is typical of an operating system, for example the abstraction of a hierarchical filesystem.
Filesystem access may then be conditioned on the ability to \emph{prove} some property of the dynamic system state at the point of access, allowing the kernel to enforce a range of expressive runtime security and correctness policies.
In short, untrusted code executing under the supervision of our kernel becomes \emph{proof-generating code}, exhibiting aspects of both static and dynamic (run-time) verification.

\end{abstract}

\section{Introduction}
\label{sect.introduction}

This paper is concerned with the intersection of the foundations of mathematics and systems-software design.

Modern interactive theorem proving software---such as the Isabelle, Coq, and HOL Light systems---implement a small proof-checking \emph{kernel}.
This kernel is the sole system component that is able to authenticate a purported proof as legitimate, and is typically small and self-contained, making it suitable for manual auditing.
By design, the kernel is the only component within the system that need be trusted by a skeptic, as ultimately all proofs constructed by the system, no matter how large or complex, are eventually passed through the kernel for authentication.
This implementation strategy---commonly known as the LCF-style, after Milner's LCF system that first implemented it---is a reliability mechanism, aiming to ensure that any theorem produced by the system is indeed a theorem.
That is, LCF is a design pattern that helps to write theorem proving software that remains \emph{sound}.

Shifting focus, we see an interesting series of analogies.
Typical commodity operating systems---such as Microsoft Windows and desktop Linux distributions---also implement a small \emph{kernel}.
This kernel is the sole component on the compute system that has unfettered access to hardware.

\section{Kernel design}
\label{sect.kernel.design}

\section{Proof-space support libraries}
\label{sect.proof-space.support.libraries}

\end{document}

### 2021-05-14 src/kernel/_type.rs
kernel: more runtime logging

@@ -231,7 +231,6 @@ mod test {
             tags, Handle, PREALLOCATED_HANDLE_TYPE_FORMER_ARROW,
             PREALLOCATED_HANDLE_TYPE_FORMER_PROP, PREALLOCATED_HANDLE_TYPE_PROP,
         },
-        term::Term,
     };
 
     /// Tests the various type construction methods align with the

### 2021-05-14 src/kernel/kernel_panic.rs

@@ -40,7 +40,7 @@ where
 /// Kernel logging for fatal cases: logs the provided error on the kernel's
 /// error log, then panics with the same error message.
 #[inline]
-pub fn kernel_panic<T>(message: T)
+pub fn kernel_panic<T>(message: T) -> !
 where
     T: AsRef<str>,
 {

### 2021-05-14 src/kernel/runtime_state.rs

@@ -15,7 +15,6 @@
 //! [Dominic Mulligan]: https://dominic-mulligan.co.uk
 //! [Arm Research]: http://www.arm.com/research
 
-use crate::kernel::kernel_panic::kernel_error;
 use crate::kernel::{
     _type::{
         Type, TYPE_ALPHA, TYPE_BETA, TYPE_BINARY_CONNECTIVE, TYPE_POLYMORPHIC_BINARY_PREDICATE,
@@ -41,7 +40,7 @@ use crate::kernel::{
         PREALLOCATED_HANDLE_TYPE_UNARY_PREDICATE, PREALLOCATED_HANDLE_UPPER_BOUND,
     },
     kernel_panic::{
-        kernel_info, kernel_panic, DANGLING_HANDLE_ERROR, HANDLE_EXHAUST_ERROR,
+        kernel_error, kernel_info, kernel_panic, DANGLING_HANDLE_ERROR, HANDLE_EXHAUST_ERROR,
         PRIMITIVE_CONSTRUCTION_ERROR,
     },
     name::Name,
@@ -174,6 +173,8 @@ impl RuntimeState {
     /// Functions calling this should ensure that the type argument, `tau`, is
     /// well-formed before calling.
     fn admit_type(&mut self, tau: Type) -> Handle<tags::Type> {
+        kernel_info(format!("Admitting type: {:?}.", tau));
+
         for (handle, registered) in self.types.iter() {
             if registered == &tau {
                 kernel_info(format!("Type already registered with handle: {}.", handle));
@@ -263,12 +264,11 @@ impl RuntimeState {
 
         let former = former.into();
 
-        let arity = self
-            .type_former_resolve(former.clone())
-            .ok_or({
-                kernel_error("Type-former handle is not registered.")
-                ErrorCode::NoSuchTypeFormerRegistered
-            })?;
+        let arity = self.type_former_resolve(former.clone()).ok_or({
+            kernel_error("Type-former handle is not registered.");
+
+            ErrorCode::NoSuchTypeFormerRegistered
+        })?;
 
         if !arguments
             .iter()
@@ -310,7 +310,10 @@ impl RuntimeState {
     where
         T: Into<Handle<tags::Type>> + Display,
     {
-        kernel_info(format!("Registering function type with domain {} and range {}.", domain, range));
+        kernel_info(format!(
+            "Registering function type with domain {} and range {}.",
+            domain, range
+        ));
 
[…]

### 2021-05-14 src/kernel/theorem.rs

@@ -74,7 +74,6 @@ impl Theorem {
 mod test {
     use crate::kernel::{
         handle::{tags, Handle, PREALLOCATED_HANDLE_TERM_TRUE},
-        name::Name,
         theorem::Theorem,
     };
 

### 2021-05-14 src/wasmi/host_interface.rs

@@ -25,7 +25,7 @@ use crate::kernel::{
 use byteorder::{ByteOrder, LittleEndian};
 use std::{
     borrow::Borrow,
-    fmt::{Display, Error as DisplayError, Formatter},
+    fmt::{Debug, Display, Error as DisplayError, Formatter},
     mem::size_of,
 };
 use wasmi::{
@@ -816,8 +816,8 @@ impl WasmiRuntimeState {
     ) -> Result<Handle<tags::Type>, KernelErrorCode>
     where
         T: Borrow<Handle<tags::Type>> + Clone,
-        U: Into<Name> + Clone,
-        V: Into<Handle<tags::Type>> + Clone,
+        U: Into<Name> + Clone + Debug,
+        V: Into<Handle<tags::Type>> + Clone + Debug,
     {
         self.kernel.type_substitute(handle, sigma)
     }
@@ -870,8 +870,8 @@ impl WasmiRuntimeState {
     ) -> Result<Handle<tags::Term>, KernelErrorCode>
     where
         T: Into<Handle<tags::Constant>> + Clone,
-        U: Into<Name> + Clone,
-        V: Into<Handle<tags::Type>> + Clone,
+        U: Into<Name> + Clone + Debug,
+        V: Into<Handle<tags::Type>> + Clone + Debug,
     {
         self.kernel.term_register_constant(constant, substitution)
     }

### 2021-06-19 src/wasmi/host_interface.rs
host-interface: minor amount of work on host interface

@@ -607,7 +607,7 @@ impl WasmiRuntimeState {
     ///
     /// Returns `Err(RuntimeTrap::MemoryWriteFailed)` if the write to memory at
     /// address, `address`, failed.
-    pub fn write_i32<T, U>(&self, address: T, value: U) -> Result<(), RuntimeTrap>
+    fn write_i32<T, U>(&self, address: T, value: U) -> Result<(), RuntimeTrap>
     where
         T: Into<Address>,
         U: Into<i32>,
@@ -628,7 +628,7 @@ impl WasmiRuntimeState {
     ///
     /// Returns `Err(RuntimeTrap::MemoryWriteFailed)` if the write to memory at
     /// address, `address`, failed.
-    pub fn write_i64<T, U>(&self, address: T, value: U) -> Result<(), RuntimeTrap>
+    fn write_i64<T, U>(&self, address: T, value: U) -> Result<(), RuntimeTrap>
     where
         T: Into<Address>,
         U: Into<i64>,
@@ -649,7 +649,7 @@ impl WasmiRuntimeState {
     ///
     /// Returns `Err(RuntimeTrap::MemoryWriteFailed)` if the write to memory at
     /// address, `address`, failed.
-    pub fn write_u32<T, U>(&self, address: T, value: U) -> Result<(), RuntimeTrap>
+    fn write_u32<T, U>(&self, address: T, value: U) -> Result<(), RuntimeTrap>
     where
         T: Into<Address>,
         U: Into<u32>,
@@ -670,7 +670,7 @@ impl WasmiRuntimeState {
     ///
     /// Returns `Err(RuntimeTrap::MemoryWriteFailed)` if the write to memory at
     /// address, `address`, failed.
-    pub fn write_u64<T, U>(&self, address: T, value: U) -> Result<(), RuntimeTrap>
+    fn write_u64<T, U>(&self, address: T, value: U) -> Result<(), RuntimeTrap>
     where
         T: Into<Address>,
         U: Into<u64>,
@@ -691,7 +691,7 @@ impl WasmiRuntimeState {
     ///
     /// Returns `Err(RuntimeTrap::MemoryWriteFailed)` if the write to memory at
     /// address, `address`, failed.
-    pub fn write_u64s<T, U>(&self, address: T, values: Vec<U>) -> Result<(), RuntimeTrap>
[…]

### 2021-06-21 paper/paper.tex
kernel: more work on kernel tracing and error reporting
...also added prooftree.sty and uncommitted changes from paper.tex.

@@
 
+\usepackage{amsfonts}
+\usepackage{amsmath}
+\usepackage{amssymb}
+\usepackage{amsthm}
+\usepackage[english]{babel}
+\usepackage[T1]{fontenc}
 \usepackage[colorlinks]{hyperref}
+\usepackage{listings}
 \usepackage{microtype}
+\usepackage{prooftree}
+
+\usepackage{tgpagella}
+
+\newtheorem{theorem}{Theorem}[section]
+\newtheorem{corollary}{Corollary}[theorem]
+\newtheorem{lemma}[theorem]{Lemma}
+\newtheorem{definition}{Definition}[section]
+
+\newcommand{\deffont}[1]{\ensuremath{\textbf{#1}}}
+\newcommand{\fall}[1]{\forall{#1}.\ }
+\newcommand{\ite}[3]{\mathtt{if}\ {#1}\ \mathtt{then}\ {#2}\ \mathtt{else}\ {#3}}
+\newcommand{\lam}[1]{\lambda{#1}.\ }
+\newcommand{\letin}[4]{\mathtt{let}{\;}#1{:}#2{\;}{:=}{\;}#4{\;}\mathtt{in}{\;}#3}
+\newcommand{\rulefont}[1]{(\ensuremath{\mathbf{#1}})}
+\newcommand{\hthe}[1]{\mathtt{Some}\ {#1}.\ }
+\newcommand{\xsts}[1]{\exists{#1}.\ }
 
@@
 \begin{abstract}
-To enforce system-wide security policies, operating systems implement a self-contained \emph{kernel} which serves as the root of all trust within the computer system.
+To enforce system-wide security policies, many operating systems implement a self-contained \emph{kernel} which serves as the root of all trust within the system.
 The kernel is the sole component able to directly interact with hardware, and provides services to executing user-space software, for example the abstraction of a hierarchical filesystem.
@@
 
-In this paper, we explore the consequences of structuring proof-checking software as \emph{supervisory software}.
-We develop a proof-checking kernel for Gordon's HOL---an expressive typed higher-order logic---that is exposed to untrusted code through a defined ABI.
+In this paper, we explore the consequences of structuring proof-checking software as \emph{supervisory software}, adopting system organisation principles typical of operating systems and similar privileged system software.
+We develop a proof-checking kernel for Gordon's HOL called \emph{Supervisionary} that is exposed to untrusted code through a defined ABI.
 Kernel objects are protected---and important invariants are maintained---through machine-oriented notions of separation, akin to ``privilege rings'' or ``exception levels'' on typical commodity hardware, rather than programming language module boundaries.
-
-We find that this structuring has some interesting consequences.
-The typical proof-checker's \emph{meta-language}---usually some derivative of Milner's ML, privileged amongst all programming languages---is absent: any programming language able to produce code that can interact with the kernel's binary interface can be used to guide proof search, with no risk of logical unsoundness.
-Moreover, the proof-checking kernel may expose functionality to untrusted code that is typical of an operating system, for example the abstraction of a hierarchical filesystem.
-Filesystem access may then be conditioned on the ability to \emph{prove} some property of the dynamic system state at the point of access, allowing the kernel to enforce a range of expressive runtime security and correctness policies.
-In short, untrusted code executing under the supervision of our kernel becomes \emph{proof-generating code}, exhibiting aspects of both static and dynamic (run-time) verification.
-
+We find that this novel structuring has some interesting consequences, which are discussed.
 \end{abstract}
@@
 
-This paper is concerned with the intersection of the foundations of mathematics and systems-software design.
-
-Modern interactive theorem proving software---such as the Isabelle, Coq, and HOL Light systems---implement a small proof-checking \emph{kernel}.
+Modern interactive theorem proving software---such as Isabelle, Coq, and HOL Light, amongst others---implement a small proof-checking \emph{kernel}.
 This kernel is the sole system component that is able to authenticate a purported proof as legitimate, and is typically small and self-contained, making it suitable for manual auditing.
@@
 
-\section{Kernel design}
-\label{sect.kernel.design}
+\section{Supervisionary}
+\label{sect.supervisionary}
 
-\section{Proof-space support libraries}
-\label{sect.proof-space.support.libraries}
+\subsection{Logic}
+\label{subsect.logic}
+
+Supervisionary implements a version of Gordon's higher-order logic (or HOL, henceforth), a close relative of Church's Simple Theory of Types that extends Church's system with polymorphic types, albeit with no ML-style \texttt{let}-polymorphism.
+We describe HOL here, introducing enough material for the reader to understand the remainder of the paper---motivating HOL or describing its expressiveness is outside of the scope of this paper.
+
+We fix a set of \deffont{type-formers}, and use $\mathtt{F}$, $\mathtt{G}$, $\mathtt{H}$, and so on, to range arbitrarily over type-formers.
+To each type-former, we associate an \deffont{arity}, and write $arity(\mathtt{F}) = m$ to assert that the arity of type-former $\mathsf{F}$ is $m$.
+
+\begin{definition}
+\label{defn.types}
+Define \deffont{types} with the following recursive grammar:
+\begin{gather*}
+\tau, \tau', \tau'' ::= \alpha \mid \mathsf{F}(\tau_1, \ldots, \tau_n)
+\end{gather*}
+\end{definition}
+
+We work only with \deffont{well-formed types} where each type-former $\mathtt{F}$ is applied to a list of arguments of length $m$ where $arity(\mathtt{F}) = m$.
+We fix two \emph{primitive} type-formers that are built-in to the logic: $\mathtt{bool}$ of arity $0$ and $\mathtt{fun}$ of arity $2$.
+We write $\mathtt{bool}$ and $\tau \Rightarrow \tau'$ as shorthand for $\mathtt{bool}()$ and $\mathtt{fun}(\tau, \tau')$, the Boolean and function space arrow types, respectively.
+Note that the function space arrow associates to the right, and we write $\tau \Rightarrow \tau' \Rightarrow \tau''$ as a more convenient rendering of the type $\tau \Rightarrow (\tau' \Rightarrow \tau'')$.
+We write $\tau[\alpha := \tau']$ for the \deffont{substitution action} on types, and write $tv(\tau)$ for the type variables of a type.
+
+To each type $\tau$ we fix a countably infinite set of \deffont{variables}.
+We write $x_\tau$, $y_\tau$, $z_\tau$, and so on, to range arbitrarily over variables associated with the type $\tau$.
+We adopt the convention that distinctly named variables denote distinct variables, so $x_\tau \neq y_\tau$.
+Note that $x_\tau \neq x_{\tau'}$ whenever $\tau \neq \tau'$.
+
+We fix a set of \deffont{constants} and use $\mathtt{C}$, $\mathtt{D}$, $\mathtt{E}$, and so on, to range arbitrarily over constants.
+To each constant we associate a type, and write $type(\mathtt{C}) = \tau$ to assert that constant $\mathtt{C}$ was declared with type $\tau$.
+
+\begin{definition}
+\label{defn.terms}
+Define \deffont{terms} with the following recursive grammar:
+\begin{gather*}
+r,s,t ::= x_\tau \mid \mathtt{C} \mid rs \mid \lam{x_\tau}r
+\end{gather*}
+\end{definition}
+
+We write $fv(r)$ for the set of \deffont{free variables} appearing within a term, noting that $x_\tau$ is \emph{bound} in $r$ in the term $\lam{x_\tau}r$, and write $r[x_\tau := t]$ for the usual \deffont{capture-avoiding substitution action} on terms.
+We henceforth work with terms quotiented by $\alpha$-equivalence---that is, terms are considered equal up-to a permutative renaming of their bound variables.
+
+We write $tv(r)$ for the set of \deffont{type variables} appearing within a term, and write $r[\alpha := \tau]$ for the extension of the type substitution action on types to terms.
+We fix the following \emph{primitive} constants, built into the logic, with declared types:
+
+\begin{displaymath}
+\begin{array}{ccc}
+\top, \bot & & \mathtt{bool} \\
+\neg & & \mathtt{bool} \Rightarrow \mathtt{bool} \\
+\vee, \wedge, \longrightarrow, \longleftrightarrow &  & \mathtt{bool} \Rightarrow \mathtt{bool} \Rightarrow \mathtt{bool} \\
+= & \quad\text{ with declared type }\quad & \alpha \Rightarrow \alpha \Rightarrow \mathtt{bool} \\
+\exists, \forall & & (\alpha \Rightarrow \mathtt{bool}) \Rightarrow \mathtt{bool} \\
+\mathtt{Some} & & (\alpha \Rightarrow \mathtt{bool}) \Rightarrow \alpha \\
+\end{array}
+\end{displaymath}
+
+Note that constants with polymorphic declared types such as $=$ and $\exists$ describe a family of constants related by their type.
+We write $=_\tau$ for a specialised form of the primitive equality constant where $type(=_\tau) = \tau \Rightarrow \tau \Rightarrow \mathtt{bool}$---similarly, for constants $\exists_\tau$, $\forall_\tau$, and $\mathtt{Some}_\tau$.
+We will, however, often drop the type $\tau$ from the name of constants where no ambiguity arises.
+
+We render $\exists_\tau(\lam{x_\tau}\phi)$ as $\xsts{x_\tau}\phi$, $(\vee \phi)\psi$ as $\phi \vee \psi$, and so on, following standard mathematical practice.
+We write $\phi \longleftrightarrow \psi$ as an abbreviation for $\phi =_\mathtt{bool} \psi$.
+Note that $\mathtt{Some}$ is our rendering of Hilbert's $\epsilon$-operator, a mechanism for \emph{choosing} elements of a type based on a property.
+
+\begin{figure}[t]
+\begin{gather*}
+\begin{prooftree}
+\phantom{h}
+\justifies
+x_\tau : \tau
+\end{prooftree}
+\qquad
+\begin{prooftree}
+\phantom{h}
+\justifies
+\mathtt{C}_\tau : \tau
+\end{prooftree}
+\qquad
+\begin{prooftree}
+r : \tau \Rightarrow \tau'
+\quad
+s : \tau
+\justifies
+rs : \tau'
+\end{prooftree}
+\qquad
+\begin{prooftree}
+r : \tau'
+\justifies
+\lam{x_\tau}r : \tau \Rightarrow \tau'
+\end{prooftree}
+\end{gather*}
+\caption{The typing relation on terms}
+\label{fig.typing.relation}
+\end{figure}
+
+\begin{definition}
+\label{defn.typing.relation}
+Define a \deffont{typing relation} on terms by the rules in Figure~\ref{fig.typing.relation}.
+We write $r : \tau$ to assert that a derivation tree, rooted at $r : \tau$, and constructed using the rules in Figure~\ref{fig.typing.relation}, exists.
+\end{definition}
+
+If $r : \mathtt{bool}$ that we call $r$ a \deffont{formula}.
+We will additionally use $\phi$, $\psi$, $\xi$, and so on, to range over formulae.
+We call a finite set of terms a \deffont{context}, and use $\Gamma$, $\Gamma'$, $\Gamma''$, and so on, to range over contexts.
+We write $\{\}$ for the empty context.
+
+\begin{figure}[t]
+\begin{gather*}
+\begin{prooftree}
+\phantom{h}
+\justifies
+\{\} \vdash \top
+\end{prooftree}
+\qquad
+\begin{prooftree}
+(\phi : \mathtt{bool})
+\justifies
+\{\phi\} \vdash \phi
+\end{prooftree}
+\qquad
+\begin{prooftree}
+\Gamma \vdash \bot
+\quad
+(\phi : \mathtt{bool})
+\justifies
+\Gamma \vdash \phi
+\end{prooftree}
+\qquad
+\begin{prooftree}
+\Gamma \vdash \phi
+\quad
+(\psi : \mathtt{bool})
+\justifies
+\Gamma \cup \{\psi\} \vdash \phi
+\end{prooftree}
+\qquad
+\begin{prooftree}
+(r : \tau)
+\justifies
+\{\} \vdash r = r
+\end{prooftree}
+\\[1ex]
+\begin{prooftree}
+\Gamma \vdash r = s
+\justifies
+\Gamma \vdash s = r
+\end{prooftree}
+\qquad
+\begin{prooftree}
+\Gamma \vdash r = s
+\quad
+\Gamma' \vdash s = t
+\justifies
+\Gamma \cup \Gamma' \vdash r = t
+\end{prooftree}
+\qquad
+\begin{prooftree}
+\Gamma \vdash r = s
+\quad
+x_\tau \notin fv(\Gamma)
+\justifies
+\Gamma \vdash \lam{x_\tau}r = \lam{x_\tau}s
+\end{prooftree}
+\\[1ex]
+\begin{prooftree}
+\Gamma \vdash f = g
+\quad
+\Gamma' \vdash r = s
+\quad
+(f : \tau \Rightarrow \tau', r : \tau)
+\justifies
+\Gamma \cup \Gamma' \vdash f(r) = g(s)
+\end{prooftree}
+\qquad
+\begin{prooftree}
+\Gamma \vdash \phi
+\quad
+\Gamma' \vdash \psi
+\justifies
+\Gamma \cup \Gamma' \vdash \phi \wedge \psi
+\end{prooftree}
+\qquad
+\begin{prooftree}
+\Gamma \vdash \phi \wedge \psi
+\justifies
+\Gamma \vdash \phi
+\end{prooftree}
+\\[1ex]
+\begin{prooftree}
+\Gamma \vdash \phi \wedge \psi
+\justifies
+\Gamma \vdash \psi
+\end{prooftree}
+\qquad
+\begin{prooftree}
+\Gamma \cup \{ \phi \} \vdash \psi
+\justifies
+\Gamma \vdash \phi \longrightarrow \psi
+\end{prooftree}
+\qquad
+\begin{prooftree}
+\Gamma \vdash \phi \longrightarrow \psi
+\quad
+\Gamma' \vdash \phi
+\justifies
+\Gamma \cup \Gamma' \vdash \psi
+\end{prooftree}
+\qquad
+\begin{prooftree}
+\Gamma \cup \{ \phi \} \vdash \bot
+\justifies
+\Gamma \vdash \neg\phi
+\end{prooftree}
+\\[1ex]
+\begin{prooftree}
+\Gamma \vdash \phi
+\quad
+\Gamma' \vdash \neg\phi
+\justifies
+\Gamma \cup \Gamma' \vdash \bot
+\end{prooftree}
+\qquad
+\begin{prooftree}
+\Gamma \vdash \phi
+\justifies
+\Gamma \vdash \phi \vee \psi
+\end{prooftree}
+\qquad
+\begin{prooftree}
+\Gamma \vdash \psi
+\justifies
+\Gamma \vdash \phi \vee \psi
+\end{prooftree}
+\qquad
+\begin{prooftree}
+\Gamma \vdash \phi \vee \psi
+\quad
+\Gamma \cup \{ \phi \} \vdash \xi
+\quad
+\Gamma \cup \{ \psi \} \vdash \xi
+\justifies
+\Gamma \vdash \xi
+\end{prooftree}
+\\[1ex]
+\begin{prooftree}
+\Gamma \vdash \phi \longrightarrow \psi
+\quad
+\Gamma \vdash \psi \longrightarrow \phi
+\justifies
+\Gamma \vdash \phi \longleftrightarrow \psi
+\end{prooftree}
+\qquad
+\begin{prooftree}
+\Gamma \vdash \phi \longleftrightarrow \psi
+\justifies
+\Gamma \vdash \phi \longrightarrow \psi
+\end{prooftree}
+\qquad
+\begin{prooftree}
+\Gamma \vdash \phi \longleftrightarrow \psi
+\justifies
+\Gamma \vdash \psi \longrightarrow \phi
+\end{prooftree}
+\qquad
+\begin{prooftree}
+(r : \tau \Rightarrow \tau', x_\tau \notin fv(r))
+\justifies
+\{\} \vdash \lam{x_\tau}r(x_\tau) = r
+\end{prooftree}
+\\[1ex]
+\begin{prooftree}
+((\lam{x_\tau}r)t : \tau')
+\justifies
+\{\} \vdash (\lam{x_\tau}r)t = r[x_\tau := t]
+\end{prooftree}
+\qquad
+\begin{prooftree}
+\Gamma \vdash \phi
+\justifies
+\Gamma[\alpha := \tau] \vdash \phi[\alpha := \tau]
+\end{prooftree}
+\qquad
+\begin{prooftree}
+\Gamma \vdash \phi
+\quad
+(r : \tau)
+\justifies
+\Gamma[x_\tau := r] \vdash \phi[x_\tau := r]
+\end{prooftree}
+\\[1ex]
+\begin{prooftree}
+\Gamma \vdash \phi
+\quad
+(x_\tau \notin fv(\Gamma))
+\justifies
+\Gamma \vdash \fall{x_\tau}\phi
+\end{prooftree}
+\qquad
+\begin{prooftree}
+\Gamma \vdash \fall{x_\tau}\phi
+\quad
+(t : \tau)
+\justifies
+\Gamma \vdash \phi[x_\tau := t]
+\end{prooftree}
+\qquad
+\begin{prooftree}
+\Gamma \vdash \phi[x_\tau := t]
+\justifies
+\Gamma \vdash \xsts{x_\tau}\phi
+\end{prooftree}
+\\[1ex]
+\begin{prooftree}
+\Gamma \vdash \xsts{x_\tau}\phi
+\quad
+\Gamma \cup \{ \phi[x_\tau := y_\tau] \} \vdash \psi
+\quad
+(y_\tau \notin fv(\Gamma) \cup fv(\psi))
+\justifies
+\Gamma \vdash \psi
+\end{prooftree}
+\qquad
+\begin{prooftree}
+\Gamma \vdash \xsts{x_\alpha}p(x)
+\justifies
+\Gamma \vdash p(\mathtt{Some}_\alpha\ p)
+\end{prooftree}
+\end{gather*}
+\caption{The Natural Deduction relation}
+\label{fig.natural.deduction.relation}
+\end{figure}
+
+\begin{definition}
+\label{defn.natural.deduction}
+Define a \deffont{Natural Deduction relation} between contexts and terms, $\Gamma \vdash r$, using the rules in Figure~\ref{fig.natural.deduction.relation}.
+We write $\Gamma \vdash r$ to assert that a derivation tree, rooted at $\Gamma \vdash r$, and constructed using the rules in Figure~\ref{fig.natural.deduction.relation}, exists.
+\end{definition}
+
+% Note that the Natural Deduction relation maintains a series of invariants---for example, $\Gamma \vdash r$ implies that $r : \mathtt{bool}$ and also $s : \mathtt{bool}$ for each $s \in \Gamma$.
+
+The set of primitive type-formers and constants defines an implicit \deffont{theory}.
+Our version of HOL also includes two \emph{primitive definitional} principles for extending this theory with new defined constants and types, respectively.
+Specifically, for a previously-undefined constant $\mathtt{C}$ and a term $r : \tau$ where $fv(r) = \{\}$ the constant $\mathtt{C}$ with declared type $\tau$ is added to the theory, as is an additional axiom $\{\} \vdash \mathtt{C} = r$.
+TODO: type-definition
+
+\subsection{The kernel runtime state}
+\label{subsect.the.kernel.runtime.state}
+
+The Supervisionary kernel maintains a mutable runtime state consisting of a series of mutable \deffont{heaps}.
+These heaps are indexed by \deffont{kernel handles}---unsigned integers generated by the kernel---and which can be thought of as pointers to \deffont{kernel objects} in the various heaps.
+Importantly, the kernel never exposes kernel objects to untrusted code directly, but performs all manipulations on these objects itself, under instruction by untrusted code.
+
+The Supervisionary kernel uses its heaps as a definitive ``source of truth''.
+Whereas an LCF system represents a theorem as an immutable and opaque element of a dedicated theorem \emph{type} within the system's meta-language, Supervisionary represents authenticated theorems as objects within its internal \deffont{heap of theorems}.
+Note, for the heap of theorems to be a reliable record of authenticated theorems, any terms, constants, types, and type-formers, mentioned by an entry in the theorem heap (or transitively, thereof) must also be made immutable and put under the control of the kernel---it would be decidedly \emph{inconvenient} if the meaning of a theorem $\{ \} \vdash \phi$ were to change under the nose of the kernel.
+As a result, the kernel also maintains separate heaps for registered constants, type-formers, well-formed types, and type-checked terms.
+
+The simplest---and most foundational---of these additional heaps is the \deffont{heap of type-formers}, which records the arities of type-formers registered with the kernel.
+When untrusted code requests that a new type-former is registered with the kernel, it calls an ABI function---\texttt{TypeFormer.Handle.Register}---with a supplied arity.
+In response, the kernel invents a fresh handle and associates it with the supplied arity, before returning the new handle back to untrusted code to use to refer to the type-former again in future.
+Note that we can describe the action of the various ABI functions exposed by the kernel with Hoare triples, using an informal Separation Logic as the assertion language.
+
+Another ABI call, \texttt{TypeFormer.Resolve}, allows one to dereference a handle to obtain the arity of a type-former.
+At kernel initialisation, the type-former heap is populated with the arities of the Boolean and function space arrow type-formers, registered at pre-provisioned handles.
+These handles are part of the ABI itself, and are required ``common knowledge'' between the kernel and untrusted code, similar to the standard file descriptors \texttt{stdin} and \texttt{stdout} forming part of the Unix ABI.
+
+\begin{figure}[t]
+\caption{The internal structure of the kernel's type heap}
+\label{fig.type.heap}
+\end{figure}
+
+Building on the heap of type-formers is the \deffont{heap of types}.
+A key consideration in designing the Supervisionary kernel is handling the recursive structure of definitions like Definitions~\ref{defn.types} and~\ref{defn.terms}.
+To this end, cells in the type heap store either a type-variable or a type-former, following the grammar of types.
+However, the recursive structure has been captured by replacing all recursive occurrences in the grammar by handles pointing back into the type heap---presented schematically in Figure~\ref{fig.type.heap}---with the reference to a type-former in the grammar of types also replaced by a handle pointing-to a type-former in the relevant heap.
+
+New types are entered into the heap via \texttt{Type.Register.Variable} and \texttt{Type.Register.Combination}, respectively.
+The latter of these is the most interesting as the kernel must be careful when entering new type-former entries, owing to its recursive character.
+In particular, it must ensure that the heap remains \deffont{inductive} in the sense that no entry in the heap contains ``dangling pointers'', or handles pointing to an unregistered type-former or type.
+
+For types, the Supervisionary kernel enforces \deffont{maximal sharing}, in the sense that trying to register a type with the kernel that has already been registered (up-to syntactic equality) does not make a duplicate copy, but simply returns the handle to the already-registered type.
+This allows code inside the kernel to quickly compare types for syntactic equality, by simply comparing handles for equality at the expense of making the registering of types slower.
+
+In addition to commands for registering types, and dereferencing handles, we also provide mechanisms for ``testing'' and ``destructuring'' objects that are pointed-to by handles.
+For example \texttt{Type.Split.Combination} takes a handle pointing-to a type-former applied to type arguments as argument and returns a type-former handle and list of type handles as output.
+
+\subsection{Formal specification of the ABI}
+\label{subsect.formal.specification.of.the.abi}
+
+\begin{definition}
+\label{defn.handle}
+Fix a countably infinite set of \deffont{handles}.
+We use $h$, $h'$, $h''$, and so on, to range arbitrarily over handles, and write $\vec{h}$ to denote a finite list of handles.
+\end{definition}
+
+For a partial function $f : A \rightharpoonup B$ we write $dom(f)$ for the \deffont{domain} of $f$, the set $\{ a \in A \mid f a \text{ is defined} \}$.
+If $dom(f)$ is finite then we call $f$ \deffont{finitely supported}.
+
+We write $a \mapsto b$ for the \deffont{singleton function} that maps $a \in A$ to $b \in B$ and is everywhere else undefined, and write $\emptyset$ for the \deffont{empty function} that is everywhere undefined.
+Note that $dom(a \mapsto b) = \{a \}$ and $dom(\emptyset) = \{\}$.
+For a function $f : A \rightharpoonup B$ we write $f(a \mapsto b)$ where $f(a \mapsto b)(a) = b$ and $f(a \mapsto b)(c) = f(c)$.
+
+We call two functions $f, g : A \rightharpoonup B$ \deffont{disjoint} whenever $dom(f) \cap dom(g) = \{\}$.
+For two disjoint functions $f, g : A \rightharpoonup B$ we write $f \uplus g$ for the \deffont{disjoint union} of the two functions, where $(f \uplus g)(a) = f(a)$ if $a \in dom(f)$, $(f \uplus g)(a) = g(a)$ if $a \in dom(g)$, and is otherwise undefined.
+
+\begin{definition}
+\label{defn.kernel.objects}
+Define \deffont{type}, \deffont{term}, and \deffont{theorem kernel objects} by the following grammars, respectively:
+\begin{gather*}
+\mathtt{var}(n) \mid \mathtt{comb}(h, \vec{h}) \\
+\mathtt{var}(n, h) \mid \mathtt{const}(h, h) \mid \mathtt{app}(h, h) \mid \mathtt{lam}(n, h, h) \\
+\mathtt{theorem}(\vec{h}, h)
+\end{gather*}
+\end{definition}
+
+\begin{definition}
+\label{defn.heaps}
+Call a finitely supported partial function from handles to natural number arities a \deffont{type-former heap}.
+Call a finitely-supported partial function from handles to type kernel objects a \deffont{type heap} (or a \deffont{constant heap}, depending on context).
+Call a finitely supported partial function from handles to term kernel objects a \deffont{term heap}.
+Call a finitely supported partial function from handles to theorem kernel objects a \deffont{theorem heap}.
+\end{definition}
+
+\begin{definition}
+\label{defn.runtime.state}
+Call a 5-tuple $\langle \frak{F}, \frak{T}, \frak{C}, \frak{R}, \frak{D} \rangle$ a \deffont{runtime state} whenever $\frak{F}$ is a type-former heap, $\frak{T}$ is a type heap, $\frak{C}$ is a constant heap, $\frak{R}$ is a term heap, and $\frak{D}$ is a theorem heap.
+We use $\sigma$, $\sigma'$, $\sigma''$, and so on, to range arbitrarily over runtime states.
+\end{definition}
+
+\begin{definition}
+\label{defn.command.result}
+\end{definition}
+
+We have:
+\begin{displaymath}
+\{ \top \}\quad \mathtt{TypeFormer.Register}(a)\quad \{ \mathtt{result} = \mathtt{Ok}() \otimes \exists h.\ h \mapsto_{tf} a \}
+\end{displaymath}
+Here, $h \mapsto_{tf} a$ asserts that the handle $h$ points-to a slot in the type-former heap containing the arity, $a$, the precondition $\top$, or truth, indicates that the ABI function may be invoked in any kernel runtime state, $\phi \otimes \psi$ is the \emph{separating conjunction}, and $\mathtt{result} = \mathtt{Ok}()$ indicates that the command terminated successfully.
+Commands do not ``fail'' in the typical Hoare-logic sense, but indicate errors by successfully terminating with an error code.
+
+where $\mathtt{result} = \mathtt{Ok}(a)$ asserts the command successfully terminates with $a$:
+\begin{displaymath}
+\{ h \mapsto_{tf} a \}\quad \mathtt{TypeFormer.Resolve}(h) \quad\{ h \mapsto_{tf} a \otimes \mathtt{result} = \mathtt{Ok}(a) \}
+\end{displaymath}
+
+\begin{minipage}[t]{0.5\linewidth}
+\begin{gather*}
+\{ h \mapsto_{t} \mathtt{Comb}(f, as) \} \\
+\mathtt{Type.Register.Combination}(f, as) \\
+\{ h \mapsto_{t} \mathtt{Comb}(f, as) \otimes \mathtt{result} = \mathtt{Ok}(h) \}
+\end{gather*}
+\end{minipage}
+\;
+\begin{minipage}[t]{0.5\linewidth}
+\begin{gather*}
+\{ \fall{a\in as}\xsts{t} a \mapsto_t t \otimes f \mapsto_{tf} len(as) \} \\
+\mathtt{Type.Register.Combination}(f, as) \\
+\{ \xsts{h}\mathtt{result} = \mathtt{Ok}(h) \otimes h \mapsto_t \mathtt{Comb}(f, as)\ \otimes \\ \fall{a\in as}\xsts{t} a \mapsto_t t \otimes f \mapsto_{tf} len(as)\}
+\end{gather*}
+\end{minipage}
+
+\begin{minipage}[t]{0.5\linewidth}
+\begin{gather*}
+\{ h \mapsto_t \mathtt{Comb}(f, as) \} \\
+\mathtt{Type.Split.Combination}(h) \\
+\{ h \mapsto_t \mathtt{Comb}(f, as) \otimes \mathtt{result} = \mathtt{Ok}(f, as) \}
+\end{gather*}
+\end{minipage}
+\begin{minipage}[t]{0.5\linewidth}
+\begin{gather*}
+\{ h \not\mapsto_t \mathtt{Comb}(f, as) \} \\
+\mathtt{Type.Split.Combination}(h) \\
+\{ \mathtt{result} = \mathtt{ErrDangle} \}
+\end{gather*}
+\end{minipage}
 

### 2021-06-21 src/kernel/runtime_state.rs

@@ -655,7 +655,14 @@ impl RuntimeState {
     where
         T: Into<Handle<tags::Type>> + Clone,
     {
+        kernel_info(format!(
+            "Registering constant with type handle: {}.",
+            handle.clone().into()
+        ));
+
         if !self.type_is_registered(handle.clone().into()) {
+            kernel_error("Unknown handle.");
+
             return Err(ErrorCode::NoSuchTypeRegistered);
         }
 
@@ -676,6 +683,11 @@ impl RuntimeState {
     where
         T: Borrow<Handle<tags::Constant>>,
     {
+        kernel_info(format!(
+            "Resolving constant with handle: {}.",
+            handle.borrow()
+        ));
+
         self.constants
             .get(handle.borrow())
             .ok_or(ErrorCode::NoSuchConstantRegistered)
@@ -688,6 +700,11 @@ impl RuntimeState {
     where
         T: Borrow<Handle<tags::Constant>>,
     {
+        kernel_info(format!(
+            "Checking constant {} is registered.",
+            handle.borrow()
+        ));
+
         self.constant_resolve(handle).is_ok()
     }
 
@@ -729,10 +746,18 @@ impl RuntimeState {
         handle: U,
     ) -> Result<Handle<tags::Term>, ErrorCode>
     where
-        T: Into<Name>,
+        T: Into<Name> + Clone,
         U: Into<Handle<tags::Type>> + Clone,
     {
+        kernel_info(format!(
+            "Registering variable with name {} and type handle {}.",
+            name.clone().into(),
+            handle.clone().into()
+        ));
+
         if !self.type_is_registered(handle.clone().into()) {
+            kernel_error("Unknown handle.");
+
             return Err(ErrorCode::NoSuchTypeRegistered);
         }
 
@@ -762,6 +787,11 @@ impl RuntimeState {
         U: Into<Name> + Clone + Debug,
         V: Into<Handle<tags::Type>> + Clone + Debug,
     {
+        kernel_info(format!(
+            "Registering constant with handle: {}.",
+            handle.clone().into()
+        ));
+
         let cnst = self.constant_resolve(handle.clone().into())?.clone();
 
         let tau = self.type_substitute(cnst, type_substitution)?;
@@ -792,11 +822,19 @@ impl RuntimeState {
         T: Into<Handle<tags::Term>> + Clone,
         U: Into<Handle<tags::Term>> + Clone,
     {
+        kernel_info(format!(
[…]

### 2021-06-21 src/kernel/runtime_state.rs
kernel: fixed compile-time errors from recent changes

@@ -1382,7 +1382,7 @@ impl RuntimeState {
                 ErrorCode::NotANegation
             })?;
 
-        let (constant, _tau) = self.term_split_constant(left).ok_or({
+        let (constant, _tau) = self.term_split_constant(left).map_err(|_e| {
             kernel_error("Term is not an application.");
 
             ErrorCode::NotANegation
@@ -1428,7 +1428,7 @@ impl RuntimeState {
                 ErrorCode::NotAnEquality
             })?;
 
-        let (constant, _tau) = self.term_split_constant(left).ok_or({
+        let (constant, _tau) = self.term_split_constant(left).map_err(|_e| {
             kernel_error("Term is not a constant.");
 
             ErrorCode::NotAnEquality
@@ -1466,7 +1466,7 @@ impl RuntimeState {
 
         let (left, mid) = self
             .resolve_term_handle(left)
-            .unwrap_or_else(kernel_panic(DANGLING_HANDLE_ERROR))
+            .unwrap_or_else(|_e| kernel_panic(DANGLING_HANDLE_ERROR))
             .split_application()
             .ok_or({
                 kernel_error("Term is not an application.");
@@ -1474,7 +1474,7 @@ impl RuntimeState {
                 ErrorCode::NotADisjunction
             })?;
 
-        let (constant, _tau) = self.term_split_constant(left).or_or({
+        let (constant, _tau) = self.term_split_constant(left).map_err(|_e| {
             kernel_error("Term is not a constant.");
 
             ErrorCode::NotADisjunction
@@ -1512,7 +1512,7 @@ impl RuntimeState {
 
         let (left, mid) = self
             .resolve_term_handle(left)
-            .unwrap_or_else(kernel_panic(DANGLING_HANDLE_ERROR))
+            .unwrap_or_else(|_e| kernel_panic(DANGLING_HANDLE_ERROR))
             .split_application()
             .ok_or({
                 kernel_error("Term is not an application.");
@@ -1520,7 +1520,7 @@ impl RuntimeState {
                 ErrorCode::NotAConjunction
             })?;
 
-        let (constant, _tau) = self.term_split_constant(left).ok_or({
+        let (constant, _tau) = self.term_split_constant(left).map_err(|_e| {
             kernel_error("Term is not a constant.");
 
             ErrorCode::NotAConjunction
@@ -1558,7 +1558,7 @@ impl RuntimeState {
 
         let (left, mid) = self
             .resolve_term_handle(left)
[…]

### 2021-06-21 src/wasmi/host_interface.rs

@@ -1039,7 +1039,7 @@ impl WasmiRuntimeState {
         tau: U,
     ) -> Result<Handle<tags::Term>, KernelErrorCode>
     where
-        T: Into<Name>,
+        T: Into<Name> + Clone,
         U: Into<Handle<tags::Type>> + Clone,
     {
         self.kernel.borrow_mut().term_register_variable(name, tau)
@@ -1084,7 +1084,7 @@ impl WasmiRuntimeState {
         body: V,
     ) -> Result<Handle<tags::Term>, KernelErrorCode>
     where
-        T: Into<Name>,
+        T: Into<Name> + Clone,
         U: Into<Handle<tags::Type>> + Clone,
         V: Into<Handle<tags::Term>> + Clone,
     {
@@ -1167,7 +1167,7 @@ impl WasmiRuntimeState {
         body: V,
     ) -> Result<Handle<tags::Term>, KernelErrorCode>
     where
-        T: Into<Name>,
+        T: Into<Name> + Clone,
         U: Into<Handle<tags::Type>> + Clone,
         V: Into<Handle<tags::Term>> + Clone,
     {
@@ -1184,7 +1184,7 @@ impl WasmiRuntimeState {
         body: V,
     ) -> Result<Handle<tags::Term>, KernelErrorCode>
     where
-        T: Into<Name>,
+        T: Into<Name> + Clone,
         U: Into<Handle<tags::Type>> + Clone,
         V: Into<Handle<tags::Term>> + Clone,
     {

### 2021-06-22 src/kernel/_type.rs
kernel: changed names from strings to u64 values

@@ -160,12 +160,12 @@ impl Type {
 ////////////////////////////////////////////////////////////////////////////////
 
 lazy_static! {
-    /// The "alpha" type-variable, `A`.
+    /// The "alpha" type-variable, `0`.
     pub static ref TYPE_ALPHA: Type = Type::Variable {
         name: 0u64,
     };
 
-    /// The "beta" type-variable, `B`.
+    /// The "beta" type-variable, `1`.
     pub static ref TYPE_BETA: Type = Type::Variable {
         name: 1u64,
     };
@@ -264,7 +264,7 @@ mod test {
             .is_variable()
         );
         assert!(
-            Type::function(PREALLOCATED_HANDLE_TYPE_PROP, PREALLOCATED_HANDLE_TYPE_PROP)
+            !Type::function(PREALLOCATED_HANDLE_TYPE_PROP, PREALLOCATED_HANDLE_TYPE_PROP)
                 .is_variable()
         );
     }

### 2021-06-22 src/kernel/name.rs

@@ -17,7 +17,7 @@
 // Miscellaneous material.
 ////////////////////////////////////////////////////////////////////////////////
 
-use crate::kernel::kernel_panic::kernel_info;
+use crate::kernel::kernel_panic::{kernel_info, kernel_panic};
 
 ////////////////////////////////////////////////////////////////////////////////
 // Names and related material.
@@ -33,11 +33,15 @@ fn fresh<T>(mut avoid: T) -> Name
 where
     T: Iterator<Item = Name>,
 {
-    let mut counter = 0u64;
+    let mut counter = 0;
 
     loop {
         if avoid.any(|x| x == counter) {
-            counter += 1;
+            if let Some(next) = counter.checked_add(1) {
+                counter = next;
+            } else {
+                kernel_panic("Exhausted fresh name generation.");
+            }
         } else {
             kernel_info(format!("Fresh name generated: {}.", counter));
             return counter;
@@ -56,8 +60,8 @@ mod test {
     /// Tests that fresh-name generation is indeed fresh.
     #[test]
     pub fn name_test0() {
-        let n = fresh(0u64..100u64);
+        let n = fresh(0..100);
 
-        assert!(!(0u64..100u64).contains(&n));
+        assert!(!(0..100).contains(&n));
     }
 }

### 2021-06-22 paper/paper.tex
paper: more work on introduction

@@
 
-Modern interactive theorem proving software---such as Isabelle, Coq, and HOL Light, amongst others---implement a small proof-checking \emph{kernel}.
-This kernel is the sole system component that is able to authenticate a purported proof as legitimate, and is typically small and self-contained, making it suitable for manual auditing.
-By design, the kernel is the only component within the system that need be trusted by a skeptic, as ultimately all proofs constructed by the system, no matter how large or complex, are eventually passed through the kernel for authentication.
-This implementation strategy---commonly known as the LCF-style, after Milner's LCF system that first implemented it---is a reliability mechanism, aiming to ensure that any theorem produced by the system is indeed a theorem.
-That is, LCF is a design pattern that helps to write theorem proving software that remains \emph{sound}.
+Interactive theorem proving software typically implements a proof-checking \emph{kernel}.
+This kernel is the sole system component that is able to authenticate a purported proof as legitimate, and is typically kept small and self-contained, making it suitable for manual auditing.
+By design, the kernel is the only component within the system that need be trusted by a skeptic, as all proofs constructed by the system, no matter how large or complex, are ultimately passed through the kernel for authentication.
+This implementation strategy---known as the ``LCF-style'', after Milner's system that first implemented it---is a reliability mechanism, aiming to ensure that any theorem produced by the system is indeed a theorem.
+In short, the kernel acts as the \emph{root of all trust} within the system.
 
-Shifting focus, we see an interesting series of analogies.
-Typical commodity operating systems---such as Microsoft Windows and desktop Linux distributions---also implement a small \emph{kernel}.
-This kernel is the sole component on the compute system that has unfettered access to hardware.
+Shifting focus, we see an interesting analogy.
+Typical commodity operating systems---such as Microsoft Windows and desktop Linux distributions---also implement a \emph{kernel}.
+This kernel is also the sole system component that has unfettered access to the host machine's hardware\footnote{Barring other privileged systems software, such as the hypervisor, which we consider as another form of operating system}, and will typically provide various abstractions, such as a hierarchically-structured filesystem, sockets, and so on, to untrusted, user-space code.
+System-wide security policies are enforced by the kernel using a variety of mechanisms, including password-protected user accounts, file and directory permissions, access control lists, firewalls, and runtime-monitoring facilities.
+From a security point-of-view, an operating system kernel again acts as the \emph{root of all trust} within the system.
+
+At heart---though both types of kernels are intended for very different use-cases---both types of kernel are intended to enforce a series of \emph{kernel invariants} even in the face of unbounded interaction with untrusted code.
+However, despite this common cause, the implementation techniques employed by authors of the two type of kernels tends to be very different.
+In particular, an LCF-style kernel is little other than a library implemented in a high-level programming language---typically some variant of ML, such as Standard ML or OCaml.
+Invariants are maintained by the kernel judiciously exposing functionality as an API that offers functions to manipulate elements of a collection of abstract types, thereby exploiting the soundness of the implementation language's type-system.
+Untrusted code then interacts with the kernel through this API.
 

### 2021-06-22 paper/paper.tex
paper: more work on introduction

@@
 
-Shifting focus, we see an interesting analogy.
-Typical commodity operating systems---such as Microsoft Windows and desktop Linux distributions---also implement a \emph{kernel}.
-This kernel is also the sole system component that has unfettered access to the host machine's hardware\footnote{Barring other privileged systems software, such as the hypervisor, which we consider as another form of operating system}, and will typically provide various abstractions, such as a hierarchically-structured filesystem, sockets, and so on, to untrusted, user-space code.
+Shifting focus, we see an interesting analogy: desktop operating systems---such as Microsoft Windows and desktop Linux distributions---also implement a \emph{kernel}.
+This kernel is also the sole system component that has unfettered access to the host machine's hardware\footnote{Barring other privileged systems software, such as the hypervisor, which we consider as another form of operating system}, and will typically provide various abstractions, such as a hierarchically-structured filesystem, sockets, and so on, to untrusted user-space code.
 System-wide security policies are enforced by the kernel using a variety of mechanisms, including password-protected user accounts, file and directory permissions, access control lists, firewalls, and runtime-monitoring facilities.
@@
 
-At heart---though both types of kernels are intended for very different use-cases---both types of kernel are intended to enforce a series of \emph{kernel invariants} even in the face of unbounded interaction with untrusted code.
-However, despite this common cause, the implementation techniques employed by authors of the two type of kernels tends to be very different.
+At heart---though intended for very different use-cases---both types of kernel are intended to enforce a series of \emph{kernel invariants} even in the face of unbounded interaction with untrusted code.
+However, despite this common cause between the two families of systems, the protection mechanisms used to enforce these invariants tend to be very different.
+
 In particular, an LCF-style kernel is little other than a library implemented in a high-level programming language---typically some variant of ML, such as Standard ML or OCaml.
-Invariants are maintained by the kernel judiciously exposing functionality as an API that offers functions to manipulate elements of a collection of abstract types, thereby exploiting the soundness of the implementation language's type-system.
-Untrusted code then interacts with the kernel through this API.
+Invariants are maintained by the kernel judiciously exposing functionality through an API---offering functions to manipulate elements of certain abstract types---and through which untrusted code interacts with the kernel.
+Logical soundness is therefore contingent on the type soundness of the implementation meta-language's type-system.
+
+In contrast, desktop operating systems use hardware separation mechanisms to protect various \emph{kernel objects}---that is, book-keeping structures used by the kernel---from unfettered access by untrusted code.
+These hardware separation mechanisms include elevated \emph{privilege rings} (or \emph{exception levels})---granting the kernel access to protected hardware resources, such as system registers---and memory protection mechanisms through which the kernel can ``carve out'' a private scratch space to safely stash kernel objects.
 
@@
 
-Supervisionary implements a version of Gordon's higher-order logic (or HOL, henceforth), a close relative of Church's Simple Theory of Types that extends Church's system with polymorphic types, albeit with no ML-style \texttt{let}-polymorphism.
+Supervisionary implements a version of Gordon's higher-order logic (or HOL, henceforth), a close relative of Church's Simple Theory of Types that extends Church's system with top-level polymorphism.
 We describe HOL here, introducing enough material for the reader to understand the remainder of the paper---motivating HOL or describing its expressiveness is outside of the scope of this paper.

### 2021-06-24 paper/paper.tex
paper: added Nick as an author, reorganized the paper's structure

@@
 \title{All watched over by machines of loving grace}
-\author{Dominic P. Mulligan}
+\author{Dominic P. Mulligan and Nick Spinale}
 \date{Systems Research Group, Arm Research, Cambridge}
@@
 
-\section{Supervisionary}
-\label{sect.supervisionary}
+\paragraph{Contributions}
+We make the following contributions in this paper:
 
-\subsection{Logic}
-\label{subsect.logic}
+\begin{enumerate}
+\item
+We design a proof-checking kernel for Gordon's HOL structured around a series of tables of kernel objects indexed by handles.
+With this, we introduce \emph{Supervisionary}, a prototype interactive theorem proving system which exposes its kernel as an ABI to WebAssembly programs.
+\item
+Noting that Supervisionary is little other than an exercise in pointer manipulation, we briefly explore how our ABI can be formally specified, using a variant of Separation Logic.
+\item
+We exploit Supervisionary's proof checking capabilities to define expressive dynamic security policies for WebAssembly programs.
+\end{enumerate}
+
+The remainder of the paper is structured as follows.
+In Section~\ref{sect.gordons.hol} we briefly introduce Gordon's HOL.
+In Section~\ref{sect.design.supervisionary.kernel} we describe the design of the Supervisionary kernel.
+In Section~\ref{sect.supervisionary.system} we describe the Supervisionary system proper, including untrusted support code---the analogue of Unix's \texttt{libc}---that allows WebAssembly programs to interact with the kernel.
+In Section~\ref{sect.separation.logic.an.interlude} we briefly explore how Separation Logic can be used to provide a formal description of the action of the Supervisionary ABI.
+In Section~\ref{sect.security.policies.for.webassembly} we show how expressive security policies can be captured, and enforced, for arbitrary WebAssembly programs using Supervisionary.
+Finally, in Section~\ref{sect.conclusions} we conclude.
+
+\section{Gordon's HOL}
+\label{sect.gordons.hol}
 
@@
 
-\subsection{The kernel runtime state}
-\label{subsect.the.kernel.runtime.state}
+\section{The design of the Supervisionary kernel}
+\label{sect.design.supervisionary.kernel}
 
@@
 For example \texttt{Type.Split.Combination} takes a handle pointing-to a type-former applied to type arguments as argument and returns a type-former handle and list of type handles as output.
+
+\section{The Supervisionary system}
+\label{sect.supervisionary.system}
+
+\section{Separation Logic: an interlude}
+\label{sect.separation.logic.an.interlude}
+
+\section{Security policies for WebAssembly}
+\label{sect.security.policies.for.webassembly}
+
+\section{Conclusions}
+\label{sect.conclusions}
+
+\end{document}
 
@@
 \end{minipage}
-
-\end{document}

### 2021-06-24 paper/paper.tex
paper: more work on introduction

@@
 In contrast, desktop operating systems use hardware separation mechanisms to protect various \emph{kernel objects}---that is, book-keeping structures used by the kernel---from unfettered access by untrusted code.
-These hardware separation mechanisms include elevated \emph{privilege rings} (or \emph{exception levels})---granting the kernel access to protected hardware resources, such as system registers---and memory protection mechanisms through which the kernel can ``carve out'' a private scratch space to safely stash kernel objects.
+These hardware separation mechanisms include \emph{privilege rings} (or \emph{exception levels})---granting the kernel access to protected hardware resources, such as system registers---and memory protection mechanisms through which the kernel can ``carve out'' a private scratch space to safely stash kernel objects.
+The kernel offers services to untrusted user-space code via a binary interface, or an ABI, through the use of \emph{system calls}---made more usable to programmers by libraries such as the C standard library, \texttt{libc}---and which provoke a context switch into the kernel.
 
-\paragraph{Contributions}
+What are the consequences, therefore, of organising a proof-checking system along the lines of a typical operating system, exposing services to untrusted code through an ABI and using privilege, rather than module boundaries, to maintain kernel invariants?
+In this paper, we introduce a prototype system---called \emph{Supervisionary}---to explore the consequences of doing just that.
+
+Supervisionary is implemented as a runtime host for a WebAssembly (or Wasm, henceforth) virtual machine.
+Wasm is a synthetic instruction set and acts as a common target for a number of high-level programming languages, including Rust, C, and C++ at the time of writing.
+Though originally intended as a sandboxing mechanism for use within web browsers, Wasm is increasingly finding application outside of the browser.
+
+\noindent
 We make the following contributions in this paper:
@@
 \item
-We design a proof-checking kernel for Gordon's HOL structured around a series of tables of kernel objects indexed by handles.
+We design a novel proof-checking kernel for Gordon's HOL structured around a series of tables of kernel objects indexed by handles.
 With this, we introduce \emph{Supervisionary}, a prototype interactive theorem proving system which exposes its kernel as an ABI to WebAssembly programs.

### 2021-06-25 paper/paper.tex
paper: more on introduction

--- 2021-06-24 paper/paper.tex
@@
 
-Interactive theorem proving software typically implements a proof-checking \emph{kernel}.
-This kernel is the sole system component that is able to authenticate a purported proof as legitimate, and is typically kept small and self-contained, making it suitable for manual auditing.
-By design, the kernel is the only component within the system that need be trusted by a skeptic, as all proofs constructed by the system, no matter how large or complex, are ultimately passed through the kernel for authentication.
+Interactive theorem proving software typically implements a proof-checking \emph{kernel} which is the sole system component able to authenticate a purported proof as genuine.
+The kernel is the only component within the system that need be trusted by a skeptic, as all proofs constructed by the system, no matter how large or complex, are ultimately passed through the kernel for authentication.
 This implementation strategy---known as the ``LCF-style'', after Milner's system that first implemented it---is a reliability mechanism, aiming to ensure that any theorem produced by the system is indeed a theorem.
@@
 
-Shifting focus, we see an interesting analogy: desktop operating systems---such as Microsoft Windows and desktop Linux distributions---also implement a \emph{kernel}.
-This kernel is also the sole system component that has unfettered access to the host machine's hardware\footnote{Barring other privileged systems software, such as the hypervisor, which we consider as another form of operating system}, and will typically provide various abstractions, such as a hierarchically-structured filesystem, sockets, and so on, to untrusted user-space code.
-System-wide security policies are enforced by the kernel using a variety of mechanisms, including password-protected user accounts, file and directory permissions, access control lists, firewalls, and runtime-monitoring facilities.
+Shifting focus, we see an interesting analogy: desktop operating systems also implement a \emph{kernel}, which is the sole system component with unfettered access to hardware\footnote{Barring other privileged systems software, such as the hypervisor, which we consider as another form of operating system}, and provides useful abstractions, such as a filesystem, sockets, and so on, to untrusted user-space code.
+Security policies are enforced by the kernel using a variety of mechanisms, file and directory permissions, access control lists, and runtime-monitoring.
 From a security point-of-view, an operating system kernel again acts as the \emph{root of all trust} within the system.
 
-At heart---though intended for very different use-cases---both types of kernel are intended to enforce a series of \emph{kernel invariants} even in the face of unbounded interaction with untrusted code.
-However, despite this common cause between the two families of systems, the protection mechanisms used to enforce these invariants tend to be very different.
+Note that, though intended for very different use-cases, at heart both types of kernel are intended to enforce a series of \emph{kernel invariants} even in the face of unbounded interaction with untrusted code.
+Yet, despite this common cause, the mechanisms used to maintain invariants is very different.
 
-In particular, an LCF-style kernel is little other than a library implemented in a high-level programming language---typically some variant of ML, such as Standard ML or OCaml.
+An LCF-style kernel is little other than a library implemented in a high-level programming language---typically some dialect of ML, such as Standard ML or OCaml.
 Invariants are maintained by the kernel judiciously exposing functionality through an API---offering functions to manipulate elements of certain abstract types---and through which untrusted code interacts with the kernel.
-Logical soundness is therefore contingent on the type soundness of the implementation meta-language's type-system.
+Logical soundness follows from the type soundness of the metalanguage.
 
 In contrast, desktop operating systems use hardware separation mechanisms to protect various \emph{kernel objects}---that is, book-keeping structures used by the kernel---from unfettered access by untrusted code.
-These hardware separation mechanisms include elevated \emph{privilege rings} (or \emph{exception levels})---granting the kernel access to protected hardware resources, such as system registers---and memory protection mechanisms through which the kernel can ``carve out'' a private scratch space to safely stash kernel objects.
+These hardware separation mechanisms include \emph{privilege rings} (or \emph{exception levels})---granting the kernel access to protected hardware resources, such as system registers---and memory protection mechanisms through which the kernel can ``carve out'' a private scratch space to safely stash kernel objects.
+The kernel offers services to untrusted user-space code via a binary interface, or an ABI, through the use of \emph{system calls} which invoke a context switch into the kernel.
+This system call ABI needs to be carefully designed, in order to avoid inadvertent security weaknesses.
+
+\emph{What, then, if we design theorem proving software to look less like typical theorem proving software, and more like a typical operating system?}
+
+In this paper, we introduce a prototype system---called \emph{Supervisionary}---to explore the consequences of this novel design.
+Supervisionary is implemented as a \emph{runtime} for a WebAssembly (or Wasm, henceforth) virtual machine.
+A runtime looks and acts like a kernel from the point-of-view of executing Wasm code: it possesses a monopoly on all system resources, and is manipulated and queried through an ABI that causes execution to synchronously switch into the runtime whenever invoked.
+Moreover, the runtime is separated from Wasm code via the use of distinct address spaces, with the runtime able to query and manipulate the Wasm heap directly but the runtime heap being kept inaccessible to Wasm.
+Immediately, there are two interesting consequences:
+
+\begin{enumerate}
+\item
+Supervisionary no longer has a \emph{metalanguage} in the typical understanding of that phrase.
+Rather, the Supervisionary kernel has an implementation language---Rust, in our case---that need not be a programming language with strong type-soundness guarantees.
+Logical soundness no longer follows from the type soundness of our implementation language, but rather follows from the separation of the kernel's address space from untrusted code, and our ABI design, which we will discuss further, below.
+\item
+The kernel and untrusted code are now decoupled.
+Untrusted code that directs proof search now need not be written in the same programming language as the kernel (or, interact with the kernel via shims written in that language).
+Rather, any programming language targeting the Wasm instruction set can be used, even a mix of them.
+\end{enumerate}
 

### 2021-06-25 paper/paper.tex
paper: finished introduction

@@
 The kernel is the only component within the system that need be trusted by a skeptic, as all proofs constructed by the system, no matter how large or complex, are ultimately passed through the kernel for authentication.
-This implementation strategy---known as the ``LCF-style'', after Milner's system that first implemented it---is a reliability mechanism, aiming to ensure that any theorem produced by the system is indeed a theorem.
-In short, the kernel acts as the \emph{root of all trust} within the system.
+This implementation strategy---known as the ``LCF-style'', after Milner's system that first implemented it---is a reliability mechanism, aiming to ensure that any purported theorem produced by the system is indeed a theorem.
+In short, the kernel acts as the \emph{root of all trust} within the wide theorem proving system.
 
-Shifting focus, we see an interesting analogy: desktop operating systems also implement a \emph{kernel}, which is the sole system component with unfettered access to hardware\footnote{Barring other privileged systems software, such as the hypervisor, which we consider as another form of operating system}, and provides useful abstractions, such as a filesystem, sockets, and so on, to untrusted user-space code.
+Shifting focus, we see an interesting similarity: desktop operating systems also implement a \emph{kernel}, which is the sole system component with unfettered access to hardware\footnote{Barring other privileged systems software---such as the hypervisor, separation kernels, security monitors, and so on ---which we consider just another form of operating system}, and provides useful abstractions, such as a filesystem, sockets, and so on, to untrusted user-space code.
 Security policies are enforced by the kernel using a variety of mechanisms, file and directory permissions, access control lists, and runtime-monitoring.
@@
 Supervisionary is implemented as a \emph{runtime} for a WebAssembly (or Wasm, henceforth) virtual machine.
-A runtime looks and acts like a kernel from the point-of-view of executing Wasm code: it possesses a monopoly on all system resources, and is manipulated and queried through an ABI that causes execution to synchronously switch into the runtime whenever invoked.
+A runtime looks and acts like a kernel from the point-of-view of executing Wasm code, in particular it is manipulated and queried through an ABI that causes execution to synchronously switch into the runtime whenever invoked.
 Moreover, the runtime is separated from Wasm code via the use of distinct address spaces, with the runtime able to query and manipulate the Wasm heap directly but the runtime heap being kept inaccessible to Wasm.
@@
 Untrusted code that directs proof search now need not be written in the same programming language as the kernel (or, interact with the kernel via shims written in that language).
-Rather, any programming language targeting the Wasm instruction set can be used, even a mix of them.
+Rather, any programming language targeting the Wasm instruction set can be used, or even a mixture of them.
+In this sense, Supervisionary again behaves like a real operating system kernel, where code written in any programming language is able to interact directly with the kernel, as long as it abides by its binary interface.
 \end{enumerate}
 
-\paragraph{Contributions}
-We make the following contributions in this paper:
+The Supervisionary ABI is also interesting in its own right as it differs markedly from the kernel API of a typical LCF system.
+Whilst all of the familiar objects of an implementation of Gordon's HOL---type-formers, types, terms, and theorems---are still present in Supervisionary, these are never \emph{directly} exposed to untrusted code for manipulation.
+Rather, the Supervisionary kernel maintains internal heaps of \emph{kernel objects}, indexed by \emph{handles}, which are akin to pointers.
+These handles are allocated and issued to untrusted code by the kernel upon creation of a new object.
+To manipulate or query an existing object, a handle is passed back to the kernel, which uses it to select an object to act on, on behalf of untrusted code.
+Interestingly, this allows us to provide an elegant formal specification of the Supervisionary ABI using Separation Logic, a modal logic used to reason about pointer-manipulation programs.
 
-\begin{enumerate}
-\item
-We design a novel proof-checking kernel for Gordon's HOL structured around a series of tables of kernel objects indexed by handles.
-With this, we introduce \emph{Supervisionary}, a prototype interactive theorem proving system which exposes its kernel as an ABI to WebAssembly programs.
-\item
-Noting that Supervisionary is little other than an exercise in pointer manipulation, we briefly explore how our ABI can be formally specified, using a variant of Separation Logic.
-\item
-We exploit Supervisionary's proof checking capabilities to define expressive dynamic security policies for WebAssembly programs.
-\end{enumerate}
+Here, we take an interlude and return to the subject of operating system design.
+Note that, for an operating system kernel to effectively act as the root of all trust within a system it must monopolise a system's resources.
+In particular, a program executing under the supervision of an operating system is unable to exhibit any side-effect\footnote{Other than non-termination, or generating heat by burning CPU cycles, naturally...} without the co-operation of the operating system kernel---all access to file-systems for reading and writing files, access to the network, and so on and so forth, must pass through the kernel, which permits or denies them depending on the prevailing security policy in force.
+A Wasm runtime like Supervisionary also has a monopoly on the system's resources from the point-of-view of executing Wasm code.
+However---as of the description above---the only resources it can access, create, or otherwise manipulate are kernel objects specific to the task of theorem proving in HOL.
+
+\emph{What if we grant Supervisionary a monopoly not just on authenticating theorems within HOL, but also on system resources such as the filesystem?}
+
+Now, Supervisionary starts to look \emph{very} different from a typical LCF system, as, whilst the latter has a monopoly on authenticating theorems, it cannot stand in the way of untrusted code within the system accessing the filesystem, network, or other resource, which is the role of the underlying operating system.
+On the other hand, Supervisionary can make access to wider system resources such as the filesystem contingent on untrusted code \emph{proving} a property at the point of access.
+Moreover, by varying the property to prove, Supervisionary can capture a wide range of expressive security and correctness policies.
+Interestingly, Supervisionary then straddles both static and dynamic verification techniques, with the system acting as both a proof assistant for HOL and a runtime verification mechanism for Wasm code.
 
@@
 In Section~\ref{sect.design.supervisionary.kernel} we describe the design of the Supervisionary kernel.
-In Section~\ref{sect.supervisionary.system} we describe the Supervisionary system proper, including untrusted support code---the analogue of Unix's \texttt{libc}---that allows WebAssembly programs to interact with the kernel.
+In Section~\ref{sect.supervisionary.system} we describe the Supervisionary system proper, including untrusted support code---the analogue of Unix's \texttt{libc}---that allows Wasm programs to interact with the kernel.
 In Section~\ref{sect.separation.logic.an.interlude} we briefly explore how Separation Logic can be used to provide a formal description of the action of the Supervisionary ABI.
-In Section~\ref{sect.security.policies.for.webassembly} we show how expressive security policies can be captured, and enforced, for arbitrary WebAssembly programs using Supervisionary.
+In Section~\ref{sect.security.policies.for.webassembly} we show how expressive security and correctness policies can be captured, and enforced, for arbitrary Wasm programs using Supervisionary.
 Finally, in Section~\ref{sect.conclusions} we conclude.

### 2021-06-25 paper/paper.tex
paper: finished off section on HOL

@@
 
-Supervisionary implements a version of Gordon's higher-order logic (or HOL, henceforth), a close relative of Church's Simple Theory of Types that extends Church's system with top-level polymorphism.
-We describe HOL here, introducing enough material for the reader to understand the remainder of the paper---motivating HOL or describing its expressiveness is outside of the scope of this paper.
+Supervisionary implements a version of Gordon's higher-order logic (or HOL, henceforth), a close relative of Church's Simple Theory of Types.
+We briefly describe HOL here.
+Motivating HOL or describing its expressiveness is beyond the scope of this paper.
 
@@
 Our version of HOL also includes two \emph{primitive definitional} principles for extending this theory with new defined constants and types, respectively.
-Specifically, for a previously-undefined constant $\mathtt{C}$ and a term $r : \tau$ where $fv(r) = \{\}$ the constant $\mathtt{C}$ with declared type $\tau$ is added to the theory, as is an additional axiom $\{\} \vdash \mathtt{C} = r$.
-TODO: type-definition
+
+For a previously-undefined constant $\mathtt{C}$ and a term $r : \tau$ where $fv(r) = \{\}$ the constant $\mathtt{C}$ with declared type $\tau$ is added to the theory, as is an additional axiom $\{\} \vdash \mathtt{C} = r$.
+
+New types are ``carved out'' of non-empty subsets of existing types.
+Given an existing type $\tau$ and a predicate $p : \tau \Rightarrow \mathtt{bool}$ we introduce a new type $\tau'$ whenever $\{\} \vdash \xsts{x_\tau}p(x_\tau)$ holds.
+In addition, two new constants are entered into the global theory---$\mathtt{in}_{\tau'}$ with type $\tau \Rightarrow \tau'$ and $\mathtt{out}_{\tau'}$ with type $\tau' \Rightarrow \tau$, respectively---representing \deffont{injection} and \deffont{projection} functions for the new type.
+These two functions are related by also entering $\{\} \vdash \fall{x_\tau}p(x_\tau) \longrightarrow out_{\tau'}(in_{\tau'}(x_\tau)) = x_\tau$ and $\{\} \vdash \fall{x_{\tau'}}in_{\tau'}(out_{\tau'}(x_{\tau'})) = x_{\tau'}$ as additional axioms.
 

### 2021-08-01 src/kernel/_type.rs
host-interface: more work on WASMI host interface

@@ -162,12 +162,12 @@ impl Type {
 lazy_static! {
     /// The "alpha" type-variable, `0`.
     pub static ref TYPE_ALPHA: Type = Type::Variable {
-        name: 0u64,
+        name: 0_u64,
     };
 
     /// The "beta" type-variable, `1`.
     pub static ref TYPE_BETA: Type = Type::Variable {
-        name: 1u64,
+        name: 1_u64,
     };
 
     /// The type of propositions, `Prop`.
@@ -237,7 +237,7 @@ mod test {
     /// discriminators, both in positive and negative forms.
     #[test]
     pub fn type_test0() {
-        assert!(Type::variable(0u64).is_variable());
+        assert!(Type::variable(0_u64).is_variable());
         assert!(
             Type::combination::<Handle<tags::TypeFormer>, Handle<tags::Type>>(
                 PREALLOCATED_HANDLE_TYPE_FORMER_PROP,
@@ -254,8 +254,8 @@ mod test {
                 .is_combination()
         );
 
-        assert!(!Type::variable(0u64).is_combination());
-        assert!(!Type::variable(0u64).is_function());
+        assert!(!Type::variable(0_u64).is_combination());
+        assert!(!Type::variable(0_u64).is_function());
         assert!(
             !Type::combination::<Handle<tags::TypeFormer>, Handle<tags::Type>>(
                 PREALLOCATED_HANDLE_TYPE_FORMER_PROP,
@@ -272,8 +272,8 @@ mod test {
     /// Tests that splitting a variable gets you back to where you started.
     #[test]
     pub fn type_test1() {
-        let v = Type::variable(0u64);
-        assert_eq!(v.split_variable(), Some(&0u64));
+        let v = Type::variable(0_u64);
+        assert_eq!(v.split_variable(), Some(&0_u64));
     }
 
     /// Tests that splitting a combination gets you back to where you started.

### 2021-08-01 src/kernel/kernel_panic.rs

@@ -13,41 +13,6 @@
 //! [Dominic Mulligan]: https://dominic-mulligan.co.uk
 //! [Arm Research]: http://www.arm.com/research
 
-use log::{error, info};
-
-////////////////////////////////////////////////////////////////////////////////
-// Kernel logging.
-////////////////////////////////////////////////////////////////////////////////
-
-/// Kernel logging for information cases.
-#[inline]
-pub fn kernel_info<T>(message: T)
-where
-    T: AsRef<str>,
-{
-    info!("[KERNEL]: {}", message.as_ref())
-}
-
-/// Kernel logging for non-fatal error cases.
-#[inline]
-pub fn kernel_error<T>(message: T)
-where
-    T: AsRef<str>,
-{
-    error!("[KERNEL]: {}", message.as_ref())
-}
-
-/// Kernel logging for fatal cases: logs the provided error on the kernel's
-/// error log, then panics with the same error message.
-#[inline]
-pub fn kernel_panic<T>(message: T) -> !
-where
-    T: AsRef<str>,
-{
-    error!("[KERNEL]: {}", message.as_ref());
-    panic!()
-}
-
 ////////////////////////////////////////////////////////////////////////////////
 // Kernel panic messages.
 ////////////////////////////////////////////////////////////////////////////////

### 2021-08-01 src/kernel/name.rs

@@ -17,7 +17,7 @@
 // Miscellaneous material.
 ////////////////////////////////////////////////////////////////////////////////
 
-use crate::kernel::kernel_panic::{kernel_info, kernel_panic};
+use log::info;
 
 ////////////////////////////////////////////////////////////////////////////////
 // Names and related material.
@@ -40,10 +40,10 @@ where
             if let Some(next) = counter.checked_add(1) {
                 counter = next;
             } else {
-                kernel_panic("Exhausted fresh name generation.");
+                panic!("Exhausted fresh name generation.");
             }
         } else {
-            kernel_info(format!("Fresh name generated: {}.", counter));
+            info!("Fresh name generated: {}.", counter);
             return counter;
         }
     }

### 2021-08-01 src/kernel/runtime_state.rs

@@ -15,6 +15,8 @@
 //! [Dominic Mulligan]: https://dominic-mulligan.co.uk
 //! [Arm Research]: http://www.arm.com/research
 
+use log::{info, error};
+
 use crate::kernel::{
     _type::{
         Type, TYPE_ALPHA, TYPE_BETA, TYPE_BINARY_CONNECTIVE, TYPE_POLYMORPHIC_BINARY_PREDICATE,
@@ -40,8 +42,7 @@ use crate::kernel::{
         PREALLOCATED_HANDLE_TYPE_UNARY_PREDICATE, PREALLOCATED_HANDLE_UPPER_BOUND,
     },
     kernel_panic::{
-        kernel_error, kernel_info, kernel_panic, DANGLING_HANDLE_ERROR, HANDLE_EXHAUST_ERROR,
-        PRIMITIVE_CONSTRUCTION_ERROR,
+        DANGLING_HANDLE_ERROR, HANDLE_EXHAUST_ERROR, PRIMITIVE_CONSTRUCTION_ERROR,
     },
     name::Name,
     term::{
@@ -51,8 +52,7 @@ use crate::kernel::{
     },
     theorem::Theorem,
 };
-use std::fmt::Debug;
-use std::{borrow::Borrow, collections::HashMap, fmt::Display, iter::FromIterator};
+use std::{borrow::Borrow, collections::HashMap, fmt::{Debug, Display}, iter::FromIterator};
 
 ////////////////////////////////////////////////////////////////////////////////
 // The runtime state.
@@ -105,11 +105,11 @@ impl RuntimeState {
         let next = self.next_handle;
 
         match self.next_handle.checked_add(1) {
-            None => kernel_panic(HANDLE_EXHAUST_ERROR),
+            None => panic!(HANDLE_EXHAUST_ERROR),
             Some(next) => self.next_handle = next,
         }
 
-        kernel_info(format!("Generating fresh handle: {}.", next));
+        info!("Generating fresh handle: {}.", next);
 
         return Handle::from(next);
     }
@@ -124,10 +124,10 @@ impl RuntimeState {
     where
         T: Into<usize> + Clone,
     {
-        kernel_info(format!(
+        info!(
             "Registering new type-former with arity: {}.",
             arity.clone().into()
-        ));
+        );
         let handle = self.issue_handle();
         self.type_formers.insert(handle.clone(), arity.into());
         handle
@@ -140,10 +140,10 @@ impl RuntimeState {
     where
[…]

### 2021-08-01 src/kernel/term.rs

@@ -20,9 +20,9 @@ use crate::kernel::{
         PREALLOCATED_HANDLE_CONSTANT_EXISTS, PREALLOCATED_HANDLE_CONSTANT_FALSE,
         PREALLOCATED_HANDLE_CONSTANT_FORALL, PREALLOCATED_HANDLE_CONSTANT_IMPLICATION,
         PREALLOCATED_HANDLE_CONSTANT_NEGATION, PREALLOCATED_HANDLE_CONSTANT_TRUE,
-        PREALLOCATED_HANDLE_TERM_DISJUNCTION, PREALLOCATED_HANDLE_TYPE_BINARY_CONNECTIVE,
-        PREALLOCATED_HANDLE_TYPE_BINARY_PREDICATE, PREALLOCATED_HANDLE_TYPE_PROP,
-        PREALLOCATED_HANDLE_TYPE_QUANTIFIER, PREALLOCATED_HANDLE_TYPE_UNARY_CONNECTIVE,
+        PREALLOCATED_HANDLE_TYPE_BINARY_CONNECTIVE, PREALLOCATED_HANDLE_TYPE_BINARY_PREDICATE,
+        PREALLOCATED_HANDLE_TYPE_PROP, PREALLOCATED_HANDLE_TYPE_QUANTIFIER,
+        PREALLOCATED_HANDLE_TYPE_UNARY_CONNECTIVE,
     },
     name::Name,
 };

### 2021-08-01 src/wasmi/host_interface.rs

@@ -106,11 +106,11 @@ impl AbiType {
 /* Type-former related calls. */
 
 /// The name of the `TypeFormer.Resolve` ABI call.
-pub const ABI_TYPE_FORMER_RESOLVE_NAME: &'static str = "__type_former_resolve";
+pub const ABI_TYPE_FORMER_RESOLVE_NAME: &str = "__type_former_resolve";
 /// The name of the `TypeFormer.IsRegistered` ABI call.
-pub const ABI_TYPE_FORMER_IS_REGISTERED_NAME: &'static str = "__type_former_is_registered";
+pub const ABI_TYPE_FORMER_IS_REGISTERED_NAME: &str = "__type_former_is_registered";
 /// The name of the `TypeFormer.Register` ABI call.
-pub const ABI_TYPE_FORMER_REGISTER_NAME: &'static str = "__type_former_register";
+pub const ABI_TYPE_FORMER_REGISTER_NAME: &str = "__type_former_register";
 
 /// The host-call number of the `TypeFormer.Resolve` ABI call.
 pub const ABI_TYPE_FORMER_RESOLVE_INDEX: usize = 0;
@@ -122,32 +122,32 @@ pub const ABI_TYPE_FORMER_REGISTER_INDEX: usize = 2;
 /* Type-related calls. */
 
 /// The name of the `Type.IsRegistered` ABI call.
-pub const ABI_TYPE_IS_REGISTERED_NAME: &'static str = "__type_is_registered";
+pub const ABI_TYPE_IS_REGISTERED_NAME: &str = "__type_is_registered";
 /// The name of the `Type.Register.Variable` ABI call.
-pub const ABI_TYPE_REGISTER_VARIABLE_NAME: &'static str = "__type_register_variable";
+pub const ABI_TYPE_REGISTER_VARIABLE_NAME: &str = "__type_register_variable";
 /// The name of the `Type.Register.Combination` ABI call.
-pub const ABI_TYPE_REGISTER_COMBINATION_NAME: &'static str = "__type_register_combination_name";
+pub const ABI_TYPE_REGISTER_COMBINATION_NAME: &str = "__type_register_combination_name";
 /// The name of the `Type.Register.Function` ABI call.
-pub const ABI_TYPE_REGISTER_FUNCTION_NAME: &'static str = "__type_register_function_name";
+pub const ABI_TYPE_REGISTER_FUNCTION_NAME: &str = "__type_register_function_name";
 
[…]

### 2021-08-01 src/kernel/_type.rs
host-interface: more work on filling out Theorem.* section of host interface

@@ -15,9 +15,11 @@
 
 use crate::kernel::{
     handle::{
-        tags, Handle, PREALLOCATED_HANDLE_TYPE_ALPHA, PREALLOCATED_HANDLE_TYPE_FORMER_ARROW,
+        tags, Handle, PREALLOCATED_HANDLE_TYPE_ALPHA,
+        PREALLOCATED_HANDLE_TYPE_FORMER_ARROW,
         PREALLOCATED_HANDLE_TYPE_FORMER_PROP, PREALLOCATED_HANDLE_TYPE_PROP,
-        PREALLOCATED_HANDLE_TYPE_UNARY_CONNECTIVE, PREALLOCATED_HANDLE_TYPE_UNARY_PREDICATE,
+        PREALLOCATED_HANDLE_TYPE_UNARY_CONNECTIVE,
+        PREALLOCATED_HANDLE_TYPE_UNARY_PREDICATE,
     },
     name::Name,
 };
@@ -90,7 +92,10 @@ impl Type {
     where
         T: Into<Handle<tags::Type>> + Clone,
     {
-        Type::combination(PREALLOCATED_HANDLE_TYPE_FORMER_ARROW, vec![domain, range])
+        Type::combination(
+            PREALLOCATED_HANDLE_TYPE_FORMER_ARROW,
+            vec![domain, range],
+        )
     }
 
     /// Returns `Some(name)` iff the type is a type-variable with name, `name`.
@@ -117,7 +122,9 @@ impl Type {
 
     /// Returns `Some((dom, rng))` iff the type is a function type between
     /// domain type `dom` and range type `rng`.
-    pub fn split_function(&self) -> Option<(&Handle<tags::Type>, &Handle<tags::Type>)> {
+    pub fn split_function(
+        &self,
+    ) -> Option<(&Handle<tags::Type>, &Handle<tags::Type>)> {
         let (handle, args) = self.split_combination()?;
 
         if handle == &PREALLOCATED_HANDLE_TYPE_FORMER_ARROW && args.len() == 2 {
@@ -229,7 +236,8 @@ mod test {
         _type::Type,
         handle::{
             tags, Handle, PREALLOCATED_HANDLE_TYPE_FORMER_ARROW,
-            PREALLOCATED_HANDLE_TYPE_FORMER_PROP, PREALLOCATED_HANDLE_TYPE_PROP,
+            PREALLOCATED_HANDLE_TYPE_FORMER_PROP,
[…]

### 2021-08-01 src/kernel/error_code.rs

@@ -111,13 +111,23 @@ impl Display for ErrorCode {
             ErrorCode::Success => write!(f, "Success"),
             ErrorCode::SignatureFailure => write!(f, "SignatureFailure"),
             ErrorCode::NoSuchFunction => write!(f, "NoSuchFunction"),
-            ErrorCode::NoSuchConstantRegistered => write!(f, "NoSuchConstantRegistered"),
-            ErrorCode::NoSuchTermRegistered => write!(f, "NoSuchTermRegistered"),
-            ErrorCode::NoSuchTheoremRegistered => write!(f, "NoSuchTheoremRegistered"),
-            ErrorCode::NoSuchTypeFormerRegistered => write!(f, "NoSuchTypeFormerRegistered"),
+            ErrorCode::NoSuchConstantRegistered => {
+                write!(f, "NoSuchConstantRegistered")
+            }
+            ErrorCode::NoSuchTermRegistered => {
+                write!(f, "NoSuchTermRegistered")
+            }
+            ErrorCode::NoSuchTheoremRegistered => {
+                write!(f, "NoSuchTheoremRegistered")
+            }
+            ErrorCode::NoSuchTypeFormerRegistered => {
+                write!(f, "NoSuchTypeFormerRegistered")
+            }
             ErrorCode::MismatchedArity => write!(f, "MismatchedArity"),
             ErrorCode::DomainTypeMismatch => write!(f, "DomainTypeMismatch"),
-            ErrorCode::NoSuchTypeRegistered => write!(f, "NoSuchTypeRegistered"),
+            ErrorCode::NoSuchTypeRegistered => {
+                write!(f, "NoSuchTypeRegistered")
+            }
             ErrorCode::NotAFunctionType => write!(f, "NotAFunctionType"),
             ErrorCode::NotATypeCombination => write!(f, "NotATypeCombination"),
             ErrorCode::NotATypeVariable => write!(f, "NotATypeVariable"),
@@ -136,7 +146,9 @@ impl Display for ErrorCode {
             ErrorCode::NotAVariable => write!(f, "NotAVariable"),
             ErrorCode::TermNotWellformed => write!(f, "TermNotWellformed"),
             ErrorCode::ShapeMismatch => write!(f, "ShapeMismatch"),
-            ErrorCode::TheoremNotWellformed => write!(f, "TheoremNotWellformed"),
[…]

### 2021-08-01 src/kernel/handle.rs

@@ -93,15 +93,17 @@ where
 ////////////////////////////////////////////////////////////////////////////////
 
 /// A pre-allocated handle used to refer to the `Prop` type-former.
-pub const PREALLOCATED_HANDLE_TYPE_FORMER_PROP: Handle<tags::TypeFormer> = Handle {
-    handle: 0,
-    marker: PhantomData,
-};
+pub const PREALLOCATED_HANDLE_TYPE_FORMER_PROP: Handle<tags::TypeFormer> =
+    Handle {
+        handle: 0,
+        marker: PhantomData,
+    };
 /// A pre-allocated handle used to refer to the function-space type-former.
-pub const PREALLOCATED_HANDLE_TYPE_FORMER_ARROW: Handle<tags::TypeFormer> = Handle {
-    handle: 1,
-    marker: PhantomData,
-};
+pub const PREALLOCATED_HANDLE_TYPE_FORMER_ARROW: Handle<tags::TypeFormer> =
+    Handle {
+        handle: 1,
+        marker: PhantomData,
+    };
 /// A pre-allocated handle used to refer to the type-variable `A`.
 pub const PREALLOCATED_HANDLE_TYPE_ALPHA: Handle<tags::Type> = Handle {
     handle: 2,
@@ -118,25 +120,29 @@ pub const PREALLOCATED_HANDLE_TYPE_PROP: Handle<tags::Type> = Handle {
     marker: PhantomData,
 };
 /// A pre-allocated handle used to refer to the type of unary predicates.
-pub const PREALLOCATED_HANDLE_TYPE_UNARY_PREDICATE: Handle<tags::Type> = Handle {
-    handle: 5,
-    marker: PhantomData,
-};
+pub const PREALLOCATED_HANDLE_TYPE_UNARY_PREDICATE: Handle<tags::Type> =
+    Handle {
+        handle: 5,
+        marker: PhantomData,
+    };
 /// A pre-allocated handle used to refer to the type of binary predicates.
-pub const PREALLOCATED_HANDLE_TYPE_BINARY_PREDICATE: Handle<tags::Type> = Handle {
-    handle: 6,
-    marker: PhantomData,
-};
+pub const PREALLOCATED_HANDLE_TYPE_BINARY_PREDICATE: Handle<tags::Type> =
+    Handle {
+        handle: 6,
+        marker: PhantomData,
+    };
 /// A pre-allocated handle used to refer to the type of unary connectives.
[…]

### 2021-08-01 src/kernel/kernel_panic.rs

@@ -19,12 +19,14 @@
 
 /// Error message produced during a kernel panic due to the kernel running out
 /// of fresh handles.
-pub const HANDLE_EXHAUST_ERROR: &str = "Kernel invariant failed: handles have been exhausted";
+pub const HANDLE_EXHAUST_ERROR: &str =
+    "Kernel invariant failed: handles have been exhausted";
 
 /// Error message produced during a kernel panic due to the kernel encountering
 /// a registered kernel-object with a dangling handle.
 
-pub const DANGLING_HANDLE_ERROR: &str = "Kernel invariant failed: dangling handle.";
+pub const DANGLING_HANDLE_ERROR: &str =
+    "Kernel invariant failed: dangling handle.";
 
 /// Error message produced during a kernel panic due to the kernel failing to
 /// build a kernel primitive.

### 2021-08-01 src/kernel/runtime_state.rs

@@ -15,44 +15,60 @@
 //! [Dominic Mulligan]: https://dominic-mulligan.co.uk
 //! [Arm Research]: http://www.arm.com/research
 
-use log::{info, error};
+use log::{error, info};
 
 use crate::kernel::{
     _type::{
-        Type, TYPE_ALPHA, TYPE_BETA, TYPE_BINARY_CONNECTIVE, TYPE_POLYMORPHIC_BINARY_PREDICATE,
-        TYPE_POLYMORPHIC_QUANTIFIER, TYPE_POLYMORPHIC_UNARY_PREDICATE, TYPE_PROP,
-        TYPE_UNARY_CONNECTIVE,
+        Type, TYPE_ALPHA, TYPE_BETA, TYPE_BINARY_CONNECTIVE,
+        TYPE_POLYMORPHIC_BINARY_PREDICATE, TYPE_POLYMORPHIC_QUANTIFIER,
+        TYPE_POLYMORPHIC_UNARY_PREDICATE, TYPE_PROP, TYPE_UNARY_CONNECTIVE,
     },
     error_code::ErrorCode,
     handle::{
         tags, Handle, PREALLOCATED_HANDLE_CONSTANT_CONJUNCTION,
-        PREALLOCATED_HANDLE_CONSTANT_DISJUNCTION, PREALLOCATED_HANDLE_CONSTANT_EQUALITY,
-        PREALLOCATED_HANDLE_CONSTANT_EXISTS, PREALLOCATED_HANDLE_CONSTANT_FALSE,
-        PREALLOCATED_HANDLE_CONSTANT_FORALL, PREALLOCATED_HANDLE_CONSTANT_IMPLICATION,
-        PREALLOCATED_HANDLE_CONSTANT_NEGATION, PREALLOCATED_HANDLE_CONSTANT_TRUE,
-        PREALLOCATED_HANDLE_TERM_CONJUNCTION, PREALLOCATED_HANDLE_TERM_DISJUNCTION,
+        PREALLOCATED_HANDLE_CONSTANT_DISJUNCTION,
+        PREALLOCATED_HANDLE_CONSTANT_EQUALITY,
+        PREALLOCATED_HANDLE_CONSTANT_EXISTS,
+        PREALLOCATED_HANDLE_CONSTANT_FALSE,
+        PREALLOCATED_HANDLE_CONSTANT_FORALL,
[…]

### 2021-08-01 src/kernel/term.rs

@@ -16,11 +16,16 @@
 use crate::kernel::{
     handle::{
         tags, Handle, PREALLOCATED_HANDLE_CONSTANT_CONJUNCTION,
-        PREALLOCATED_HANDLE_CONSTANT_DISJUNCTION, PREALLOCATED_HANDLE_CONSTANT_EQUALITY,
-        PREALLOCATED_HANDLE_CONSTANT_EXISTS, PREALLOCATED_HANDLE_CONSTANT_FALSE,
-        PREALLOCATED_HANDLE_CONSTANT_FORALL, PREALLOCATED_HANDLE_CONSTANT_IMPLICATION,
-        PREALLOCATED_HANDLE_CONSTANT_NEGATION, PREALLOCATED_HANDLE_CONSTANT_TRUE,
-        PREALLOCATED_HANDLE_TYPE_BINARY_CONNECTIVE, PREALLOCATED_HANDLE_TYPE_BINARY_PREDICATE,
+        PREALLOCATED_HANDLE_CONSTANT_DISJUNCTION,
+        PREALLOCATED_HANDLE_CONSTANT_EQUALITY,
+        PREALLOCATED_HANDLE_CONSTANT_EXISTS,
+        PREALLOCATED_HANDLE_CONSTANT_FALSE,
+        PREALLOCATED_HANDLE_CONSTANT_FORALL,
+        PREALLOCATED_HANDLE_CONSTANT_IMPLICATION,
+        PREALLOCATED_HANDLE_CONSTANT_NEGATION,
+        PREALLOCATED_HANDLE_CONSTANT_TRUE,
+        PREALLOCATED_HANDLE_TYPE_BINARY_CONNECTIVE,
+        PREALLOCATED_HANDLE_TYPE_BINARY_PREDICATE,
         PREALLOCATED_HANDLE_TYPE_PROP, PREALLOCATED_HANDLE_TYPE_QUANTIFIER,
         PREALLOCATED_HANDLE_TYPE_UNARY_CONNECTIVE,
     },
@@ -155,7 +160,9 @@ impl Term {
     /// handle pointing to a type, `opt`.  If `opt` is `None` then the constant
     /// has the type registered in the constant-table under the handle,
     /// `handle`.
-    pub fn split_constant(&self) -> Option<(&Handle<tags::Constant>, &Handle<tags::Type>)> {
+    pub fn split_constant(
+        &self,
+    ) -> Option<(&Handle<tags::Constant>, &Handle<tags::Type>)> {
[…]

### 2021-08-01 src/wasmi/host_interface.rs

@@ -25,8 +25,9 @@ use std::{
 
 use byteorder::{ByteOrder, LittleEndian};
 use wasmi::{
-    Error as WasmiError, Externals, FuncInstance, FuncRef, HostError, MemoryInstance,
-    ModuleImportResolver, RuntimeArgs, RuntimeValue, Signature, Trap, TrapKind, ValueType,
+    Error as WasmiError, Externals, FuncInstance, FuncRef, HostError,
+    MemoryInstance, ModuleImportResolver, RuntimeArgs, RuntimeValue, Signature,
+    Trap, TrapKind, ValueType,
 };
 
 use crate::kernel::{
@@ -48,10 +49,6 @@ mod semantic_types {
     pub type Pointer = u32;
     /// An arity of a type-former.
     pub type Arity = u64;
-    /// A Boolean value.
-    pub type Boolean = u32;
-    /// A kernel error code.
-    pub type ErrorCode = u32;
     /// A handle to a kernel object.
     pub type Handle = u64;
     /// A name of a variable (e.g. a lambda-abstracted variable, or
@@ -108,7 +105,8 @@ impl AbiType {
 /// The name of the `TypeFormer.Resolve` ABI call.
 pub const ABI_TYPE_FORMER_RESOLVE_NAME: &str = "__type_former_resolve";
 /// The name of the `TypeFormer.IsRegistered` ABI call.
-pub const ABI_TYPE_FORMER_IS_REGISTERED_NAME: &str = "__type_former_is_registered";
+pub const ABI_TYPE_FORMER_IS_REGISTERED_NAME: &str =
+    "__type_former_is_registered";
 /// The name of the `TypeFormer.Register` ABI call.
 pub const ABI_TYPE_FORMER_REGISTER_NAME: &str = "__type_former_register";
 
@@ -126,14 +124,17 @@ pub const ABI_TYPE_IS_REGISTERED_NAME: &str = "__type_is_registered";
 /// The name of the `Type.Register.Variable` ABI call.
 pub const ABI_TYPE_REGISTER_VARIABLE_NAME: &str = "__type_register_variable";
 /// The name of the `Type.Register.Combination` ABI call.
-pub const ABI_TYPE_REGISTER_COMBINATION_NAME: &str = "__type_register_combination_name";
[…]

### 2021-08-02 paper/paper.tex
paper: minor changes to paper

@@
 \usepackage{microtype}
+\usepackage{tikz}
 \usepackage{prooftree}
+
+\usetikzlibrary{matrix,backgrounds}
 
@@
 \end{prooftree}
-\qquad
+\\[1ex]
 \begin{prooftree}
@@
 \end{prooftree}
-\\[1ex]
+\qquad
 \begin{prooftree}
@@
 
-New types are ``carved out'' of non-empty subsets of existing types.
-Given an existing type $\tau$ and a predicate $p : \tau \Rightarrow \mathtt{bool}$ we introduce a new type $\tau'$ whenever $\{\} \vdash \xsts{x_\tau}p(x_\tau)$ holds.
-In addition, two new constants are entered into the global theory---$\mathtt{in}_{\tau'}$ with type $\tau \Rightarrow \tau'$ and $\mathtt{out}_{\tau'}$ with type $\tau' \Rightarrow \tau$, respectively---representing \deffont{injection} and \deffont{projection} functions for the new type.
-These two functions are related by also entering $\{\} \vdash \fall{x_\tau}p(x_\tau) \longrightarrow out_{\tau'}(in_{\tau'}(x_\tau)) = x_\tau$ and $\{\} \vdash \fall{x_{\tau'}}in_{\tau'}(out_{\tau'}(x_{\tau'})) = x_{\tau'}$ as additional axioms.
+A new type $\tau$ is ``carved out'' of an existing type $\tau'$ by identifying the non-empty collection of elements of $\tau'$ that belong to $\tau$.
+Specifically, given a predicate $p : \tau \Rightarrow \mathtt{bool}$ we introduce a new type $\tau'$ whenever we can show $\{\} \vdash \xsts{x_\tau}p(x_\tau)$ holds.
+In addition, two new constants are also entered into the global theory---$\mathtt{in}_{\tau'}$ with type $\tau \Rightarrow \tau'$ and $\mathtt{out}_{\tau'}$ with type $\tau' \Rightarrow \tau$, respectively---representing \deffont{injection} and \deffont{projection} functions for the new type.
+These two functions are related by a pair of additional axioms, $\{\} \vdash \fall{x_\tau}p(x_\tau) \longrightarrow out_{\tau'}(in_{\tau'}(x_\tau)) = x_\tau$ and $\{\} \vdash \fall{x_{\tau'}}in_{\tau'}(out_{\tau'}(x_{\tau'})) = x_{\tau'}$, that are added to the global theory.
 
@@
 
+In this section, we describe the design and implementation of the Supervisionary kernel.
+We use a semi-formal approach, adopting a dialect of Separation Logic to describe the action of the kernel ABI.
+We begin by introducing some foundational material that will be used in the remainder of the section.
+
+\subsection{Preliminary material}
+
+\begin{definition}
+\label{defn.kernel.handles}
+We fix a countably infinite set of \deffont{handles}.
+We use $h$, $i$, $j$, and so on, to range arbitrarily over handles.
+\end{definition}
+
+\begin{definition}
+\label{defn.partial.function}
+Write $f : A \rightharpoonup B$ to denote a \deffont{partial function} from $A$ to $B$.
+\end{definition}
+
+We write $dom(f)$ for the set $\{ x \mid f(x) \text{ is defined} \}$.
+We write $\emptyset$ for the \deffont{empty} partial function where $dom(\emptyset) = \{\}$ and write $a \mapsto b$ for the \deffont{singleton} partial function where $dom(a \mapsto b) = \{a\}$.
+
+If $dom(f) \cap dom(g) = \{\}$ we write $f \uplus g$ for the partial function where $(f \uplus g)(x) = f(x)$ if $x \in dom(f)$, $(f \uplus g)(x) = g(x)$ if $x \in dom(g)$, and undefined otherwise.
+Note that $f \uplus g$ is well-defined, owing to the assumption that $dom(f) \cap dom(g) = \{\}$.
+
+\subsection{The kernel ABI}
+
 The Supervisionary kernel maintains a mutable runtime state consisting of a series of mutable \deffont{heaps}.
-These heaps are indexed by \deffont{kernel handles}---unsigned integers generated by the kernel---and which can be thought of as pointers to \deffont{kernel objects} in the various heaps.
+These heaps are indexed by \deffont{kernel handles} which can be thought of as pointers to \deffont{kernel objects} stored in the heaps.
 Importantly, the kernel never exposes kernel objects to untrusted code directly, but performs all manipulations on these objects itself, under instruction by untrusted code.
+
+Note that in our implementation of Supervisionary handles are implemented as unsigned integers.
+These are generated fresh by the kernel.
+Client code may compare handles for equality and inequality, but the order in which handles are generated is undefined.
 
@@
 The simplest---and most foundational---of these additional heaps is the \deffont{heap of type-formers}, which records the arities of type-formers registered with the kernel.
+
 When untrusted code requests that a new type-former is registered with the kernel, it calls an ABI function---\texttt{TypeFormer.Handle.Register}---with a supplied arity.
@@
 \begin{figure}[t]
-\caption{The internal structure of the kernel's type heap}
+\caption{The internal structure of the kernel's type-former and type heap}
 \label{fig.type.heap}
@@
 \label{sect.separation.logic.an.interlude}
+
+Separation Logic is a modal logic typically used to reason about pointer-programs and dynamic memory allocation.
+In this section, we sketch how the Supervisionary kernel ABI can be formally specified using Separation Logic.
+
+\begin{definition}
+\label{defn.type.kernel.objects}
+Define \deffont{type kernel objects} by the following grammar:
+\begin{gather*}
+\tau, \tau', \tau'' ::= \mathtt{TV}(n) \mid \mathtt{TC}(h, h_1, \ldots, h_n)
+\end{gather*}
+\end{definition}
+
+Here, $\mathtt{TV}(n)$ represents a HOL type-variable with name $n$, whilst $\mathtt{TC}(h, h_1, \ldots, h_n)$ represents a HOL compound type, wherein $h$ points-to a registered type-former and $h_i$, for $1 \leq i \leq n$, point-to type arguments.
+
+\begin{definition}
+\label{defn.term.kernel.objects}
+Define \deffont{term kernel objects} by the following grammar:
+\begin{gather*}
+r,s,t ::= \mathtt{V}(n, h) \mid \mathtt{C}(h, i) \mid \mathtt{A}(h, i) \mid \mathtt{L}(n, h, i)
+\end{gather*}
+\end{definition}
+
+\noindent
+Intuitively, we have:
+\begin{itemize}
+\item
+$\mathtt{V}(n, h)$ represents a HOL term variable, with name $n$, wherein $h$ points-to the type of a variable.
+\item
+$\mathtt{C}(h, i)$ represents a HOL term constant, wherein $h$ points-to a registered constant, and $i$ points-to a type.
+\item
+$\mathtt{A}(h, i)$ represents the application of a functional term to an argument, wherein both $h$ and $i$ point-to terms.
+\item
+$\mathtt{L}(n, h, i)$ represents a $\lambda$-abstraction, with bound name $n$, wherein $h$ points-to a type and $i$ points-to a term.
+\end{itemize}
+
+\begin{definition}
+\label{defn.theorem.kernel.objects}
+Define \deffont{theorem kernel objects} by the following grammar:
+\begin{gather*}
+t, u, v ::= \mathtt{Thm}(h_1, \ldots, h_n, h)
+\end{gather*}
+\end{definition}
+
+Here, $\mathtt{Thm}(h_1, \ldots, h_n, h)$ represents a HOL theorem, wherein $h_i$, for $1 \leq i \leq n$, and $h$ point-to terms---the hypotheses and the conclusions of the theorem, respectively.
+
+\begin{definition}
+\label{defn.heaps}
+We define the type-former heap 
+\end{definition}
 

### 2021-08-03 src/kernel/runtime_state.rs
host-interface: more work on host interface for theorems

@@ -2842,7 +2842,7 @@ impl RuntimeState {
         Ok(self.admit_theorem(Theorem::new(hypotheses, conclusion)))
     }
 
-    pub fn register_implication_introduction_theorem<T, U>(
+    pub fn theorem_register_implication_introduction<T, U>(
         &mut self,
         handle: T,
         intro: U,
@@ -2878,7 +2878,7 @@ impl RuntimeState {
         Ok(self.admit_theorem(Theorem::new(hypotheses, conclusion)))
     }
 
-    pub fn register_implication_elimination_theorem<T>(
+    pub fn theorem_register_implication_elimination<T>(
         &mut self,
         left: T,
         right: T,

### 2021-08-03 src/wasmi/host_interface.rs

@@ -1990,6 +1990,73 @@ impl WasmiRuntimeState {
             )
     }
 
+    /// Lifting of the `theorem_register_negation_introduction` function.
+    #[inline]
+    fn theorem_register_negation_introduction<T, U>(
+        &self,
+        theorem_handle: T,
+        term_handle: U,
+    ) -> Result<Handle<tags::Theorem>, KernelErrorCode>
+    where
+        T: Borrow<Handle<tags::Theorem>>,
+        U: Into<Handle<tags::Term>> + Clone,
+    {
+        self.kernel
+            .borrow_mut()
+            .theorem_register_negation_introduction(theorem_handle, term_handle)
+    }
+
+    /// Lifting of the `theorem_register_negation_elimination` function.
+    #[inline]
+    fn theorem_register_negation_introduction<T, U>(
+        &self,
+        left_handle: T,
+        right_handle: U,
+    ) -> Result<Handle<tags::Theorem>, KernelErrorCode>
+    where
+        T: Borrow<Handle<tags::Theorem>>,
+        U: Borrow<Handle<tags::Theorem>>,
+    {
+        self.kernel
+            .borrow_mut()
+            .theorem_register_negation_elimination(theorem_handle, term_handle)
+    }
+
+    /// Lifting of the `theorem_register_implication_introduction` function.
+    #[inline]
+    fn theorem_register_implication_introduction<T, U>(
+        &self,
+        theorem_handle: T,
+        term_handle: U,
+    ) -> Result<Handle<tags::Theorem>, KernelErrorCode>
+    where
+        T: Borrow<Handle<tags::Theorem>>,
+        U: Into<Handle<tags::Term>> + Clone,
+    {
+        self.kernel
+            .borrow_mut()
+            .theorem_register_implication_introduction(
+                theorem_handle,
+                term_handle,
+            )
+    }
+
+    /// Lifting of the `theorem_register_implication_elimination` function.
+    #[inline]
+    fn theorem_register_implication_elimination<T, U>(
+        &self,
+        left_handle: T,
+        right_handle: U,
+    ) -> Result<Handle<tags::Theorem>, KernelErrorCode>
+    where
+        T: Borrow<Handle<tags::Theorem>>,
+        U: Borrow<Handle<tags::Theorem>>,
+    {
+        self.kernel
+            .borrow_mut()
+            .theorem_register_implication_elimination(left_handle, right_handle)
+    }
+
[…]

### 2021-08-03 src/kernel/runtime_state.rs
host-interface: finished runtime state integration for Theorem.*

@@ -2541,13 +2541,14 @@ impl RuntimeState {
         Ok(self.admit_theorem(Theorem::new(hypotheses, conclusion)))
     }
 
-    pub fn register_equality_introduction_theorem<T>(
+    pub fn theorem_register_iff_introduction<T, U>(
         &mut self,
         left: T,
-        right: T,
+        right: U,
     ) -> Result<Handle<tags::Theorem>, ErrorCode>
     where
         T: Borrow<Handle<tags::Theorem>>,
+        U: Borrow<Handle<tags::Theorem>>,
     {
         let left = self
             .resolve_theorem_handle(left)
@@ -2579,7 +2580,7 @@ impl RuntimeState {
         Ok(self.admit_theorem(Theorem::new(hypotheses, conclusion)))
     }
 
-    pub fn register_equality_elimination_theorem<T>(
+    pub fn theorem_register_iff_left_elimination<T>(
         &mut self,
         handle: T,
     ) -> Result<Handle<tags::Theorem>, ErrorCode>
@@ -2785,14 +2786,16 @@ impl RuntimeState {
         Ok(self.admit_theorem(Theorem::new(hypotheses, conclusion)))
     }
 
-    pub fn register_disjunction_elimination_theorem<T>(
+    pub fn theorem_register_disjunction_elimination<T, U, V>(
         &mut self,
         left: T,
-        mid: T,
-        right: T,
+        mid: U,
+        right: V,
     ) -> Result<Handle<tags::Theorem>, ErrorCode>
     where
         T: Borrow<Handle<tags::Theorem>>,
+        U: Borrow<Handle<tags::Theorem>>,
+        V: Borrow<Handle<tags::Theorem>>,
     {
         let left = self
             .resolve_theorem_handle(left)
@@ -2878,13 +2881,14 @@ impl RuntimeState {
         Ok(self.admit_theorem(Theorem::new(hypotheses, conclusion)))
     }
 
-    pub fn theorem_register_implication_elimination<T>(
+    pub fn theorem_register_implication_elimination<T, U>(
         &mut self,
         left: T,
-        right: T,
+        right: U,
     ) -> Result<Handle<tags::Theorem>, ErrorCode>
     where
         T: Borrow<Handle<tags::Theorem>>,
+        U: Borrow<Handle<tags::Theorem>>,
     {
         let left = self
             .resolve_theorem_handle(left)
[…]

### 2021-08-03 src/wasmi/host_interface.rs

@@ -462,8 +462,6 @@ pub const ABI_THEOREM_REGISTER_EXISTS_INTRODUCTION_NAME: &str =
 /// The name of the `Theorem.Register.ExistsElimination` ABI call.
 pub const ABI_THEOREM_REGISTER_EXISTS_ELIMINATION_NAME: &str =
     "__theorem_register_exists_elimination";
-/// The name of the `Theorem.Register.Weaken` ABI call.
-pub const ABI_THEOREM_REGISTER_WEAKEN_NAME: &str = "__theorem_register_weaken";
 
 /// The name of the `Theorem.Split.Hypotheses` ABI call.
 pub const ABI_THEOREM_SPLIT_HYPOTHESES_NAME: &str =
@@ -540,13 +538,11 @@ pub const ABI_THEOREM_REGISTER_FORALL_ELIMINATION_INDEX: usize = 83;
 pub const ABI_THEOREM_REGISTER_EXISTS_INTRODUCTION_INDEX: usize = 84;
 /// The index of the `Theorem.Register.ExistsElimination` ABI call.
 pub const ABI_THEOREM_REGISTER_EXISTS_ELIMINATION_INDEX: usize = 85;
-/// The index of the `Theorem.Register.Weaken` ABI call.
-pub const ABI_THEOREM_REGISTER_WEAKEN_INDEX: usize = 86;
 
 /// The index of the `Theorem.Split.Hypotheses` ABI call.
-pub const ABI_THEOREM_SPLIT_HYPOTHESES_INDEX: usize = 87;
+pub const ABI_THEOREM_SPLIT_HYPOTHESES_INDEX: usize = 86;
 /// The index of the `Theorem.Split.Conclusion` ABI call.
-pub const ABI_THEOREM_SPLIT_CONCLUSION_INDEX: usize = 88;
+pub const ABI_THEOREM_SPLIT_CONCLUSION_INDEX: usize = 87;
 
 ////////////////////////////////////////////////////////////////////////////////
 // Errors and traps.
@@ -1889,7 +1885,7 @@ impl WasmiRuntimeState {
     fn theorem_register_falsity_elimination<T, U>(
         &self,
         theorem_handle: T,
-        conclusion_handle: U,
+        term_handle: U,
     ) -> Result<Handle<tags::Theorem>, KernelErrorCode>
     where
         T: Borrow<Handle<tags::Theorem>>,
@@ -1897,10 +1893,7 @@ impl WasmiRuntimeState {
     {
         self.kernel
             .borrow_mut()
[…]

### 2021-08-03 kernel/src/_type.rs
supervisionary: split project up into separate sub-projects using workspaces

@@ -13,7 +13,7 @@
 //! [Dominic Mulligan]: https://dominic-mulligan.co.uk
 //! [Arm Research]: http://www.arm.com/research
 
-use crate::kernel::{
+use crate::{
     handle::{
         tags, Handle, PREALLOCATED_HANDLE_TYPE_ALPHA,
         PREALLOCATED_HANDLE_TYPE_FORMER_ARROW,
@@ -232,7 +232,7 @@ lazy_static! {
 
 #[cfg(test)]
 mod test {
-    use crate::kernel::{
+    use crate::{
         _type::Type,
         handle::{
             tags, Handle, PREALLOCATED_HANDLE_TYPE_FORMER_ARROW,

### 2021-08-03 kernel/src/error_code.rs

@@ -23,8 +23,11 @@ use std::{
     fmt::{Display, Error as DisplayError, Formatter},
 };
 
+#[cfg(feature = "wasmi-hosterror")]
+use wasmi::HostError;
+
 /// The upper limit (exclusive) on the encoding space of the `ErrorCode` type.
-pub const ERRORCODE_ENCODING_UPPER_BOUND: usize = 29;
+pub const ERRORCODE_ENCODING_UPPER_BOUND: usize = 28;
 
 /// Error codes, used for passing back information on why a kernel operation
 /// failed to prover-space.  These codes are intra-convertible between the `i32`
@@ -34,8 +37,6 @@ pub enum ErrorCode {
     /* ABI errors. */
     /// The operation completed successfully.
     Success,
-    /// The type-signature of an ABI function was not as expected.
-    SignatureFailure,
     /// The WASM guest program tried to call a host function that does not
     /// exist.
     NoSuchFunction,
@@ -71,7 +72,11 @@ pub enum ErrorCode {
     /// A term passed to a function was expected to be a constant but it was
     /// not.
     NotAConstant,
+    /// A term passed to a function was expected to be a universal quantifier
+    /// but it was not.
     NotAForall,
+    /// A term passed to a function was expected to be a disjunction but it was
+    /// not.
     NotADisjunction,
     /// A term passed to a function was expected to be a lambda-abstraction but
     /// it was not.
@@ -79,8 +84,14 @@ pub enum ErrorCode {
     /// A term passed to a function was expected to be an application but it was
     /// not.
     NotAnApplication,
+    /// A term passed to a function was expected to be an equality but it was
+    /// not.
     NotAnEquality,
+    /// A term passed to a function was expected to be an existential quantifier
+    /// but it was not.
     NotAnExists,
+    /// A term passed to a function was expected to be an implication but it was
+    /// not.
     NotAnImplication,
     /// A term passed to a function was expected to be a negation but it was
     /// not.
@@ -109,7 +120,6 @@ impl Display for ErrorCode {
     fn fmt(&self, f: &mut Formatter) -> Result<(), DisplayError> {
         match self {
[…]

### 2021-08-03 kernel/src/runtime_state.rs

@@ -15,9 +15,9 @@
 //! [Dominic Mulligan]: https://dominic-mulligan.co.uk
 //! [Arm Research]: http://www.arm.com/research
 
-use log::{error, info};
+use log::info;
 
-use crate::kernel::{
+use crate::{
     _type::{
         Type, TYPE_ALPHA, TYPE_BETA, TYPE_BINARY_CONNECTIVE,
         TYPE_POLYMORPHIC_BINARY_PREDICATE, TYPE_POLYMORPHIC_QUANTIFIER,
@@ -274,27 +274,18 @@ impl RuntimeState {
 
         let former = former.into();
 
-        let arity = self.type_former_resolve(former.clone()).ok_or({
-            error!("Type-former handle is not registered.");
-
-            ErrorCode::NoSuchTypeFormerRegistered
-        })?;
+        let arity = self
+            .type_former_resolve(former.clone())
+            .ok_or(ErrorCode::NoSuchTypeFormerRegistered)?;
 
         if !arguments
             .iter()
             .all(|a| self.type_is_registered(a.clone().into()))
         {
-            error!("Not all type argument handles are registered.");
-
             return Err(ErrorCode::NoSuchTypeRegistered);
         }
 
         if arguments.len() != *arity {
-            error!(
-                "Number of arguments provided does not match registered arity: {}.",
-                arity,
-            );
-
             return Err(ErrorCode::MismatchedArity);
         }
 
@@ -329,14 +320,10 @@ impl RuntimeState {
         let range = range.into();
 
         if !self.type_is_registered(&domain) {
-            error!("Domain type handle is not registered.");
-
             return Err(ErrorCode::NoSuchTypeRegistered);
         }
 
         if !self.type_is_registered(&range) {
-            error!("Range type handle is not registered.");
-
             return Err(ErrorCode::NoSuchTypeRegistered);
         }
 
@@ -363,14 +350,8 @@ impl RuntimeState {
         info!("Splitting handle {} into type variable.", handle.borrow());
 
         if let Some(tau) = self.resolve_type_handle(handle) {
-            tau.split_variable().ok_or({
-                error!("Handle is not a type variable.");
-
-                ErrorCode::NotATypeVariable
-            })
+            tau.split_variable().ok_or(ErrorCode::NotATypeVariable)
         } else {
-            error!("Unknown handle.");
-
             Err(ErrorCode::NoSuchTypeRegistered)
         }
     }
@@ -400,12 +381,9 @@ impl RuntimeState {
         );
 
[…]

### 2021-08-03 kernel/src/term.rs

@@ -13,7 +13,7 @@
 //! [Dominic Mulligan]: https://dominic-mulligan.co.uk
 //! [Arm Research]: http://www.arm.com/research
 
-use crate::kernel::{
+use crate::{
     handle::{
         tags, Handle, PREALLOCATED_HANDLE_CONSTANT_CONJUNCTION,
         PREALLOCATED_HANDLE_CONSTANT_DISJUNCTION,

### 2021-08-03 kernel/src/theorem.rs

@@ -13,7 +13,7 @@
 //! [Dominic Mulligan]: https://dominic-mulligan.co.uk
 //! [Arm Research]: http://www.arm.com/research
 
-use crate::kernel::handle::{tags, Handle};
+use crate::handle::{tags, Handle};
 
 ////////////////////////////////////////////////////////////////////////////////
 // Theorems, proper.
@@ -72,7 +72,7 @@ impl Theorem {
 
 #[cfg(test)]
 mod test {
-    use crate::kernel::{
+    use crate::{
         handle::{tags, Handle, PREALLOCATED_HANDLE_TERM_TRUE},
         theorem::Theorem,
     };

### 2021-08-03 src/lib.rs

@@ -1,272 +1,0 @@
-//! # Supervisionary: a privileged proof checker
-//!
-//! *Supervisionary* is a proof-checker for Classical Higher-order Logic (HOL)
-//! in a similar vein to [Isabelle/HOL], [HOL4], [HOL Light], and similar
-//! systems.  Supervisionary is structured in a broadly similar sense to the
-//! aforementioned systems, in the sense that it provides a trusted,
-//! distinguished *kernel* which is the only system component capable of
-//! authoritatively authenticating or rejecting all claimed theorems in the
-//! implemented logic.
-//!
-//! ## LCF: the *status quo*
-//!
-//! Existing implementations of HOL (and other logics) are organised along lines
-//! pioneered by [Robin Milner] in the late-1970s.  This is the so-called "LCF
-//! approach" named after the [LCF] proof-checking system that Milner was
-//! designing when he and his team invented it, wherein the system's
-//! proof-checking kernel is implemented as a module in a *meta-language* which
-//! exposes an abstract type of theorems, with functions for constructing basic
-//! theorems, corresponding to the axioms of the logic, and other functions
-//! implementing the inference rules of the logic mapping existing theorems to
-//! new theorems.  The soundness of the system, therefore depends crucially on
-//! the soundness of the type- and module-system of the meta-language.  As a
-//! result, most such systems are today written in strongly-typed functional
-//! programming languages, such as [OCaml], [Standard ML], or [Haskell], all of
-//! which descended from Milner's original meta-language, [ML], which was designed
-//! explicitly for this task.
-//!
-//! Whilst this approach to system design obviously works well, and is widely
-//! used, it does suffer from a number of disadvantages.
-//!
-//! First, the system's meta-language is privileged amongst all programming
-//! languages.  This is because the system kernel, the component through which
-//! all logical inference must be routed, is little other than a module written
-//! in that meta-language.  As a result, all components wishing to communicate
-//! with the system's kernel must also be written in that programming language,
-//! or use a shim-layer written in that language to facilitate the interaction.
-//! Whilst this need not necessarily be a problem, what *is* a problem is the
-//! fact that proof-checking systems tend to be written in relatively niche
-//! programming languages, with relatively small communities of users, like the
-//! aforementioned Standard ML, OCaml, and Haskell.  Standard ML in particular
-//! is a cause for concern, as it is to all intents and purposes a programming
-//! language on "life support", barely kept alive for use in the Isabelle and
-//! HOL4 systems.  Indeed, Standard ML has a single maintained compiler
[…]

### 2021-08-03 src/wasmi/mod.rs

@@ -1,2 +1,0 @@
-pub mod host_interface;
-pub mod panic;

### 2021-08-03 src/wasmi/panic.rs

@@ -1,17 +1,0 @@
-//! # Panic messages.
-//!
-//! Various user-facing messages that are printed when the host panics due to an
-//! unrecoverable runtime error.
-//!
-//! # Authors
-//!
-//! [Dominic Mulligan], Systems Research Group, [Arm Research] Cambridge.
-//!
-//! # Copyright
-//!
-//! Copyright (c) Arm Limited, 2021.  All rights reserved (r).  Please see the
-//! `LICENSE.markdown` file in the *Supervisionary* root directory for licensing
-//! information.
-//!
-//! [Dominic Mulligan]: https://dominic-mulligan.co.uk
-//! [Arm Research]: http://www.arm.com/research

### 2021-08-03 wasmi-bindings/src/abi_types.rs

@@ -1,0 +1,74 @@
+//! # ABI types
+//!
+//! Type aliases and other related functionality for describing values passed
+//! across the ABI boundary.
+//!
+//! # Authors
+//!
+//! [Dominic Mulligan], Systems Research Group, [Arm Research] Cambridge.
+//!
+//! # Copyright
+//!
+//! Copyright (c) Arm Limited, 2021.  All rights reserved (r).  Please see the
+//! `LICENSE.markdown` file in the *Supervisionary* root directory for licensing
+//! information.
+//!
+//! [Dominic Mulligan]: https://dominic-mulligan.co.uk
+//! [Arm Research]: http://www.arm.com/research
+
+use wasmi::ValueType;
+
+/// Type-synonyms for declaratively describing the intended purpose of WASM
+/// types passed across the ABI boundary.
+pub(crate) mod semantic_types {
+    /// A WASM address, used for reading-from and writing-to the guest WASM
+    /// program heap, assuming the `wasm32-abi`.
+    pub type Pointer = u32;
+    /// An arity of a type-former.
+    pub type Arity = u64;
+    /// A handle to a kernel object.
+    pub type Handle = u64;
+    /// A name of a variable (e.g. a lambda-abstracted variable, or
+    /// type-variable).
+    pub type Name = u64;
+    /// A size of a buffer, or other heap-allocated structure, used for
+    /// reading-from and writing-to the guest WASM program heap, assuming the
+    /// `wasm32-abi`.
+    pub type Size = u64;
+}
+
+/// A type capturing semantic types of the ABI, more descriptive than the base
+/// types of WASM.  Note that the constructors of this type are intended to shadow
+/// the type-synyonyms defined in the `semantic_types` module.
+pub(crate) enum AbiType {
+    /// A handle pointing-to a kernel object.
+    Handle,
+    /// A name (e.g. of a lambda-abstracted variable, or similar).
+    Name,
+    /// An arity for a type-former.
+    Arity,
+    /// A pointer into the host WASM program's heap.
+    Pointer,
+    /// A size (or length) of an object appearing in the WASM program's heap.
+    Size,
+    /// A Boolean value.
+    Boolean,
+    /// An error code returned from an ABI function.
+    ErrorCode,
+}
+
+impl AbiType {
+    /// Returns `true` iff the current `AbiType` is implemented by the WASM
+    /// value type, `tau`.
[…]

### 2021-08-03 wasmi-bindings/src/hostcall_number.rs

@@ -1,0 +1,477 @@
+//! # Hostcall numbers
+//!
+//! Defines the number (and name) of each hostcall.
+//!
+//! # Authors
+//!
+//! [Dominic Mulligan], Systems Research Group, [Arm Research] Cambridge.
+//!
+//! # Copyright
+//!
+//! Copyright (c) Arm Limited, 2021.  All rights reserved (r).  Please see the
+//! `LICENSE.markdown` file in the *Supervisionary* root directory for licensing
+//! information.
+//!
+//! [Dominic Mulligan]: https://dominic-mulligan.co.uk
+//! [Arm Research]: http://www.arm.com/research
+
+/// The name of the `TypeFormer.Resolve` ABI call.
+pub(crate) const ABI_TYPE_FORMER_RESOLVE_NAME: &str = "__type_former_resolve";
+/// The name of the `TypeFormer.IsRegistered` ABI call.
+pub(crate) const ABI_TYPE_FORMER_IS_REGISTERED_NAME: &str =
+    "__type_former_is_registered";
+/// The name of the `TypeFormer.Register` ABI call.
+pub(crate) const ABI_TYPE_FORMER_REGISTER_NAME: &str = "__type_former_register";
+
+/// The host-call number of the `TypeFormer.Resolve` ABI call.
+pub(crate) const ABI_TYPE_FORMER_RESOLVE_INDEX: usize = 0;
+/// The host-call number of the `TypeFormer.IsRegistered` ABI call.
+pub(crate) const ABI_TYPE_FORMER_IS_REGISTERED_INDEX: usize = 1;
+/// The host-call number of the `TypeFormer.Register` ABI call.
+pub(crate) const ABI_TYPE_FORMER_REGISTER_INDEX: usize = 2;
+
+/// The name of the `Type.IsRegistered` ABI call.
+pub(crate) const ABI_TYPE_IS_REGISTERED_NAME: &str = "__type_is_registered";
+/// The name of the `Type.Register.Variable` ABI call.
+pub(crate) const ABI_TYPE_REGISTER_VARIABLE_NAME: &str =
+    "__type_register_variable";
+/// The name of the `Type.Register.Combination` ABI call.
+pub(crate) const ABI_TYPE_REGISTER_COMBINATION_NAME: &str =
+    "__type_register_combination_name";
+/// The name of the `Type.Register.Function` ABI call.
[…]

### 2021-08-03 wasmi-bindings/src/lib.rs

@@ -1,0 +1,27 @@
+//! # WASMI host interface
+//!
+//! # Authors
+//!
+//! [Dominic Mulligan], Systems Research Group, [Arm Research] Cambridge.
+//!
+//! # Copyright
+//!
+//! Copyright (c) Arm Limited, 2021.  All rights reserved (r).  Please see the
+//! `LICENSE.markdown` file in the *Supervisionary* root directory for licensing
+//! information.
+//!
+//! [Dominic Mulligan]: https://dominic-mulligan.co.uk
+//! [Arm Research]: http://www.arm.com/research
+
+/// Type-aliases and other functionality related to passing values acrossed the
+/// ABI boundary.
+mod abi_types;
+/// The numbers (and names) of hostcall functions exposed to the running WASM
+/// program.
+mod hostcall_number;
+/// Binding the kernel's functionality to the WASMI host interface.
+pub mod runtime_state;
+/// Functionality related to runtime traps (an analogue of a traditional "kernel
+/// panic").
+mod runtime_trap;
+mod type_checking;

### 2021-08-03 wasmi-bindings/src/runtime_state.rs

@@ -1,4 +1,4 @@
-//! # WASMI host interface
+//! # WASMI runtime state
 //!
 //! This binds the kernel code, proper, to the guest WASM program-facing ABI
 //! interface, routing host-calls as appropriate.
@@ -16,585 +16,149 @@
 //! [Dominic Mulligan]: https://dominic-mulligan.co.uk
 //! [Arm Research]: http://www.arm.com/research
 
-use std::{
-    borrow::Borrow,
-    cell::RefCell,
-    fmt::{Debug, Display, Error as DisplayError, Formatter},
-    mem::size_of,
-};
+use std::{borrow::Borrow, cell::RefCell, fmt::Debug, mem::size_of};
 
 use byteorder::{ByteOrder, LittleEndian};
 use wasmi::{
     Error as WasmiError, Externals, FuncInstance, FuncRef, HostError,
     MemoryInstance, ModuleImportResolver, RuntimeArgs, RuntimeValue, Signature,
-    Trap, TrapKind, ValueType,
+    Trap,
 };
 
-use crate::kernel::{
+use kernel::{
     error_code::ErrorCode as KernelErrorCode,
     handle::{tags, Handle},
     name::Name,
     runtime_state::RuntimeState as KernelRuntimeState,
 };
 
-////////////////////////////////////////////////////////////////////////////////
-// Semantic types for the WASM ABI.
-////////////////////////////////////////////////////////////////////////////////
-
-/// Type-synonyms for declaratively describing the intended purpose of WASM
-/// types passed across the ABI boundary.
-mod semantic_types {
-    /// A WASM address, used for reading-from and writing-to the guest WASM
-    /// program heap, assuming the `wasm32-abi`.
-    pub type Pointer = u32;
-    /// An arity of a type-former.
-    pub type Arity = u64;
-    /// A handle to a kernel object.
-    pub type Handle = u64;
-    /// A name of a variable (e.g. a lambda-abstracted variable, or
-    /// type-variable).
-    pub type Name = u64;
-    /// A size of a buffer, or other heap-allocated structure, used for
-    /// reading-from and writing-to the guest WASM program heap, assuming the
-    /// `wasm32-abi`.
-    pub type Size = u64;
-}
-
[…]

### 2021-08-03 wasmi-bindings/src/runtime_trap.rs

@@ -1,0 +1,52 @@
+use std::fmt::{Display, Error as DisplayError, Formatter};
+
+use wasmi::{Error as WasmiError, HostError, Trap, TrapKind};
+
+use kernel::error_code::ErrorCode;
+
+/// Runtime traps are unrecoverable errors raised by the WASM program host.
+/// These are equivalent, essentially, to kernel panics in a typical operating
+/// system.
+#[derive(Clone, Debug, Eq, Hash, Ord, PartialEq, PartialOrd)]
+pub enum RuntimeTrap {
+    /// The WASM guest's memory was not registered with the runtime state.
+    MemoryNotRegistered,
+    /// An attempted read from the WASM guest's heap failed.
+    MemoryReadFailed,
+    /// An attempted write to the WASM guest's heap failed.
+    MemoryWriteFailed,
+    /// The WASM guest program tried to call a function that does not exist.
+    NoSuchFunction,
+    /// A type-signature check on a host-function failed.
+    SignatureFailure,
+}
+
+/// Pretty-printing for `RuntimeTrap` values.
+impl Display for RuntimeTrap {
+    fn fmt(&self, f: &mut Formatter) -> Result<(), DisplayError> {
+        match self {
+            RuntimeTrap::NoSuchFunction => write!(f, "NoSuchFunction"),
+            RuntimeTrap::SignatureFailure => write!(f, "SignatureFailure"),
+            RuntimeTrap::MemoryNotRegistered => {
+                write!(f, "MemoryNotRegistered")
+            }
+            RuntimeTrap::MemoryReadFailed => write!(f, "MemoryReadFailed"),
+            RuntimeTrap::MemoryWriteFailed => write!(f, "MemoryWriteFailed"),
+        }
+    }
+}
+
+impl HostError for RuntimeTrap {}
+
+/// Lifts a kernel error into an error that can be passed back to the WASM
+/// program.
+#[inline]
+pub fn host_error(code: ErrorCode) -> WasmiError {
+    WasmiError::Host(Box::new(code))
+}
+
+/// Creates a WASMI `Trap` type from a `RuntimeTrap`.
+#[inline]
+pub fn host_trap(trap: RuntimeTrap) -> Trap {
+    Trap::new(TrapKind::Host(Box::new(trap)))
+}

### 2021-08-03 wasmi-bindings/src/type_checking.rs

@@ -1,0 +1,1116 @@
+//! # Type-checking for host-calls
+//!
+//! # Authors
+//!
+//! [Dominic Mulligan], Systems Research Group, [Arm Research] Cambridge.
+//!
+//! # Copyright
+//!
+//! Copyright (c) Arm Limited, 2021.  All rights reserved (r).  Please see the
+//! `LICENSE.markdown` file in the *Supervisionary* root directory for licensing
+//! information.
+//!
+//! [Dominic Mulligan]: https://dominic-mulligan.co.uk
+//! [Arm Research]: http://www.arm.com/research
+
+use wasmi::Signature;
+
+use crate::abi_types::AbiType;
+
+/// Returns `true` iff the semantic function signature type described by
+/// `params` and `ret` is implemented by the WASM type described by
+/// `signature`.
+fn check_signature(
+    signature: &Signature,
+    params: &[AbiType],
+    ret: &Option<AbiType>,
+) -> bool {
+    let params = signature
+        .params()
+        .iter()
+        .zip(params)
+        .all(|(a, w)| w.implemented_by(a));
+
+    match (ret, signature.return_type()) {
+        (None, None) => params,
+        (Some(a), Some(w)) => a.implemented_by(&w) && params,
+        _otherwise => false,
+    }
+}
+
+/// Checks the signature of the `TypeFormer.Resolve` ABI function.
+#[inline]
+pub(crate) fn check_type_former_resolve_signature(
+    signature: &Signature,
+) -> bool {
+    check_signature(
+        signature,
+        &[AbiType::Handle, AbiType::Pointer],
+        &Some(AbiType::ErrorCode),
+    )
+}
+
+/// Checks the signature of the `TypeFormer.Register` ABI function.
+#[inline]
+pub(crate) fn check_type_former_register_signature(
+    signature: &Signature,
+) -> bool {
+    check_signature(signature, &[AbiType::Arity], &Some(AbiType::Handle))
+}
+
+/// Checks the signature of the `TypeFormer.IsRegistered` ABI function.
+#[inline]
+pub(crate) fn check_type_former_is_registered_signature(
+    signature: &Signature,
+) -> bool {
+    check_signature(signature, &[AbiType::Handle], &Some(AbiType::Boolean))
+}
+
+/// Checks the signature of the `Type.Register.Variable` ABI function.
+#[inline]
+pub(crate) fn check_type_register_variable_signature(
+    signature: &Signature,
+) -> bool {
[…]

### 2021-08-03 driver/src/main.rs
libsupervisionary+driver: introduced libsupervisionary and driver crates

@@ -1,0 +1,78 @@
+//! # Entry point for the driver application
+//!
+//! # Authors
+//!
+//! [Dominic Mulligan], Systems Research Group, [Arm Research] Cambridge.
+//!
+//! # Copyright
+//!
+//! Copyright (c) Arm Limited, 2021.  All rights reserved (r).  Please see the
+//! `LICENSE.markdown` file in the *Supervisionary* root directory for licensing
+//! information.
+//!
+//! [Dominic Mulligan]: https://dominic-mulligan.co.uk
+//! [Arm Research]: http://www.arm.com/research
+
+use anyhow::Result;
+use clap::{App, Arg};
+use env_logger;
+use log::info;
+use std::{path::PathBuf, process::exit};
+
+////////////////////////////////////////////////////////////////////////////////
+// Useful constants.
+////////////////////////////////////////////////////////////////////////////////
+
+const APPLICATION_NAME: &str = "Supervisionary driver.";
+const ABOUT_MESSAGE: &str = "Main driver application for Supervisionary tests.";
+const AUTHOR_LIST: &str = "The Supervisionary Development Team.";
+const VERSION_NUMBER: &str = "0.1.0";
+
+////////////////////////////////////////////////////////////////////////////////
+// Command-line parsing.
+////////////////////////////////////////////////////////////////////////////////
+
+/// Captures the command line arguments passed to the program.
+struct CommandLineArguments {
+    /// The path of the Wasm binary to load.
+    wasm_binary_path: PathBuf,
+}
+
+/// Parses the command line arguments of the program, exiting with an error code
+/// if this cannot be done successfully.  Otherwise, packs the command line
+/// arguments into a `CommandLineArguments` value, which is returned.
+fn parse_command_line_arguments() -> CommandLineArguments {
+    let mut app = App::new(APPLICATION_NAME)
+        .about(ABOUT_MESSAGE)
+        .version(VERSION_NUMBER)
+        .author(AUTHOR_LIST);
+
+    app.arg(
+        Arg::new("wasm-binary-path")
+            .required(true)
+            .short('b')
+            .long("binary")
+            .about("Path to the Wasm binary to load"),
+    );
+
+    let matches = app.get_matches();
+
+    if let Some(path) = matches.value_of("wasm-binary-path") {
+        CommandLineArguments {
+            wasm_binary_path: PathBuf::from(path),
+        }
+    } else {
+        eprintln!("No Wasm binary path provided as argument.");
+        exit(1)
+    }
+}
+
[…]

### 2021-08-03 libsupervisionary/src/lib.rs

@@ -1,0 +1,18 @@
+//! # libsupervisionary
+//!
+//! Rust language support for interacting with the Supervisionary kernel.  This
+//! library provides a binding to the kernel's host-call interface, with some
+//! minor abstractions built on top.
+//!
+//! # Authors
+//!
+//! [Dominic Mulligan], Systems Research Group, [Arm Research] Cambridge.
+//!
+//! # Copyright
+//!
+//! Copyright (c) Arm Limited, 2021.  All rights reserved (r).  Please see the
+//! `LICENSE.markdown` file in the *Supervisionary* root directory for licensing
+//! information.
+//!
+//! [Dominic Mulligan]: https://dominic-mulligan.co.uk
+//! [Arm Research]: http://www.arm.com/research

### 2021-08-04 driver/src/main.rs
libsupervisionary+tests: started testing kernel/libsupervisionary

@@ -13,20 +13,40 @@
 //! [Dominic Mulligan]: https://dominic-mulligan.co.uk
 //! [Arm Research]: http://www.arm.com/research
 
-use anyhow::Result;
 use clap::{App, Arg};
-use env_logger;
 use log::info;
-use std::{path::PathBuf, process::exit};
+use wasmi::{
+    ExternVal, ImportsBuilder, MemoryRef, Module, ModuleInstance, ModuleRef,
+    RuntimeValue,
+};
+
+use wasmi_bindings::runtime_state::WasmiRuntimeState;
+
+use std::{
+    fs::File,
+    io::Read,
+    path::{Path, PathBuf},
+    process::exit,
+};
 
 ////////////////////////////////////////////////////////////////////////////////
 // Useful constants.
 ////////////////////////////////////////////////////////////////////////////////
 
-const APPLICATION_NAME: &str = "Supervisionary driver.";
+/// An about message/header for the help menu of the driver application.
 const ABOUT_MESSAGE: &str = "Main driver application for Supervisionary tests.";
+/// Name of the driver application.
+const APPLICATION_NAME: &str = "Supervisionary driver.";
+/// Authors of the driver application.
 const AUTHOR_LIST: &str = "The Supervisionary Development Team.";
+/// The name of the Wasm module's heap.
+const LINEAR_MEMORY_NAME: &str = "memory";
+/// The version number of the driver application.
 const VERSION_NUMBER: &str = "0.1.0";
+/// The name of the Wasm entry point.
+const WASM_ENTRY_POINT: &str = "main";
+/// The name of the module resolved by the Wasmi imports resolver.
+const WASMI_MODULE_IMPORTS_RESOLVER_NAME: &str = "env";
 
 ////////////////////////////////////////////////////////////////////////////////
 // Command-line parsing.
@@ -42,22 +62,25 @@ struct CommandLineArguments {
 /// if this cannot be done successfully.  Otherwise, packs the command line
 /// arguments into a `CommandLineArguments` value, which is returned.
 fn parse_command_line_arguments() -> CommandLineArguments {
-    let mut app = App::new(APPLICATION_NAME)
+    info!("Parsing command line arguments.");
+
+    let matches = App::new(APPLICATION_NAME)
         .about(ABOUT_MESSAGE)
         .version(VERSION_NUMBER)
-        .author(AUTHOR_LIST);
-
-    app.arg(
-        Arg::new("wasm-binary-path")
-            .required(true)
[…]

### 2021-08-04 libsupervisionary/src/_type.rs

@@ -1,0 +1,46 @@
+//! # Bindings to Supervisionary's type ABI
+//!
+//! # Authors
+//!
+//! [Dominic Mulligan], Systems Research Group, [Arm Research] Cambridge.
+//!
+//! # Copyright
+//!
+//! Copyright (c) Arm Limited, 2021.  All rights reserved (r).  Please see the
+//! `LICENSE.markdown` file in the *Supervisionary* root directory for licensing
+//! information.
+//!
+//! [Dominic Mulligan]: https://dominic-mulligan.co.uk
+//! [Arm Research]: http://www.arm.com/research
+
+use crate::handle::{tags, Handle};
+use std::marker::PhantomData;
+
+////////////////////////////////////////////////////////////////////////////////
+// Pre-allocated type-related handles.
+////////////////////////////////////////////////////////////////////////////////
+
+/// A pre-allocated handle used to refer to the type-variable `A`.
+pub const PREALLOCATED_HANDLE_TYPE_ALPHA: Handle<tags::Type> =
+    Handle::new(2usize, PhantomData);
+/// A pre-allocated handle used to refer to the type-variable `B`.
+pub const PREALLOCATED_HANDLE_TYPE_BETA: Handle<tags::Type> =
+    Handle::new(3usize, PhantomData);
+/// A pre-allocated handle used to refer to the `Prop` type.
+pub const PREALLOCATED_HANDLE_TYPE_PROP: Handle<tags::Type> =
+    Handle::new(4usize, PhantomData);
+/// A pre-allocated handle used to refer to the type of unary predicates.
+pub const PREALLOCATED_HANDLE_TYPE_UNARY_PREDICATE: Handle<tags::Type> =
+    Handle::new(5usize, PhantomData);
+/// A pre-allocated handle used to refer to the type of binary predicates.
+pub const PREALLOCATED_HANDLE_TYPE_BINARY_PREDICATE: Handle<tags::Type> =
+    Handle::new(6usize, PhantomData);
+/// A pre-allocated handle used to refer to the type of unary connectives.
+pub const PREALLOCATED_HANDLE_TYPE_UNARY_CONNECTIVE: Handle<tags::Type> =
+    Handle::new(7usize, PhantomData);
+/// A pre-allocated handle used to refer to the type of binary connectives.
+pub const PREALLOCATED_HANDLE_TYPE_BINARY_CONNECTIVE: Handle<tags::Type> =
[…]

### 2021-08-04 libsupervisionary/src/constant.rs

@@ -1,0 +1,49 @@
+//! # Bindings to Supervisionary's constant ABI
+//!
+//! # Authors
+//!
+//! [Dominic Mulligan], Systems Research Group, [Arm Research] Cambridge.
+//!
+//! # Copyright
+//!
+//! Copyright (c) Arm Limited, 2021.  All rights reserved (r).  Please see the
+//! `LICENSE.markdown` file in the *Supervisionary* root directory for licensing
+//! information.
+//!
+//! [Dominic Mulligan]: https://dominic-mulligan.co.uk
+//! [Arm Research]: http://www.arm.com/research
+
+use crate::handle::{tags, Handle};
+use std::marker::PhantomData;
+
+////////////////////////////////////////////////////////////////////////////////
+// Pre-allocated constant-related handles.
+////////////////////////////////////////////////////////////////////////////////
+
+/// A pre-allocated handle used to refer to the truth constant.
+pub const PREALLOCATED_HANDLE_CONSTANT_TRUE: Handle<tags::Constant> =
+    Handle::new(10usize, PhantomData);
+/// A pre-allocated handle used to refer to the falsity constant.
+pub const PREALLOCATED_HANDLE_CONSTANT_FALSE: Handle<tags::Constant> =
+    Handle::new(11usize, PhantomData);
+/// A pre-allocated handle used to refer to the negation constant.
+pub const PREALLOCATED_HANDLE_CONSTANT_NEGATION: Handle<tags::Constant> =
+    Handle::new(12usize, PhantomData);
+/// A pre-allocated handle used to refer to the binary conjunction connective.
+pub const PREALLOCATED_HANDLE_CONSTANT_CONJUNCTION: Handle<tags::Constant> =
+    Handle::new(13usize, PhantomData);
+/// A pre-allocated handle used to refer to the binary disjunction connective.
+pub const PREALLOCATED_HANDLE_CONSTANT_DISJUNCTION: Handle<tags::Constant> =
+    Handle::new(14usize, PhantomData);
+/// A pre-allocated handle used to refer to the binary implication connective.
+pub const PREALLOCATED_HANDLE_CONSTANT_IMPLICATION: Handle<tags::Constant> =
+    Handle::new(15usize, PhantomData);
+/// A pre-allocated handle used to refer to the universal quantifier constant.
+pub const PREALLOCATED_HANDLE_CONSTANT_FORALL: Handle<tags::Constant> =
[…]

### 2021-08-04 libsupervisionary/src/error_code.rs

@@ -1,0 +1,180 @@
+//! # Error codes
+//!
+//! # Authors
+//!
+//! [Dominic Mulligan], Systems Research Group, [Arm Research] Cambridge.
+//!
+//! # Copyright
+//!
+//! Copyright (c) Arm Limited, 2021.  All rights reserved (r).  Please see the
+//! `LICENSE.markdown` file in the *Supervisionary* root directory for licensing
+//! information.
+//!
+//! [Dominic Mulligan]: https://dominic-mulligan.co.uk
+//! [Arm Research]: http://www.arm.com/research
+
+use std::{
+    convert::TryFrom,
+    fmt::{Display, Error as DisplayError, Formatter},
+};
+
+////////////////////////////////////////////////////////////////////////////////
+// Error codes.
+////////////////////////////////////////////////////////////////////////////////
+
+/// Error codes, used for passing back information on why a kernel operation
+/// failed to prover-space.  These codes are intra-convertible between the `i32`
+/// type.
+#[derive(Clone, Debug, Eq, Hash, Ord, PartialEq, PartialOrd)]
+pub enum ErrorCode {
+    /* ABI errors. */
+    /// The WASM guest program tried to call a host function that does not
+    /// exist.
+    NoSuchFunction,
+    /* Dangling objects. */
+    /// A handle was supplied that did not reference a registered constant.
+    NoSuchConstantRegistered,
+    /// A handle was supplied that did not reference a registered term.
+    NoSuchTermRegistered,
+    /// A handle was supplied that did not reference a registered theorem.
+    NoSuchTheoremRegistered,
+    /// A handle was supplied that did not reference a registered type-former.
+    NoSuchTypeFormerRegistered,
+    /* Type-former related errors. */
+    /// A type-former was applied to the wrong number of arguments.
+    MismatchedArity,
+    /* -- Type related errors. */
+    /// A term with functional type was applied to an argument that had a
+    /// different type to the domain type of the function.
+    DomainTypeMismatch,
+    /// A handle was supplied that did not reference a registered type.
+    NoSuchTypeRegistered,
+    /// A type was expected to be a functional type, but it was not.
+    NotAFunctionType,
+    /// A type was expected to be a type-combination, but it was not.
+    NotATypeCombination,
+    /// A type was expected to be a type-variable, but it was not.
[…]

### 2021-08-04 libsupervisionary/src/handle.rs

@@ -1,0 +1,135 @@
+//! # Kernel handles
+//!
+//! # Authors
+//!
+//! [Dominic Mulligan], Systems Research Group, [Arm Research] Cambridge.
+//!
+//! # Copyright
+//!
+//! Copyright (c) Arm Limited, 2021.  All rights reserved (r).  Please see the
+//! `LICENSE.markdown` file in the *Supervisionary* root directory for licensing
+//! information.
+//!
+//! [Dominic Mulligan]: https://dominic-mulligan.co.uk
+//! [Arm Research]: http://www.arm.com/research
+
+use std::{
+    fmt,
+    fmt::{Display, Formatter},
+    marker::PhantomData,
+    ops::Deref,
+};
+
+////////////////////////////////////////////////////////////////////////////////
+// Handle tags.
+////////////////////////////////////////////////////////////////////////////////
+
+pub mod tags {
+    #[derive(Clone, Debug, Eq, Hash, Ord, PartialEq, PartialOrd)]
+    pub struct TypeFormer;
+
+    #[derive(Clone, Debug, Eq, Hash, Ord, PartialEq, PartialOrd)]
+    pub struct Type;
+
+    #[derive(Clone, Debug, Eq, Hash, Ord, PartialEq, PartialOrd)]
+    pub struct Constant;
+
+    #[derive(Clone, Debug, Eq, Hash, Ord, PartialEq, PartialOrd)]
+    pub struct Term;
+
+    #[derive(Clone, Debug, Eq, Hash, Ord, PartialEq, PartialOrd)]
+    pub struct Theorem;
+
+    pub trait IsTag {}
+
+    impl IsTag for TypeFormer {}
+
+    impl IsTag for Type {}
+
+    impl IsTag for Constant {}
+
+    impl IsTag for Term {}
+
+    impl IsTag for Theorem {}
+}
+
+////////////////////////////////////////////////////////////////////////////////
+// Tagged handles.
+////////////////////////////////////////////////////////////////////////////////
+
+#[derive(Clone, Debug, Eq, Hash, Ord, PartialEq, PartialOrd)]
+pub struct Handle<T>
+where
+    T: tags::IsTag,
+{
+    /// We use the Rust `usize` type as our handle type.  Note that on modern 64-bit
+    /// systems this is implemented as a 64-bit unsigned integer.
+    handle: usize,
+    /// The phantom data binding the tag type, `T`.
+    marker: PhantomData<T>,
+}
+
+impl<T> Handle<T>
+where
+    T: tags::IsTag,
[…]

### 2021-08-04 libsupervisionary/src/lib.rs

@@ -16,3 +16,31 @@
 //!
 //! [Dominic Mulligan]: https://dominic-mulligan.co.uk
 //! [Arm Research]: http://www.arm.com/research
+
+#![feature(const_fn_trait_bound)]
+
+/// Bindings to the Supervisionary kernel's type ABI.
+pub mod _type;
+/// Bindings to the Supervisionary kernel's constant ABI.
+pub mod constant;
+/// Kernel error codes.  Note that the material in this module must exactly
+/// match the definition of error code used in the kernel.
+pub mod error_code;
+/// The kernel handle type, and some primitive handles associated with the
+/// Supervisionary ABI.
+pub mod handle;
+/// Bindings to the Supervisionary kernel's term ABI.
+pub mod term;
+/// Bindings to the Supervisionary kernel's theorem ABI.
+pub mod theorem;
+/// Bindings to the Supervisionary kernel's type-former ABI.
+pub mod type_former;
+
+/// The kernel type of names.
+pub type Name = u64;
+/// The kernel type of arities.
+pub type Arity = usize;
+/// The "raw" representation of handles expected by the kernel.
+pub(crate) type RawHandle = usize;
+/// The "raw" representation of kernel error modes.
+pub(crate) type RawKernelFailureMode = i32;

### 2021-08-04 libsupervisionary/src/term.rs

@@ -1,0 +1,58 @@
+//! # Bindings to Supervisionary's term ABI
+//!
+//! # Authors
+//!
+//! [Dominic Mulligan], Systems Research Group, [Arm Research] Cambridge.
+//!
+//! # Copyright
+//!
+//! Copyright (c) Arm Limited, 2021.  All rights reserved (r).  Please see the
+//! `LICENSE.markdown` file in the *Supervisionary* root directory for licensing
+//! information.
+//!
+//! [Dominic Mulligan]: https://dominic-mulligan.co.uk
+//! [Arm Research]: http://www.arm.com/research
+
+use crate::handle::{tags, Handle};
+use std::marker::PhantomData;
+
+////////////////////////////////////////////////////////////////////////////////
+// Pre-allocated term-related handles.
+////////////////////////////////////////////////////////////////////////////////
+
+/// A pre-allocated handle used to refer to the truth term, the truth constant
+/// lifted into a term.
+pub const PREALLOCATED_HANDLE_TERM_TRUE: Handle<tags::Term> =
+    Handle::new(19usize, PhantomData);
+/// A pre-allocated handle used to refer to the falsity term, the falsity
+/// constant lifted into a term.
+pub const PREALLOCATED_HANDLE_TERM_FALSE: Handle<tags::Term> =
+    Handle::new(20usize, PhantomData);
+/// A pre-allocated handle used to refer to the negation term, the negation
+/// constant lifted into a term.
+pub const PREALLOCATED_HANDLE_TERM_NEGATION: Handle<tags::Term> =
+    Handle::new(21usize, PhantomData);
+/// A pre-allocated handle used to refer to the conjunction term, the
+/// conjunction constant lifted into a term.
+pub const PREALLOCATED_HANDLE_TERM_CONJUNCTION: Handle<tags::Term> =
+    Handle::new(22usize, PhantomData);
+/// A pre-allocated handle used to refer to the disjunction term, the
+/// disjunction constant lifted into a term.
+pub const PREALLOCATED_HANDLE_TERM_DISJUNCTION: Handle<tags::Term> =
+    Handle::new(23usize, PhantomData);
+/// A pre-allocated handle used to refer to the implication term, the
+/// implication constant lifted into a term.
+pub const PREALLOCATED_HANDLE_TERM_IMPLICATION: Handle<tags::Term> =
+    Handle::new(24usize, PhantomData);
[…]

### 2021-08-04 libsupervisionary/src/theorem.rs

@@ -1,0 +1,14 @@
+//! # Bindings to Supervisionary's theorem ABI
+//!
+//! # Authors
+//!
+//! [Dominic Mulligan], Systems Research Group, [Arm Research] Cambridge.
+//!
+//! # Copyright
+//!
+//! Copyright (c) Arm Limited, 2021.  All rights reserved (r).  Please see the
+//! `LICENSE.markdown` file in the *Supervisionary* root directory for licensing
+//! information.
+//!
+//! [Dominic Mulligan]: https://dominic-mulligan.co.uk
+//! [Arm Research]: http://www.arm.com/research

### 2021-08-04 libsupervisionary/src/type_former.rs

@@ -1,0 +1,84 @@
+//! # Bindings to Supervisionary's type-former ABI
+//!
+//! # Authors
+//!
+//! [Dominic Mulligan], Systems Research Group, [Arm Research] Cambridge.
+//!
+//! # Copyright
+//!
+//! Copyright (c) Arm Limited, 2021.  All rights reserved (r).  Please see the
+//! `LICENSE.markdown` file in the *Supervisionary* root directory for licensing
+//! information.
+//!
+//! [Dominic Mulligan]: https://dominic-mulligan.co.uk
+//! [Arm Research]: http://www.arm.com/research
+
+use crate::{
+    error_code::ErrorCode,
+    handle::{tags, Handle},
+    Arity, RawHandle, RawKernelFailureMode,
+};
+use std::{convert::TryFrom, marker::PhantomData};
+
+////////////////////////////////////////////////////////////////////////////////
+// Pre-allocated type-former handles.
+////////////////////////////////////////////////////////////////////////////////
+
+/// A pre-allocated handle used to refer to the `Prop` type-former.
+pub const PREALLOCATED_HANDLE_TYPE_FORMER_PROP: Handle<tags::TypeFormer> =
+    Handle::new(0usize, PhantomData);
+/// A pre-allocated handle used to refer to the function-space type-former.
+pub const PREALLOCATED_HANDLE_TYPE_FORMER_ARROW: Handle<tags::TypeFormer> =
+    Handle::new(1usize, PhantomData);
+
+////////////////////////////////////////////////////////////////////////////////
+// ABI bindings.
+////////////////////////////////////////////////////////////////////////////////
+
+extern "C" {
+    /// Raw ABI binding to the `__type_former_register` function.
+    fn __type_former_register(handle: RawHandle) -> usize;
+    /// Raw ABI binding to the `__type_former_is_registered` function.
+    fn __type_former_is_registered(handle: RawHandle) -> bool;
+    /// Raw ABI binding to the `__type_former_resolve` function.
+    fn __type_former_resolve(
+        handle: RawHandle,
+        out: *mut usize,
+    ) -> RawKernelFailureMode;
+}
+
+/// Registers a new type-former with a given `arity`.  Returns the handle to the
+/// new type-former.  Note that this function is generative, in the sense that
+/// registering two type-formers with the same arity results in two different
+/// type-formers.
[…]

### 2021-08-04 tests/type_former/src/main.rs

@@ -1,0 +1,40 @@
+//! # Tests for the Supervisionary type-former ABI
+//!
+//! # Authors
+//!
+//! [Dominic Mulligan], Systems Research Group, [Arm Research] Cambridge.
+//!
+//! # Copyright
+//!
+//! Copyright (c) Arm Limited, 2021.  All rights reserved (r).  Please see the
+//! `LICENSE.markdown` file in the *Supervisionary* root directory for licensing
+//! information.
+//!
+//! [Dominic Mulligan]: https://dominic-mulligan.co.uk
+//! [Arm Research]: http://www.arm.com/research
+
+use libsupervisionary::type_former::{
+    type_former_is_registered, type_former_resolve,
+    PREALLOCATED_HANDLE_TYPE_FORMER_ARROW,
+    PREALLOCATED_HANDLE_TYPE_FORMER_PROP,
+};
+
+fn main() {
+    assert!(type_former_is_registered(
+        PREALLOCATED_HANDLE_TYPE_FORMER_PROP
+    ));
+
+    assert!(type_former_is_registered(
+        PREALLOCATED_HANDLE_TYPE_FORMER_ARROW
+    ));
+
+    assert_eq!(
+        type_former_resolve(PREALLOCATED_HANDLE_TYPE_FORMER_PROP),
+        Ok(0usize)
+    );
+
+    assert_eq!(
+        type_former_resolve(PREALLOCATED_HANDLE_TYPE_FORMER_ARROW),
+        Ok(2usize)
+    );
+}

### 2021-08-04 wasmi-bindings/src/runtime_state.rs

@@ -1,4 +1,4 @@
-//! # WASMI runtime state
+//! # Wasmi runtime state
 //!
 //! This binds the kernel code, proper, to the guest WASM program-facing ABI
 //! interface, routing host-calls as appropriate.
@@ -20,9 +20,8 @@ use std::{borrow::Borrow, cell::RefCell, fmt::Debug, mem::size_of};
 
 use byteorder::{ByteOrder, LittleEndian};
 use wasmi::{
-    Error as WasmiError, Externals, FuncInstance, FuncRef, HostError,
-    MemoryInstance, ModuleImportResolver, RuntimeArgs, RuntimeValue, Signature,
-    Trap,
+    Error as WasmiError, Externals, FuncInstance, FuncRef, MemoryRef,
+    ModuleImportResolver, RuntimeArgs, RuntimeValue, Signature, Trap,
 };
 
 use kernel::{
@@ -161,10 +160,10 @@ use crate::{
 ////////////////////////////////////////////////////////////////////////////////
 
 ////////////////////////////////////////////////////////////////////////////////
-// The WASMI runtime state.
+// The Wasmi runtime state.
 ////////////////////////////////////////////////////////////////////////////////
 
-/// The WASMI runtime state, which is a thin wrapper around the kerne's own
+/// The Wasmi runtime state, which is a thin wrapper around the kernel's own
 /// runtime state, adding a reference to the guest WASM program's memory module,
 /// to enable host functions to read-from and write-to the memory module
 /// directly.
@@ -173,7 +172,7 @@ pub struct WasmiRuntimeState {
     /// The kernel's runtime state.
     kernel: RefCell<KernelRuntimeState>,
     /// The memory instance of the executing WASM guest program.
-    memory: Option<RefCell<MemoryInstance>>,
+    memory: Option<RefCell<MemoryRef>>,
 }
 
 impl Default for WasmiRuntimeState {
@@ -188,7 +187,7 @@ impl Default for WasmiRuntimeState {
 
 impl WasmiRuntimeState {
     /// Constructs a new instance of a `WasmiRuntimeState` with the kernel state
-    /// intitialized to its correct initial state, and the reference to the WASM
+    /// initialised to its correct initial state, and the reference to the Wasm
     /// guest's memory set to `None`.
     #[inline]
     pub fn new() -> Self {
@@ -207,7 +206,7 @@ impl WasmiRuntimeState {
 
[…]

### 2021-08-04 wasmi-bindings/src/type_checking.rs

@@ -809,6 +809,44 @@ pub(crate) fn check_theorem_register_transitivity_signature(
     )
 }
 
+/// Checks the signature of the `Theorem.Register.Substitution` ABI function.
+#[inline]
+pub(crate) fn check_theorem_register_substitution_signature(
+    signature: &Signature,
+) -> bool {
+    check_signature(
+        signature,
+        &[
+            AbiType::Handle,
+            AbiType::Pointer,
+            AbiType::Size,
+            AbiType::Pointer,
+            AbiType::Size,
+            AbiType::Pointer,
+        ],
+        &Some(AbiType::ErrorCode),
+    )
+}
+
+/// Checks the signature of the `Theorem.Register.TypeSubstitution` ABI function.
+#[inline]
+pub(crate) fn check_theorem_register_type_substitution_signature(
+    signature: &Signature,
+) -> bool {
+    check_signature(
+        signature,
+        &[
+            AbiType::Handle,
+            AbiType::Pointer,
+            AbiType::Size,
+            AbiType::Pointer,
+            AbiType::Size,
+            AbiType::Pointer,
+        ],
+        &Some(AbiType::ErrorCode),
+    )
+}
+
 /// Checks the signature of the `Theorem.Register.Application` ABI function.
 #[inline]
 pub(crate) fn check_theorem_register_application_signature(

### 2021-08-04 driver/src/main.rs
test: first tests successfully run on kernel

@@ -170,7 +170,7 @@ fn main() {
 
     info!("Wasmi environment resolver and module instance created.");
 
-    if !not_started_module.has_start() {
+    if not_started_module.has_start() {
         eprintln!("Wasm module contains 'start' function.");
         exit(1);
     };

### 2021-08-04 libsupervisionary/src/lib.rs

@@ -39,8 +39,8 @@ pub mod type_former;
 /// The kernel type of names.
 pub type Name = u64;
 /// The kernel type of arities.
-pub type Arity = usize;
+pub type Arity = u64;
 /// The "raw" representation of handles expected by the kernel.
-pub(crate) type RawHandle = usize;
+pub(crate) type RawHandle = u64;
 /// The "raw" representation of kernel error modes.
 pub(crate) type RawKernelFailureMode = i32;

### 2021-08-04 libsupervisionary/src/type_former.rs

@@ -43,7 +43,7 @@ extern "C" {
     /// Raw ABI binding to the `__type_former_resolve` function.
     fn __type_former_resolve(
         handle: RawHandle,
-        out: *mut usize,
+        out: *mut u64,
     ) -> RawKernelFailureMode;
 }
 
@@ -53,7 +53,7 @@ extern "C" {
 /// type-formers.
 pub fn type_former_register<T>(arity: T) -> Handle<tags::TypeFormer>
 where
-    T: Into<usize>,
+    T: Into<Arity>,
 {
     let handle = unsafe { __type_former_register(arity.into()) };
 
@@ -63,7 +63,7 @@ where
 /// Returns `true` iff `handle` points-to a registered type-former in the
 /// kernel's heap.
 pub fn type_former_is_registered(handle: Handle<tags::TypeFormer>) -> bool {
-    unsafe { __type_former_is_registered(*handle) }
+    unsafe { __type_former_is_registered(*handle as u64) }
 }
 
 /// Returns the arity of the type-former pointed-to by `handle` in the kernel's
@@ -71,10 +71,11 @@ pub fn type_former_is_registered(handle: Handle<tags::TypeFormer>) -> bool {
 pub fn type_former_resolve(
     handle: Handle<tags::TypeFormer>,
 ) -> Result<Arity, ErrorCode> {
-    let mut arity: Arity = 0usize;
+    let mut arity: Arity = 0u64;
 
-    let result =
-        unsafe { __type_former_resolve(*handle, &mut arity as *mut usize) };
+    let result = unsafe {
+        __type_former_resolve(*handle as u64, &mut arity as *mut u64)
+    };
 
     if result == 0 {
         Ok(arity)

### 2021-08-04 tests/type_former/src/main.rs

@@ -30,11 +30,11 @@ fn main() {
 
     assert_eq!(
         type_former_resolve(PREALLOCATED_HANDLE_TYPE_FORMER_PROP),
-        Ok(0usize)
+        Ok(0u64)
     );
 
     assert_eq!(
         type_former_resolve(PREALLOCATED_HANDLE_TYPE_FORMER_ARROW),
-        Ok(2usize)
+        Ok(2u64)
     );
 }

### 2021-08-04 wasmi-bindings/src/runtime_state.rs

@@ -19,6 +19,7 @@
 use std::{borrow::Borrow, cell::RefCell, fmt::Debug, mem::size_of};
 
 use byteorder::{ByteOrder, LittleEndian};
+use log::{error, info};
 use wasmi::{
     Error as WasmiError, Externals, FuncInstance, FuncRef, MemoryRef,
     ModuleImportResolver, RuntimeArgs, RuntimeValue, Signature, Trap,
@@ -229,6 +230,10 @@ impl WasmiRuntimeState {
     where
         T: Into<semantic_types::Pointer>,
     {
+        let address = address.into();
+
+        info!("Writing {:?} bytes at address {}.", bytes, address);
+
         let memory = match &self.memory {
             None => return Err(RuntimeTrap::MemoryNotRegistered),
             Some(memory) => memory,
@@ -257,8 +262,13 @@ impl WasmiRuntimeState {
         T: Into<semantic_types::Pointer>,
         U: Into<u64>,
     {
-        let mut buffer = Vec::new();
-        LittleEndian::write_u64(&mut buffer, value.into());
+        let address = address.into();
+        let value = value.into();
+
+        info!("Writing u64 value {} at address {}.", value, address);
+
+        let mut buffer = vec![0u8; 8];
+        LittleEndian::write_u64(&mut buffer, value);
 
         self.write_bytes(address, &buffer)
     }
@@ -284,6 +294,12 @@ impl WasmiRuntimeState {
     {
         let mut address = address.into();
 
+        info!(
+            "Writing {} u64 values starting at address {}.",
+            values.len(),
+            address
+        );
+
         for v in values.iter().cloned() {
             self.write_u64(address, v)?;
             address += 8;
@@ -307,8 +323,13 @@ impl WasmiRuntimeState {
         T: Into<semantic_types::Pointer>,
         U: Into<bool>,
     {
-        let mut buffer = Vec::new();
-        LittleEndian::write_u32(&mut buffer, value.into() as u32);
+        let mut buffer = vec![0u8; 4];
+        let address = address.into();
+        let value = value.into();
+
+        info!("Writing bool value {} at address {}.", value, address);
+
+        LittleEndian::write_u32(&mut buffer, value as u32);
 
         self.write_bytes(address, &buffer)
     }
@@ -332,9 +353,14 @@ impl WasmiRuntimeState {
     where
         T: Into<semantic_types::Pointer>,
         U: Into<Handle<V>>,
-        V: tags::IsTag,
[…]

### 2021-08-04 libsupervisionary/src/handle.rs
test: expanded type-former tests, flushed out small bugs

@@ -85,6 +85,16 @@ where
 // Trait implementations.
 ////////////////////////////////////////////////////////////////////////////////
 
+impl<T> AsRef<Handle<T>> for Handle<T>
+where
+    T: tags::IsTag,
+{
+    #[inline]
+    fn as_ref(&self) -> &Handle<T> {
+        self
+    }
+}
+
 impl<T> Deref for Handle<T>
 where
     T: tags::IsTag,

### 2021-08-04 libsupervisionary/src/type_former.rs

@@ -37,7 +37,7 @@ pub const PREALLOCATED_HANDLE_TYPE_FORMER_ARROW: Handle<tags::TypeFormer> =
 
 extern "C" {
     /// Raw ABI binding to the `__type_former_register` function.
-    fn __type_former_register(handle: RawHandle) -> usize;
+    fn __type_former_register(handle: RawHandle) -> u64;
     /// Raw ABI binding to the `__type_former_is_registered` function.
     fn __type_former_is_registered(handle: RawHandle) -> bool;
     /// Raw ABI binding to the `__type_former_resolve` function.
@@ -57,24 +57,31 @@ where
 {
     let handle = unsafe { __type_former_register(arity.into()) };
 
-    Handle::new(handle, PhantomData)
+    Handle::new(handle as usize, PhantomData)
 }
 
 /// Returns `true` iff `handle` points-to a registered type-former in the
 /// kernel's heap.
-pub fn type_former_is_registered(handle: Handle<tags::TypeFormer>) -> bool {
-    unsafe { __type_former_is_registered(*handle as u64) }
+pub fn type_former_is_registered<H>(handle: H) -> bool
+where
+    H: AsRef<Handle<tags::TypeFormer>>,
+{
+    unsafe { __type_former_is_registered(*handle.as_ref().clone() as u64) }
 }
 
 /// Returns the arity of the type-former pointed-to by `handle` in the kernel's
 /// heap, if any.
-pub fn type_former_resolve(
-    handle: Handle<tags::TypeFormer>,
-) -> Result<Arity, ErrorCode> {
+pub fn type_former_resolve<H>(handle: H) -> Result<Arity, ErrorCode>
+where
+    H: AsRef<Handle<tags::TypeFormer>>,
+{
     let mut arity: Arity = 0u64;
 
     let result = unsafe {
-        __type_former_resolve(*handle as u64, &mut arity as *mut u64)
+        __type_former_resolve(
+            *handle.as_ref().clone() as u64,
+            &mut arity as *mut u64,
+        )
     };
 
     if result == 0 {

### 2021-08-04 tests/type_former/src/main.rs

@@ -14,7 +14,7 @@
 //! [Arm Research]: http://www.arm.com/research
 
 use libsupervisionary::type_former::{
-    type_former_is_registered, type_former_resolve,
+    type_former_is_registered, type_former_register, type_former_resolve,
     PREALLOCATED_HANDLE_TYPE_FORMER_ARROW,
     PREALLOCATED_HANDLE_TYPE_FORMER_PROP,
 };
@@ -37,4 +37,9 @@ fn main() {
         type_former_resolve(PREALLOCATED_HANDLE_TYPE_FORMER_ARROW),
         Ok(2u64)
     );
+
+    let handle = type_former_register(5u64);
+
+    assert_eq!(type_former_resolve(&handle), Ok(5u64));
+    assert!(type_former_is_registered(&handle));
 }

### 2021-08-04 wasmi-bindings/src/runtime_state.rs

@@ -451,6 +451,10 @@ impl WasmiRuntimeState {
     where
         T: Into<semantic_types::Pointer>,
     {
+        let address = address.into();
+
+        info!("Reading u64 value at address {}.", address);
+
         let buffer = self.read_bytes(address, size_of::<u64>())?;
         Ok(LittleEndian::read_u64(&buffer))
     }
@@ -476,8 +480,11 @@ impl WasmiRuntimeState {
     {
         let mut accumulator = Vec::new();
         let mut address = address.into();
+        let count = count.into();
 
-        for _c in 0..count.into() {
+        info!("Reading {} u64 values at address {}.", count, address);
+
+        for _c in 0..count {
             let handle = self.read_u64(address)?;
             accumulator.push(handle);
             address += 8;
@@ -502,6 +509,10 @@ impl WasmiRuntimeState {
         T: tags::IsTag,
         U: Into<u32>,
     {
+        let address = address.into();
+
+        info!("Reading handle at address {}.", address);
+
         Ok(Handle::from(self.read_u64(address)? as usize))
     }
 
@@ -527,8 +538,11 @@ impl WasmiRuntimeState {
     {
         let mut accumulator = Vec::new();
         let mut address = address.into();
+        let count = count.into();
 
-        for _c in 0..count.into() {
+        info!("Reading {} handles at address {}.", count, address);
+
+        for _c in 0..count {
             let handle = self.read_handle(address)?;
             accumulator.push(handle);
             address += 8;

### 2021-08-04 libsupervisionary/src/_type.rs
libsupervisionary: part-way through binding the Type.* ABI

@@ -13,7 +13,11 @@
 //! [Dominic Mulligan]: https://dominic-mulligan.co.uk
 //! [Arm Research]: http://www.arm.com/research
 
-use crate::handle::{tags, Handle};
+use crate::error_code::ErrorCode;
+use crate::{
+    handle::{tags, Handle},
+    Name, RawHandle,
+};
 use std::marker::PhantomData;
 
 ////////////////////////////////////////////////////////////////////////////////
@@ -44,3 +48,98 @@ pub const PREALLOCATED_HANDLE_TYPE_BINARY_CONNECTIVE: Handle<tags::Type> =
 /// A pre-allocated handle used to refer to the type of polymorphic quantifiers.
 pub const PREALLOCATED_HANDLE_TYPE_QUANTIFIER: Handle<tags::Type> =
     Handle::new(9usize, PhantomData);
+
+////////////////////////////////////////////////////////////////////////////////
+// ABI bindings.
+////////////////////////////////////////////////////////////////////////////////
+
+extern "C" {
+    fn __type_is_registered(handle: RawHandle) -> bool;
+    fn __type_register_variable(name: Name) -> u64;
+    fn __type_register_combination(
+        type_former_handle: RawHandle,
+        argument_base: *const RawHandle,
+        argument_length: u64,
+        result: *mut RawHandle,
+    ) -> i32;
+    fn __type_register_function(
+        domain_handle: RawHandle,
+        range_handle: RawHandle,
+        result: *mut RawHandle,
+    ) -> i32;
+    fn __type_split_variable(handle: RawHandle, result: *mut Name) -> i32;
+    fn __type_split_combination(
+        handle: RawHandle,
+        argument_base: *mut RawHandle,
+        argument_length: *mut u64,
+    ) -> i32;
+    fn __type_split_function(
+        handle: RawHandle,
+        domain_handle: *mut RawHandle,
+        range_handle: *mut RawHandle,
+    ) -> i32;
+    fn __type_test_variable(handle: RawHandle) -> bool;
+    fn __type_test_combination(handle: RawHandle) -> bool;
+    fn __type_test_function(handle: RawHandle) -> bool;
+    fn __type_variables(
+        handle: RawHandle,
+        result_base: *mut Name,
+        result_length: u64,
+    ) -> i32;
+    fn __type_substitute(
+        handle: RawHandle,
+        domain_base: *const Name,
+        domain_length: u64,
[…]

### 2021-08-04 libsupervisionary/src/type_former.rs

@@ -51,6 +51,7 @@ extern "C" {
 /// new type-former.  Note that this function is generative, in the sense that
 /// registering two type-formers with the same arity results in two different
 /// type-formers.
+#[inline]
 pub fn type_former_register<T>(arity: T) -> Handle<tags::TypeFormer>
 where
     T: Into<Arity>,
@@ -62,6 +63,7 @@ where
 
 /// Returns `true` iff `handle` points-to a registered type-former in the
 /// kernel's heap.
+#[inline]
 pub fn type_former_is_registered<H>(handle: H) -> bool
 where
     H: AsRef<Handle<tags::TypeFormer>>,

### 2021-08-04 libsupervisionary/src/_type.rs
libsupervisionary: bound Type.* ABI

@@ -13,12 +13,13 @@
 //! [Dominic Mulligan]: https://dominic-mulligan.co.uk
 //! [Arm Research]: http://www.arm.com/research
 
-use crate::error_code::ErrorCode;
 use crate::{
+    error_code::ErrorCode,
     handle::{tags, Handle},
     Name, RawHandle,
 };
-use std::marker::PhantomData;
+
+use std::{convert::TryFrom, marker::PhantomData, ptr::null_mut};
 
 ////////////////////////////////////////////////////////////////////////////////
 // Pre-allocated type-related handles.
@@ -70,6 +71,7 @@ extern "C" {
     fn __type_split_variable(handle: RawHandle, result: *mut Name) -> i32;
     fn __type_split_combination(
         handle: RawHandle,
+        type_former: *mut RawHandle,
         argument_base: *mut RawHandle,
         argument_length: *mut u64,
     ) -> i32;
@@ -78,13 +80,13 @@ extern "C" {
         domain_handle: *mut RawHandle,
         range_handle: *mut RawHandle,
     ) -> i32;
-    fn __type_test_variable(handle: RawHandle) -> bool;
-    fn __type_test_combination(handle: RawHandle) -> bool;
-    fn __type_test_function(handle: RawHandle) -> bool;
+    fn __type_test_variable(handle: RawHandle, result: *mut u32) -> i32;
+    fn __type_test_combination(handle: RawHandle, result: *mut u32) -> i32;
+    fn __type_test_function(handle: RawHandle, result: *mut u32) -> i32;
     fn __type_variables(
         handle: RawHandle,
         result_base: *mut Name,
-        result_length: u64,
+        result_length: *mut u64,
     ) -> i32;
     fn __type_substitute(
         handle: RawHandle,
@@ -134,12 +136,254 @@ where
 /// Returns `ErrorCode::NoSuchType` if any of the handles in `arguments` does
 /// not point-to an allocated type in the kernel's heaps.
 pub fn type_register_combination<T, A>(
-    _type_former: T,
-    _arguments: Vec<A>,
+    type_former: T,
+    arguments: Vec<A>,
 ) -> Result<Handle<tags::Type>, ErrorCode>
 where
     T: Into<Handle<tags::TypeFormer>>,
-    A: Into<Handle<tags::Type>>,
+    A: Into<Handle<tags::Type>> + Clone,
 {
-    unimplemented!()
+    let mut result: u64 = 0;
+
+    let status = unsafe {
+        __type_register_combination(
[…]

### 2021-08-04 libsupervisionary/src/error_code.rs

@@ -107,6 +107,53 @@ pub enum ErrorCode {
 // Trait implementations.
 ////////////////////////////////////////////////////////////////////////////////
 
+/// Pretty-printing for error codes.
+impl Display for ErrorCode {
+    fn fmt(&self, f: &mut Formatter) -> Result<(), DisplayError> {
+        match self {
+            ErrorCode::NoSuchFunction => write!(f, "NoSuchFunction"),
+            ErrorCode::NoSuchConstantRegistered => {
+                write!(f, "NoSuchConstantRegistered")
+            }
+            ErrorCode::NoSuchTermRegistered => {
+                write!(f, "NoSuchTermRegistered")
+            }
+            ErrorCode::NoSuchTheoremRegistered => {
+                write!(f, "NoSuchTheoremRegistered")
+            }
+            ErrorCode::NoSuchTypeFormerRegistered => {
+                write!(f, "NoSuchTypeFormerRegistered")
+            }
+            ErrorCode::MismatchedArity => write!(f, "MismatchedArity"),
+            ErrorCode::DomainTypeMismatch => write!(f, "DomainTypeMismatch"),
+            ErrorCode::NoSuchTypeRegistered => {
+                write!(f, "NoSuchTypeRegistered")
+            }
+            ErrorCode::NotAFunctionType => write!(f, "NotAFunctionType"),
+            ErrorCode::NotATypeCombination => write!(f, "NotATypeCombination"),
+            ErrorCode::NotATypeVariable => write!(f, "NotATypeVariable"),
+            ErrorCode::TypeNotWellformed => write!(f, "TypeNotWellformed"),
+            ErrorCode::NotAConjunction => write!(f, "NotAConjunction"),
+            ErrorCode::NotAConstant => write!(f, "NotAConstant"),
+            ErrorCode::NotAForall => write!(f, "NotAForall"),
+            ErrorCode::NotADisjunction => write!(f, "NotADisjunction"),
+            ErrorCode::NotALambda => write!(f, "NotALambda"),
+            ErrorCode::NotAnApplication => write!(f, "NotAnApplication"),
+            ErrorCode::NotAnEquality => write!(f, "NotAnEquality"),
+            ErrorCode::NotAnExists => write!(f, "NotAnExists"),
+            ErrorCode::NotAnImplication => write!(f, "NotAnImplication"),
+            ErrorCode::NotANegation => write!(f, "NotANegation"),
[…]

### 2021-08-04 libsupervisionary/src/_type.rs
libsupervisionary: added rustdoc comments for Type.* ABI

@@ -55,39 +55,51 @@ pub const PREALLOCATED_HANDLE_TYPE_QUANTIFIER: Handle<tags::Type> =
 ////////////////////////////////////////////////////////////////////////////////
 
 extern "C" {
+    /// Raw ABI binding to the `__type_is_registered` function.
     fn __type_is_registered(handle: RawHandle) -> bool;
+    /// Raw ABI binding to the `__type_register_variable` function.
     fn __type_register_variable(name: Name) -> u64;
+    /// Raw ABI binding to the `__type_register_combination` function.
     fn __type_register_combination(
         type_former_handle: RawHandle,
         argument_base: *const RawHandle,
         argument_length: u64,
         result: *mut RawHandle,
     ) -> i32;
+    /// Raw ABI binding to the `__type_register_function` function.
     fn __type_register_function(
         domain_handle: RawHandle,
         range_handle: RawHandle,
         result: *mut RawHandle,
     ) -> i32;
+    /// Raw ABI binding to the `__type_split_variable` function.
     fn __type_split_variable(handle: RawHandle, result: *mut Name) -> i32;
+    /// Raw ABI binding to the `__type_split_combination` function.
     fn __type_split_combination(
         handle: RawHandle,
         type_former: *mut RawHandle,
         argument_base: *mut RawHandle,
         argument_length: *mut u64,
     ) -> i32;
+    /// Raw ABI binding to the `__type_split_function` function.
     fn __type_split_function(
         handle: RawHandle,
         domain_handle: *mut RawHandle,
         range_handle: *mut RawHandle,
     ) -> i32;
+    /// Raw ABI binding to the `__type_test_variable` function.
     fn __type_test_variable(handle: RawHandle, result: *mut u32) -> i32;
+    /// Raw ABI binding to the `__type_test_combination` function.
     fn __type_test_combination(handle: RawHandle, result: *mut u32) -> i32;
+    /// Raw ABI binding to the `__type_test_function` function.
     fn __type_test_function(handle: RawHandle, result: *mut u32) -> i32;
+    /// Raw ABI binding to the `__type_variables` function.
     fn __type_variables(
         handle: RawHandle,
         result_base: *mut Name,
[…]

### 2021-08-04 libsupervisionary/src/_type.rs
libsupervisionary: part-way through binding of Constant.* ABI

@@ -56,7 +56,7 @@ pub const PREALLOCATED_HANDLE_TYPE_QUANTIFIER: Handle<tags::Type> =
 
 extern "C" {
     /// Raw ABI binding to the `__type_is_registered` function.
-    fn __type_is_registered(handle: RawHandle) -> bool;
+    fn __type_is_registered(handle: RawHandle) -> u32;
     /// Raw ABI binding to the `__type_register_variable` function.
     fn __type_register_variable(name: Name) -> u64;
     /// Raw ABI binding to the `__type_register_combination` function.
@@ -117,7 +117,10 @@ pub fn type_is_registered<H>(handle: H) -> bool
 where
     H: AsRef<Handle<tags::Type>>,
 {
-    unsafe { __type_is_registered(*handle.as_ref().clone() as u64) }
+    let result =
+        unsafe { __type_is_registered(*handle.as_ref().clone() as u64) };
+
+    result == 0
 }
 
 /// Allocates a new type-variable with a given `name`.  Note that this function

### 2021-08-04 libsupervisionary/src/constant.rs

@@ -14,6 +14,7 @@
 //! [Arm Research]: http://www.arm.com/research
 
 use crate::handle::{tags, Handle};
+use crate::RawHandle;
 use std::marker::PhantomData;
 
 ////////////////////////////////////////////////////////////////////////////////
@@ -47,3 +48,29 @@ pub const PREALLOCATED_HANDLE_CONSTANT_EXISTS: Handle<tags::Constant> =
 /// A pre-allocated handle used to refer to the equality constant.
 pub const PREALLOCATED_HANDLE_CONSTANT_EQUALITY: Handle<tags::Constant> =
     Handle::new(18usize, PhantomData);
+
+////////////////////////////////////////////////////////////////////////////////
+// ABI bindings.
+////////////////////////////////////////////////////////////////////////////////
+
+extern "C" {
+    /// Raw ABI binding to the `__constant_is_registered` function.
+    fn __constant_is_registered(handle: RawHandle) -> i32;
+    /// Raw ABI binding to the `__constant_resolve` function.
+    fn __constant_resolve(handle: RawHandle, result: *mut RawHandle) -> i32;
+    /// Raw ABI binding to the `__constant_register` function.
+    fn __constant_register(
+        type_handle: RawHandle,
+        result: *mut RawHandle,
+    ) -> i32;
+}
+
+#[inline]
+pub fn constant_is_registered<H>(handle: H) -> bool
+where
+    H: Into<Handle<tags::Constant>>,
+{
+    let result = unsafe { __constant_is_registered(*handle.into() as u64) };
+
+    result == 0
+}

### 2021-09-30 libsupervisionary/src/build/_type.rs
libsupervisionary: more work on libsupervisionary uncommitted on my machine

@@ -1,0 +1,4 @@
+pub enum Type {
+    Variable(String),
+    Former(),
+}

### 2021-09-30 libsupervisionary/src/handle.rs

@@ -1,145 +1,0 @@
-//! # Kernel handles
-//!
-//! # Authors
-//!
-//! [Dominic Mulligan], Systems Research Group, [Arm Research] Cambridge.
-//!
-//! # Copyright
-//!
-//! Copyright (c) Arm Limited, 2021.  All rights reserved (r).  Please see the
-//! `LICENSE.markdown` file in the *Supervisionary* root directory for licensing
-//! information.
-//!
-//! [Dominic Mulligan]: https://dominic-mulligan.co.uk
-//! [Arm Research]: http://www.arm.com/research
-
-use std::{
-    fmt,
-    fmt::{Display, Formatter},
-    marker::PhantomData,
-    ops::Deref,
-};
-
-////////////////////////////////////////////////////////////////////////////////
-// Handle tags.
-////////////////////////////////////////////////////////////////////////////////
-
-pub mod tags {
-    #[derive(Clone, Debug, Eq, Hash, Ord, PartialEq, PartialOrd)]
-    pub struct TypeFormer;
-
-    #[derive(Clone, Debug, Eq, Hash, Ord, PartialEq, PartialOrd)]
-    pub struct Type;
-
-    #[derive(Clone, Debug, Eq, Hash, Ord, PartialEq, PartialOrd)]
-    pub struct Constant;
-
-    #[derive(Clone, Debug, Eq, Hash, Ord, PartialEq, PartialOrd)]
-    pub struct Term;
-
-    #[derive(Clone, Debug, Eq, Hash, Ord, PartialEq, PartialOrd)]
-    pub struct Theorem;
-
-    pub trait IsTag {}
-
-    impl IsTag for TypeFormer {}
-
-    impl IsTag for Type {}
-
-    impl IsTag for Constant {}
-
-    impl IsTag for Term {}
-
-    impl IsTag for Theorem {}
-}
-
-////////////////////////////////////////////////////////////////////////////////
-// Tagged handles.
-////////////////////////////////////////////////////////////////////////////////
-
-#[derive(Clone, Debug, Eq, Hash, Ord, PartialEq, PartialOrd)]
-pub struct Handle<T>
-where
-    T: tags::IsTag,
-{
-    /// We use the Rust `usize` type as our handle type.  Note that on modern 64-bit
-    /// systems this is implemented as a 64-bit unsigned integer.
-    handle: usize,
-    /// The phantom data binding the tag type, `T`.
-    marker: PhantomData<T>,
-}
-
-impl<T> Handle<T>
-where
[…]

### 2021-09-30 libsupervisionary/src/lib.rs

@@ -19,28 +19,5 @@
 
 #![feature(const_fn_trait_bound)]
 
-/// Bindings to the Supervisionary kernel's type ABI.
-pub mod _type;
-/// Bindings to the Supervisionary kernel's constant ABI.
-pub mod constant;
-/// Kernel error codes.  Note that the material in this module must exactly
-/// match the definition of error code used in the kernel.
-pub mod error_code;
-/// The kernel handle type, and some primitive handles associated with the
-/// Supervisionary ABI.
-pub mod handle;
-/// Bindings to the Supervisionary kernel's term ABI.
-pub mod term;
-/// Bindings to the Supervisionary kernel's theorem ABI.
-pub mod theorem;
-/// Bindings to the Supervisionary kernel's type-former ABI.
-pub mod type_former;
-
-/// The kernel type of names.
-pub type Name = u64;
-/// The kernel type of arities.
-pub type Arity = u64;
-/// The "raw" representation of handles expected by the kernel.
-pub(crate) type RawHandle = u64;
-/// The "raw" representation of kernel error modes.
-pub(crate) type RawKernelFailureMode = i32;
+/// Raw bindings to the Supervisionary kernel ABI.
+pub mod raw;

### 2021-09-30 libsupervisionary/src/raw/_type.rs

@@ -13,14 +13,12 @@
 //! [Dominic Mulligan]: https://dominic-mulligan.co.uk
 //! [Arm Research]: http://www.arm.com/research
 
-use crate::{
-    error_code::ErrorCode,
-    handle::{tags, Handle},
-    Name, RawHandle,
+use std::{
+    collections::HashSet, convert::TryFrom, iter::FromIterator,
+    marker::PhantomData,
 };
 
-use std::iter::FromIterator;
-use std::{collections::HashSet, convert::TryFrom, marker::PhantomData};
+use crate::raw::{tags, ErrorCode, Handle, Name, RawHandle};
 
 ////////////////////////////////////////////////////////////////////////////////
 // Pre-allocated type-related handles.

### 2021-09-30 libsupervisionary/src/raw/constant.rs

@@ -13,13 +13,10 @@
 //! [Dominic Mulligan]: https://dominic-mulligan.co.uk
 //! [Arm Research]: http://www.arm.com/research
 
-use crate::{
-    error_code::ErrorCode,
-    handle::{tags, Handle},
-    RawHandle,
-};
 use std::convert::TryFrom;
 use std::marker::PhantomData;
+
+use crate::raw::{tags, ErrorCode, Handle, RawHandle};
 
 ////////////////////////////////////////////////////////////////////////////////
 // Pre-allocated constant-related handles.

### 2021-09-30 libsupervisionary/src/raw/mod.rs

@@ -1,4 +1,4 @@
-//! # Error codes
+//! # Bindings to Supervisionary's ABI
 //!
 //! # Authors
 //!
@@ -13,10 +13,30 @@
 //! [Dominic Mulligan]: https://dominic-mulligan.co.uk
 //! [Arm Research]: http://www.arm.com/research
 
+use std::marker::PhantomData;
+use std::ops::Deref;
 use std::{
     convert::TryFrom,
+    fmt,
     fmt::{Display, Error as DisplayError, Formatter},
 };
+
+pub mod _type;
+pub mod constant;
+pub mod term;
+pub mod theorem;
+pub mod type_former;
+
+////////////////////////////////////////////////////////////////////////////////
+// Raw types.
+////////////////////////////////////////////////////////////////////////////////
+
+/// The kernel type of names.
+pub type Name = u64;
+/// The kernel type of arities.
+pub type Arity = u64;
+/// The "raw" representation of handles expected by the kernel.
+pub(crate) type RawHandle = u64;
 
 ////////////////////////////////////////////////////////////////////////////////
 // Error codes.
@@ -102,10 +122,6 @@ pub enum ErrorCode {
     /// A theorem passed to a function as an argument was not well-formed.
     TheoremNotWellformed,
 }
-
-////////////////////////////////////////////////////////////////////////////////
-// Trait implementations.
-////////////////////////////////////////////////////////////////////////////////
 
 /// Pretty-printing for error codes.
 impl Display for ErrorCode {
@@ -223,5 +239,121 @@ impl TryFrom<i32> for ErrorCode {
             27 => Ok(ErrorCode::TheoremNotWellformed),
             _otherwise => Err(()),
         }
+    }
+}
+
+////////////////////////////////////////////////////////////////////////////////
+// Kernel handles.
+////////////////////////////////////////////////////////////////////////////////
+
+pub mod tags {
+    #[derive(Clone, Debug, Eq, Hash, Ord, PartialEq, PartialOrd)]
+    pub struct TypeFormer;
+
+    #[derive(Clone, Debug, Eq, Hash, Ord, PartialEq, PartialOrd)]
+    pub struct Type;
+
+    #[derive(Clone, Debug, Eq, Hash, Ord, PartialEq, PartialOrd)]
[…]

### 2021-09-30 libsupervisionary/src/raw/term.rs

@@ -13,15 +13,12 @@
 //! [Dominic Mulligan]: https://dominic-mulligan.co.uk
 //! [Arm Research]: http://www.arm.com/research
 
-use crate::{
-    error_code::ErrorCode,
-    handle::{tags, Handle},
-    Name, RawHandle,
+use std::{
+    collections::HashSet, convert::TryFrom, iter::FromIterator,
+    marker::PhantomData,
 };
-use std::collections::HashSet;
-use std::convert::TryFrom;
-use std::iter::FromIterator;
-use std::marker::PhantomData;
+
+use crate::raw::{tags, ErrorCode, Handle, Name, RawHandle};
 
 ////////////////////////////////////////////////////////////////////////////////
 // Pre-allocated term-related handles.

### 2021-09-30 libsupervisionary/src/raw/theorem.rs

@@ -13,13 +13,9 @@
 //! [Dominic Mulligan]: https://dominic-mulligan.co.uk
 //! [Arm Research]: http://www.arm.com/research
 
-use crate::{
-    error_code::ErrorCode,
-    handle::{tags, Handle},
-    Name, RawHandle,
-};
-
 use std::{convert::TryFrom, marker::PhantomData};
+
+use crate::raw::{tags, ErrorCode, Handle, Name, RawHandle};
 
 ////////////////////////////////////////////////////////////////////////////////
 // ABI bindings.

### 2021-09-30 libsupervisionary/src/raw/type_former.rs

@@ -13,12 +13,9 @@
 //! [Dominic Mulligan]: https://dominic-mulligan.co.uk
 //! [Arm Research]: http://www.arm.com/research
 
-use crate::{
-    error_code::ErrorCode,
-    handle::{tags, Handle},
-    Arity, RawHandle, RawKernelFailureMode,
-};
 use std::{convert::TryFrom, marker::PhantomData};
+
+use crate::raw::{tags, Arity, ErrorCode, Handle, RawHandle};
 
 ////////////////////////////////////////////////////////////////////////////////
 // Pre-allocated type-former handles.
@@ -41,10 +38,7 @@ extern "C" {
     /// Raw ABI binding to the `TypeFormer.IsRegistered` function.
     fn __type_former_is_registered(handle: RawHandle) -> bool;
     /// Raw ABI binding to the `TypeFormer.Resolve` function.
-    fn __type_former_resolve(
-        handle: RawHandle,
-        out: *mut u64,
-    ) -> RawKernelFailureMode;
+    fn __type_former_resolve(handle: RawHandle, out: *mut u64) -> i32;
 }
 
 /// Registers a new type-former with a given `arity`.  Returns the handle to the

### 2021-09-30 paper/prisc-2022.tex
prisc-2022: moved paper from `master` to `main`

\documentclass[sigplan, review]{acmart}
\settopmatter{printfolios=true, printccs=false, printacmref=false}

\usepackage{prooftree}

\newcommand{\rulefont}[1]{(\ensuremath{\mathbf{#1}})}

\bibliographystyle{ACM-Reference-Format}

\title{The Supervisionary proof-checking kernel}
\subtitle{Or: toward proof-generating code}

\author{Dominic P. Mulligan}
\orcid{}
\affiliation{
  \institution{Systems Research Group, Arm Research}
  \streetaddress{Fulbourn Road}
  \city{Cambridge}
  \country{United Kingdom}
}
\email{dominic.mulligan@arm.com}

\author{Nick Spinale}
\orcid{}
\affiliation{
  \institution{Systems Research Group, Arm Research}
  \streetaddress{Fulbourn Road}
  \city{Cambridge}
  \country{United Kingdom}
}
\email{nick.spinale@arm.com}

\begin{document}

\maketitle

To enforce system-wide security policies, most operating systems implement a self-contained \emph{kernel} which serves as the root of all trust within the system.
The kernel is the sole component able to directly interact with hardware, and provides services to executing user-space software, for example the abstraction of a hierarchical filesystem.
Untrusted user-space code resides outside of the kernel and interacts with it through a defined \emph{ABI}.

To enforce logical soundness, many proof-checking systems implement a self-contained \emph{kernel} which serves as the root of all trust within the system.
The kernel is the sole system component able to authenticate a proof as legitimate.
Untrusted code resides outside of the kernel and interacts with it through a defined \emph{API}.

Though ostensibly wildly different, in \emph{some} respects the goals of operating system and proof-checking kernels are the same: the maintenance of important system invariants in the face of unbounded interaction with untrusted code.
The mechanisms employed to protect these invariants, by each type of system, are different however.
For operating systems, machine-oriented notions of separation are employed, with the kernel executing at a higher prilege level than untrusted code, and cosseted by paged memory protection---hardware and privileged software work together to provide an abstract machine interface to untrusted code.
On the other hand, LCF-style proof-checking kernels are typically implemented as libraries within a \emph{meta-language}---usually an ML dialect---with invariants maintained through type abstraction and code encapsulation behind module boundaries.
Logical soundness follows from type-soundness of the meta-language.

Yet, what happens if we structure proof-checking software more like supervisory system software---such as operating systems, hypervisors, and programming language runtimes?
Now, the proof-checking kernel is not implemented as a library, but rather becomes a programming platform, exposing an abstract machine to untrusted code, in much the same way that the Linux kernel, for example, exposes an abstract machine to user-space code through its system call interface.
We are currently developing a prototype system, \emph{Supervisionary}, implemented as a WebAssembly host, to explore the consequences of this design.

One immediate consequence of Supervisionary's design is the abolition of the LCF-style meta-language.
Rather, Supervisionary's kernel is implemented in Rust---a notably \emph{unsafe} systems language.
Moreover, untrusted code which ``drives'' the Supervisionary kernel to produce an authenticated theorem need not necessarily be written in the same language that the kernel is written in: rather any language that can be compiled to code that is binary-compatible with the Supervisionary abstract machine can be used, even a mixture of them, in much the same way that Linux user-space software can be written in Haskell, Ada, or any other language that can be linked together, despite the kernel being written in C.

Note that the use of Rust to implement the Supervisionary kernel does not pose any risk to the logical soundness of the system, providing we expose a carefully designed abstract machine through the Supervisionary ABI.
The Supervisionary kernel manages various private memory spaces, or heaps, within which \emph{kernel objects} are managed.
These kernel objects correspond to the parephenalia of a typical HOL implementation---type-formers, types, term constants, terms, and theorems---but are never directly exposed to untrusted code.
Rather, kernel objects are allocated by the kernel in one of its private heaps, through ABI functions like the following:
\begin{displaymath}
\mathtt{Term.Handle.AllocateApplication(left, right, out)}
\end{displaymath}
Above, both $\mathtt{left}$ and $\mathtt{right}$ passed to the function are kernel handles, assumed to point-to allocated terms.
On the other hand, the $\mathtt{out}$ pointer points-into the heap of the code calling the function (note that the kernel can freely read from, and write to, the memory of untrusted code, given it is supervisory software).
Providing both $\mathtt{left}$ and $\mathtt{right}$ indeed point-to allocated terms, and their types match, a new kernel object corresponding to a term application is allocated, with internal pointers to the functional- and argument-terms, in the kernel's heap.
This new object is pointed-to by a freshly-generated handle, returned back to the caller via the $\mathtt{out}$ pointer.

Note that Supervisionary ABI functions do not panic, or otherwise abort, modulo bugs in the implementation of the kernel itself.
Rather, errors are signalled to calling code via a defined \emph{error code} which can be used to deduce the cause of failure: if either $\mathtt{left}$ or $\mathtt{right}$ above dangle, then the $\mathtt{DanglingHandle}$ code is returned, for example.
Note that these checks for dangling handles---which all Supervisionary ABI functions perform---ensures that the kernel heaps remain \emph{inductive}, or ``downward closed'', with nodes in the kernel object graph always pointing to previously-allocated objects at all times---a basic kernel invariant.

Post allocation, all manipulation and querying of kernel objects is performed by the kernel itself.
For example, to check if a handle points-to a $\lambda$-abstraction, untrusted code can call the following function:
\begin{displaymath}
\mathtt{Term.Handle.IsLambda(handle, out)}
\end{displaymath}
This again returns a defined error code if $\mathtt{handle}$ dangles, leaving $\mathtt{out}$ unchanged, or otherwise writing $\mathtt{true}$ or $\mathtt{false}$ to the $\mathtt{out}$ pointer as appropriate.

Unfortunately, space constraints here prevent us from providing a pr\`{e}cis of all ABI functions for allocating, querying, destructuring, and otherwise manipulating type-formers, types, term constants, and terms.
However, note that theorems are also gradually constructed in an analogous way to how terms are constructed.
For example, the HOL symmetry inference rule is exposed as the ABI function
\begin{displaymath}
\mathtt{Theorem.Handle.AllocateSym(pre, out)}
\end{displaymath}
where $\mathtt{pre}$ is assumed to point-to an allocated theorem $\Gamma \vdash r = s$ and $\mathtt{out}$ points-to a freshly-allocated handle that itself points-to a newly-allocated theorem $\Gamma \vdash s = r$, should the function succeed.
Note that the heap of theorem objects must only ever grow monotonically, and a theorem object must always remain immutable once allocated by the kernel.
This immutable aspect also applies to any kernel object mentioned by any allocated theorem---terms, types, and so on: it would be decidedly \emph{inconvenient} if the meaning of the HOL truth constant were redefined, for example.
On the other hand, it seems possible that some kernel heaps could shrink, for example allocated type-formers, types, term constants, and terms, that are not referenced transitively by any allocated theorem could be garbage collected.
This is not yet implemented in Supervisionary.

Note that one consequence of this style of implementation is the ability to provide concise \emph{specifications} of Supervisionary ABI functions.
Essentially, the Supervisionary kernel is a grand exercise in pointer manipulation, and as such our ABI specifications can be expressed as Hoare Triples, using Separation Logic as our assertion language.
Writing $\mathtt{h} \mapsto_{\mathtt{trm}} \mathtt{Application(l, r)}$ to assert that the handle $\mathtt{h}$ points-to a term application (of the term pointed-to by $l$ to the term pointed-to by $r$), and writing $\mathtt{out} \mapsto \mathtt{b}$ to assert that $\mathtt{out}$ points-to the Boolean value $\mathtt{b}$, we have, for example:
\begin{gather*}
\{ h \mapsto_{\mathtt{trm}} \mathtt{Application(l, r)} \} \\
\mathtt{Term.Handle.IsApplication(h, out)} \\
\{ \mathtt{out} \mapsto \mathtt{True} \}
\end{gather*}
(Here, the triple $\{ P \} C \{ Q \}$ asserts that if the command $C$ executes in a state concordant with $P$ then the command succeeds and produces a state concordant with $Q$.)

The material presented so far is implemented in Supervisionary.
We now speculate on opportunities for future work.

First, observe that Supervisionary is not merely a rehashing of LCF, perhaps with some minor tweaks.
In particular, Supervisionary ``internalises'' theorems as kernel objects in a way that LCF does not: the latter authenticates theorems by producing an opaque object, of a particular abstract type, which in a sense``passes out'' of the kernel.
Moreover, Supervisionary behaves like a typical operating system, or programming language runtime, in that it takes a monopoly on the system's resources from the point of view of untrusted code.
To access the filesystem, or a device for example, untrusted code generally has no choice but to interact with the operating system, which gates access based on a prevailing security policy. 
Adding ABI functions to Supervisionary for filesystem access allows WebAssembly programs executing under Supervisionary to access the system's filesystem, at the cost of passing through the Supervisionary kernel.
What, then, if we force code wishing to interact with the filesystem (or any other system resource) to provide a particular \emph{theorem} to the kernel before that access is granted?

\end{document}

### 2021-09-30 paper/prisc-2022.tex
prisc-2022: reorganising paper using paragraphs

@@
 Though ostensibly wildly different, in \emph{some} respects the goals of operating system and proof-checking kernels are the same: the maintenance of important system invariants in the face of unbounded interaction with untrusted code.
-The mechanisms employed to protect these invariants, by each type of system, are different however.
-For operating systems, machine-oriented notions of separation are employed, with the kernel executing at a higher prilege level than untrusted code, and cosseted by paged memory protection---hardware and privileged software work together to provide an abstract machine interface to untrusted code.
-On the other hand, LCF-style proof-checking kernels are typically implemented as libraries within a \emph{meta-language}---usually an ML dialect---with invariants maintained through type abstraction and code encapsulation behind module boundaries.
-Logical soundness follows from type-soundness of the meta-language.
+The mechanisms employed to protect these invariants, by each system, are different however.
+Operating systems kernels are cosseted by paged memory protection and execute at a relative level of privilege compared to untrusted code.
+On the other hand, LCF-style proof-checking kernels are typically implemented as libraries within a \emph{meta-language}---usually an ML dialect---with invariants maintained by type abstraction and module encapsulation.
+Logical soundness is ultimately guaranteed by type-soundness of the meta-language.
 
-Yet, what happens if we structure proof-checking software more like supervisory system software---such as operating systems, hypervisors, and programming language runtimes?
-Now, the proof-checking kernel is not implemented as a library, but rather becomes a programming platform, exposing an abstract machine to untrusted code, in much the same way that the Linux kernel, for example, exposes an abstract machine to user-space code through its system call interface.
+What happens if we structure proof-checking software like \emph{supervisory} system software---such as operating systems, hypervisors, and programming language runtimes?
 We are currently developing a prototype system, \emph{Supervisionary}, implemented as a WebAssembly host, to explore the consequences of this design.
+In one sense, Supervisionary can be used as a standard proof-checker for Gordon's HOL, driven by an untrusted WebAssembly program.
+However, in another sense, Supervisionary exposes a novel abstract machine to untrusted code, in much the same way that the Linux kernel exposes an abstract machine to user-space code through its system call interface, for example.
+We pr\`{e}cis some of the consequences of Supervisionary's design below, and argue that this novel abstract machine is of interest.
 
-One immediate consequence of Supervisionary's design is the abolition of the LCF-style meta-language.
-Rather, Supervisionary's kernel is implemented in Rust---a notably \emph{unsafe} systems language.
-Moreover, untrusted code which ``drives'' the Supervisionary kernel to produce an authenticated theorem need not necessarily be written in the same language that the kernel is written in: rather any language that can be compiled to code that is binary-compatible with the Supervisionary abstract machine can be used, even a mixture of them, in much the same way that Linux user-space software can be written in Haskell, Ada, or any other language that can be linked together, despite the kernel being written in C.
+\paragraph{No meta-language} One immediate consequence of Supervisionary's design is the abolition of the LCF meta-language.
+Rather, Supervisionary's kernel is implemented in Rust.
+On the other hand, untrusted code which ``drives'' the Supervisionary kernel to produce an authenticated theorem may be written in any language that can be compiled to code that is binary-compatible with the Supervisionary abstract machine, or even a mixture of them, in much the same way that Linux user-space software can be written in Haskell, Ada, or any other language that can be linked together, despite the kernel being written in C.
 
-Note that the use of Rust to implement the Supervisionary kernel does not pose any risk to the logical soundness of the system, providing we expose a carefully designed abstract machine through the Supervisionary ABI.
+\paragraph{ABI design} Rust is a notably \emph{unsafe} systems programming language, yet its use in the Supervisionary kernel does not pose any risk to the logical soundness of the system, providing we expose a carefully designed ABI.
 The Supervisionary kernel manages various private memory spaces, or heaps, within which \emph{kernel objects} are managed.
@@
 
-Note that one consequence of this style of implementation is the ability to provide concise \emph{specifications} of Supervisionary ABI functions.
-Essentially, the Supervisionary kernel is a grand exercise in pointer manipulation, and as such our ABI specifications can be expressed as Hoare Triples, using Separation Logic as our assertion language.
-Writing $\mathtt{h} \mapsto_{\mathtt{trm}} \mathtt{Application(l, r)}$ to assert that the handle $\mathtt{h}$ points-to a term application (of the term pointed-to by $l$ to the term pointed-to by $r$), and writing $\mathtt{out} \mapsto \mathtt{b}$ to assert that $\mathtt{out}$ points-to the Boolean value $\mathtt{b}$, we have, for example:
+\paragraph{Specifying the ABI}
+Supervisionary is a grand exercise in pointer manipulation.
+As such, our ABI specifications can be expressed as Hoare Triples, using Separation Logic as our assertion language.
+Write $\mathtt{h} \mapsto_{\mathtt{trm}} \mathtt{Application(l, r)}$ to assert that the handle $\mathtt{h}$ points-to a term application (of the term pointed-to by $l$ to the term pointed-to by $r$) in the kernel's private heap.
+Write $\mathtt{out} \mapsto \mathtt{b}$ to assert that $\mathtt{out}$ points-to the Boolean value $\mathtt{b}$ in untrusted code's heap.
+We then have:
 \begin{gather*}
@@
 \end{gather*}
-(Here, the triple $\{ P \} C \{ Q \}$ asserts that if the command $C$ executes in a state concordant with $P$ then the command succeeds and produces a state concordant with $Q$.)
+Here, the triple $\{ P \} C \{ Q \}$ asserts that if the command $C$ executes in a state concordant with $P$ then the command succeeds and produces a state concordant with $Q$.
 
-The material presented so far is implemented in Supervisionary.
-We now speculate on opportunities for future work.
+\paragraph{Runtime monitoring}
 
@@
 
+\paragraph{Reinventing the proof-assistant}
+
 \end{document}

### 2021-09-30 paper/prisc-2022.tex
prisc-2022: some more minor rewriting, tightening, etc.

@@
 The mechanisms employed to protect these invariants, by each system, are different however.
-Operating systems kernels are cosseted by paged memory protection and execute at a relative level of privilege compared to untrusted code.
-On the other hand, LCF-style proof-checking kernels are typically implemented as libraries within a \emph{meta-language}---usually an ML dialect---with invariants maintained by type abstraction and module encapsulation.
-Logical soundness is ultimately guaranteed by type-soundness of the meta-language.
+Operating systems kernels are cosseted by paged memory protection and execute at a relative level of privilege compared to untrusted code, which interacts with the kernel through a binary interface.
+On the other hand, LCF-style proof-checking kernels are typically implemented as libraries within a \emph{meta-language}---usually an ML dialect---with invariants maintained by type abstraction and module encapsulation, and linked against untrusted code typically written in the meta-language.
 
@@
 We are currently developing a prototype system, \emph{Supervisionary}, implemented as a WebAssembly host, to explore the consequences of this design.
-In one sense, Supervisionary can be used as a standard proof-checker for Gordon's HOL, driven by an untrusted WebAssembly program.
-However, in another sense, Supervisionary exposes a novel abstract machine to untrusted code, in much the same way that the Linux kernel exposes an abstract machine to user-space code through its system call interface, for example.
-We pr\`{e}cis some of the consequences of Supervisionary's design below, and argue that this novel abstract machine is of interest.
+In one sense, Supervisionary can be used as a standard proof-checker for Gordon's HOL, driven by an untrusted WebAssembly program that interacts with the kernel via a binary interface which defines an abstract machine, in much the same way that the Linux kernel exposes an abstract machine to user-space code through its system call interface, for example.
+However, in another sense, Supervisionary can be generalised to a programming environment that can host arbitrary programs.
+We briefly describe some of the consequences of Supervisionary's design below, and argue that this novel abstract machine is of interest.
 
-\paragraph{No meta-language} One immediate consequence of Supervisionary's design is the abolition of the LCF meta-language.
+\paragraph{No meta-language} One immediate consequence of Supervisionary's design is the lack of any meta-language.
 Rather, Supervisionary's kernel is implemented in Rust.
@@
 \end{displaymath}
-Above, both $\mathtt{left}$ and $\mathtt{right}$ passed to the function are kernel handles, assumed to point-to allocated terms.
-On the other hand, the $\mathtt{out}$ pointer points-into the heap of the code calling the function (note that the kernel can freely read from, and write to, the memory of untrusted code, given it is supervisory software).
-Providing both $\mathtt{left}$ and $\mathtt{right}$ indeed point-to allocated terms, and their types match, a new kernel object corresponding to a term application is allocated, with internal pointers to the functional- and argument-terms, in the kernel's heap.
-This new object is pointed-to by a freshly-generated handle, returned back to the caller via the $\mathtt{out}$ pointer.
+Here, $\mathtt{left}$ and $\mathtt{right}$ are kernel handles, assumed to point-to allocated terms in the kernel's heaps.
+On the other hand, the $\mathtt{out}$ pointer points-into the heap of the code calling the function---the kernel can freely read from, and write to, the memory of untrusted code.
+Assuming $\mathtt{left}$ and $\mathtt{right}$ indeed point-to allocated terms, and other checks pass, a new kernel object corresponding to a term application is allocated.
+This object contains internal pointers to the functional- and argument-terms, in the kernel's heap.
+The kernel generates a fresh handle that points-to this new object, which is returned back to the caller via the $\mathtt{out}$ pointer.
 

### 2021-10-07 paper/prisc-2022.tex
prisc-2022: moved paper material from master to main

--- 2021-09-30 paper/prisc-2022.tex
@@
 
+\newcommand{\deffont}[1]{\textbf{#1}}
+\newcommand{\fall}[1]{\forall{#1}.}
+\newcommand{\lam}[1]{\lambda{#1}.}
 \newcommand{\rulefont}[1]{(\ensuremath{\mathbf{#1}})}
@@
 \bibliographystyle{ACM-Reference-Format}
+
+\setcopyright{none}
 
@@
 \maketitle
+
+\paragraph{Some scene setting}
+
+Modern operating systems are used for a wide array of tasks.
+We focus on just two: providing a relatively portable abstract machine interface to user-space software, and enforcing system-wide security policies.
+
+Most operating systems are architected around a \deffont{kernel}, the root of all trust within a system, which engages in a grand conspiracy with its host microprocessor to take a monopoly on direct access to the system's hardware and sensitive system configuration options.
+%Relative to user-space code, the kernel executes at a high level of \emph{privilege}, using hardware \deffont{exception levels}.
+%Hardware ensures that sensitive system configuration features, such as system registers and page tables, may only be modified by software executing at a high level of privilege---in effect, only the kernel itself.
+Using this monopoly, the kernel protects its own memory from inadvertent or malicious modification, using paged memory protection, and acts as a ``choke point'' at which security policies are enforced.
+
+%Users-space code wishing to interact with a system resource must do so by interacting with the kernel.
+
+The kernel also offers an abstract machine interface to user-space software, abstracting over the vulgarities of programming against any particular peripheral, say.
+Useful abstractions, such as sockets and a file-system, are also commonly offered.
+This interface is largely exposed as \deffont{system calls}, defining an \deffont{application binary interface}, or ABI henceforth, which user-space and kernel must understand to interact.
+%Executing a system call forces a \deffont{context switch}, within which the thread of execution enters the kernel.
+The result of a system call may be written directly into the memory of the calling code, via a pointer argument, for example.
+Note an asymmetry: in theory, the kernel can read and modify all of a user-space program's state.
+
+Code written in any language can make use of services via the kernel ABI.
+However, programs and runtimes typically make extensive use of low-level libraries---\texttt{libc} being an exemplar---that provide further abstraction over the raw ABI.
+
+This pattern of interaction between trusted and untrusted code via an ABI, wherein trusted code can read and modify the state of untrusted code, is common.
+Hypervisors, the Secure World of Arm TrustZone, and virtual machines (in the Wasm sense) are structured in this way.
+Call software designed in this pattern \deffont{supervisory software}.
+
+\paragraph{Supervisionary} is a prototype proof-checking system for Gordon's HOL structured as supervisory software, specifically as a Wasm \deffont{host environment}, which allows us to explore the consequences of the desig, without becoming entangled in e.g.
+a real machine's boot ceremony.
+The ideas discussed herein could be applied to a real operating system.
+
+Supervisionary is implemented in Rust which does not pose any risk to the logical soundness of the system providing we expose a well-designed ABI.
+The kernel manages various private heaps within which \deffont{kernel objects} are allocated, corresponding to the paraphenalia of any HOL implementation---type-formers, types, term constants, terms, and theorems---and never directly exposed to untrusted code.
+Kernel objects are allocated in response to system calls like:
+\begin{displaymath}
+\mathtt{Term.Handle.AllocateApplication(left, right, out)}
+\end{displaymath}
+Here, both $\mathtt{left}$ and $\mathtt{right}$ are \deffont{kernel handles}, assumed to point-to allocated terms, whilst $\mathtt{out}$ points-to a buffer in untrusted code's memory.
+If neither $\mathtt{left}$ nor $\mathtt{right}$ \deffont{dangle}, and the types of their referents match, a fresh handle is generated which points-to a new HOL term application object, with internal pointers to the functional- and argument-terms.
+This handle is returned to the caller via the $\mathtt{out}$ pointer.
+
+Modulo bugs, system calls do not panic, but rather return \deffont{error codes} to signal failure.
+If $\mathtt{left}$ or $\mathtt{right}$ above dangle, for example, then a defined code is returned.
+These checks---which all system calls perform---ensure that the kernel heaps remain \deffont{inductive}, with nodes in the kernel object graph pointing-to allocated objects at all times, a key invariant.
+Moreover, kernel objects are only ever allocated, and \emph{never} deallocated or otherwise modified once created, thus ensuring the meaning of an object remains immutable (though it is possible that some \deffont{garbage collection} could be used).
+
+%To check if a handle points-to a $\lambda$-abstraction, for example, one may use:
+%\begin{displaymath}
+%\mathtt{Term.Handle.IsLambda(handle, out)}
+%\end{displaymath}
+%This again returns a defined error code if $\mathtt{handle}$ dangles and otherwise writes $\mathtt{true}$ or $\mathtt{false}$ to $\mathtt{out}$, as appropriate.
+
+The manipulation and querying of kernel objects is performed by the kernel itself on behalf of untrusted code.
+Space constraints prevent us from describing the entire Supervisionary ABI for working with and on kernel objects.
+However, note that theorems are also constructed in an analogous way to terms.
+The HOL symmetry rule is exposed as:
+\begin{displaymath}
+\mathtt{Theorem.Handle.AllocateSym(pre, out)}
+\end{displaymath}
+Here, $\mathtt{pre}$ points-to a theorem $\Gamma \vdash r = s$ and after succeeding, $\mathtt{out}$ contains a handle that points-to a newly-allocated theorem $\Gamma \vdash s = r$.
+
+We have described the \emph{what} of Supervisionary and now turn our attention to the \emph{why}.
+This can be approached by examining Supervisionary both as a programming platform, and as a proof-assistant kernel.
+What follows is speculative, and a work-in-progress.
+
+\paragraph{Runtime monitoring}
+
+Nothing in principle stops us from extending Supervisionary's ABI with facilities for sockets, querying a hierarchical filesystem, and other similar system tasks.
+In particular, given our use of Wasm, we could implement the Wasm System Interface, a POSIX-like interface for Wasm.
+This transforms Supervisionary from a simple programmable proof-checker into a general-purpose virtual machine, capable of supporting a wide-range of programs.
+
+\emph{What happens if we blur the lines between Supervisionary's ABIs for system access and proof-checking?}
+In particular, Supervisionary could force untrusted code to prove some theorem each time it wished to open a file on the filesystem, write to a socket, or otherwise perform some side-effect on the system.
+These theorems could be arbitrary correctness or security-related theorems, corresponding to a prevailing \deffont{policy} in force.
+Moreover, as Supervisionary is supervisory software, and capable of capturing the runtime state of untrusted code, the statements of these theorems can be arbitrary HOL predicates and functions of the reified runtime state of both untrusted code, the kernel itself, and the arguments, and name of, the invoked system call.
+
+Obvious predicates of interest include $\lam{k}\lam{u}\lam{s}\top$ and $\lam{k}\lam{u}\lam{s}\bot$.
+Here, $\top$ and $\bot$ are truth and falsity, respectively, and $k$, $u$, and $s$ the kernel and untrusted code states, and packed system call meta-data, reified as HOL types, at the point the system call was invoked.
+One can always prove $\vdash \top$, trivially, and therefore $\lam{k}\lam{u}\lam{s}\top$ represents no restriction, whereas $\vdash \bot$ is never provable in a consistent context, and $\lam{k}\lam{u}\lam{s}\bot$ therefore represents a ``closing off'' of a system call.
+Note that, by making the property a function of the system call name and arguments, this ``closing off'' can be very specific: a process could be banned from calling a particular system call, or a system call with a particular set of arguments, for example.
+This would allow us to mimic existing security mechanisms, like \texttt{seccomp} from Linux.
+
+We can go further.
+Supervisionary could maintain meta-data about the behaviour of a running process, under its supervision: for example, a trace of the system calls invoked by a process, thus far.
+This meta-data could be used in forming security or correctness challenges, for example, by forcing untrusted code to prove that only a certain number of files have been written, or writes to a socket only ever happen after a read.
+Correctness or information flow properties could also be captured: by requiring a demonstration that there exists a pure HOL function that, when applied to input from a particular file, produces an output written to another file, we are able to rule out information leakage from any other file into the result.
+In short, HOL becomes a \emph{lingua franca} between kernel and untrusted code through which arbitrarily complex security and correctness policies may be expressed.
+
+\deffont{Jailing}, through which a process voluntarily sheds capabilities, is common.
+A similar pattern can be captured by allowing a change in the prevailing policy, $O$, to a new policy $N$, as long as the process is able to prove that $N$ is a refinment of $O$: $\vdash \fall{k}\fall{u}\fall{s} O\ k\ u\ s \longrightarrow N\ k\ u\ s$.
+Note that this essentially states that the machine states described by $N$ are a subset of those described by $O$.
+
+Note that \deffont{proof carrying code} modified the operating system loader, and endowed binaries with skeleton proofs of adherence to some policy which was reconstructed by the loader, prior to execution.
+The idea sketched above represents something different, where proofs are generated dynamically, as the program executes, and is perhaps more aptly described as \deffont{proof generating code}.
+
+\paragraph{Specifications}
+
+Supervisionary is an extended exercise in pointer manipulation.
+The semantics of our system calls can therefore be expressed as Hoare Triples, using formulae of \deffont{Separation Logic} as an assertion language.
+Write $\mathtt{h} \mapsto_{\mathtt{trm}} \mathtt{Application(l, r)}$ to assert that the handle $\mathtt{h}$ points-to a term application (of the term pointed-to by $l$ to the term pointed-to by $r$).
+Write $\mathtt{out} \mapsto \mathtt{b}$ to assert that $\mathtt{out}$ points-to the Boolean value $\mathtt{b}$.
+Finally, write $\{ P \} C \{ Q \}$ to assert that $C$ executing in a state satisfying $P$ succeeds and produces a state satisfying $Q$.
+We then have, for example:
+\begin{gather*}
+\{ h \mapsto_{\mathtt{trm}} \mathtt{Application(l, r)} \} \\
+\mathtt{Term.Handle.IsApplication(h, out)} \\
+\{ \mathtt{out} \mapsto \mathtt{True} \}
+\end{gather*}
+Note that the structural \deffont{frame rule} captures the fact that any Supervisionary system call only manipulates or queries a particular limited ``footprint'' of the kernel state.
+
+\paragraph{Reinventing the proof-assistant}
+
+Proof-assistants in the LCF lineage are also architected around a \deffont{kernel}.
+Ostensibly these have little in common with the kernels discussed so far, but a closer examination shows that both are tasked with preserving important system invariants in the face of unbounded interaction with untrusted code.
+Moreover, whilst operating system kernels enjoy a monopoly on direct access to the system, proof-assistant kernels likewise have a monopoly on the \deffont{authentication} of theorems.
+In short, both act as the roots of all trust in their respective domains.
+Yet the two types of kernel are implemented in very different ways, with LCF kernels implemented as libraries in a \deffont{meta-language}---usually some functional language, a descendent of Milner's ML.
+Untrusted code interacts with the kernel via an \emph{API}, rather than an ABI.
+
+One immediate consequence of Supervisionary's design is that the meta-language is non-existent.
+Rather, as noted above, untrusted code ``driving'' the kernel can be written in any language capable of producing Wasm that is binary-compatible with our ABI.
+Arguably, the \emph{natural} choice of language for this is now not even functional, given the inherently stateful and pointer-oriented nature of the kernel.
+
+Observe also that interaction with an LCF-style proof-assistant is centred around the meta-language's REPL, or some thin GUI layer atop, itself started from an operating system shell.
+Supervisionary's design allows us to potentially dispense with this middle-man, and develop proofs interactively directly in a shell implemented on top of our kernel.
+Moreover, and in contrast to LCF, our kernel inherently internalises kernel objects and could make them available to browse as special files within a hierarchical filesystem, made available to untrusted code as discussed previously, in analogy to how Linux exposes its internals as special files (e.g.
+the \texttt{/proc/} file system).
+In this way, hierarchical theories of related material appear grouped under directories, and can be searched using Unix-like command-line utilities.
+\end{document}
 
@@
 However, note that theorems are also gradually constructed in an analogous way to how terms are constructed.
-For example, the HOL symmetry inference rule is exposed as the ABI function
+For example, the HOL symmetry inference rule is exposed as:
 \begin{displaymath}
@@
 \end{displaymath}
-where $\mathtt{pre}$ is assumed to point-to an allocated theorem $\Gamma \vdash r = s$ and $\mathtt{out}$ points-to a freshly-allocated handle that itself points-to a newly-allocated theorem $\Gamma \vdash s = r$, should the function succeed.
+Here, $\mathtt{pre}$ points-to an allocated theorem $\Gamma \vdash r = s$ and after successfully completing, $\mathtt{out}$ contains a handle that points-to a newly-allocated theorem $\Gamma \vdash s = r$.
 Note that the heap of theorem objects must only ever grow monotonically, and a theorem object must always remain immutable once allocated by the kernel.
@@
 What, then, if we force code wishing to interact with the filesystem (or any other system resource) to provide a particular \emph{theorem} to the kernel before that access is granted?
-
-\end{document}

### 2021-10-07 paper/prisc-2022.tex
prisc-2022: squashed paper into page limit

@@
 
-The kernel also offers an abstract machine interface to user-space software, abstracting over the vulgarities of programming against any particular peripheral, say.
+The kernel also offers an abstract machine interface to user-space software, smoothing over the vulgarities of programming against any particular peripheral, say.
 Useful abstractions, such as sockets and a file-system, are also commonly offered.
@@
 The result of a system call may be written directly into the memory of the calling code, via a pointer argument, for example.
-Note an asymmetry: in theory, the kernel can read and modify all of a user-space program's state.
+Note the asymmetry: in theory, the kernel can read and modify all of a user-space program's state.
 
@@
 
-\paragraph{Supervisionary} is a prototype proof-checking system for Gordon's HOL structured as supervisory software, specifically as a Wasm \deffont{host environment}, which allows us to explore the consequences of the desig, without becoming entangled in e.g.
+\paragraph{Supervisionary} is a prototype proof-checking system for Gordon's HOL structured as supervisory software, specifically as a Wasm \deffont{host environment}, which allows us to explore the consequences of the design, without becoming entangled in e.g.
 a real machine's boot ceremony.
@@
 The manipulation and querying of kernel objects is performed by the kernel itself on behalf of untrusted code.
-Space constraints prevent us from describing the entire Supervisionary ABI for working with and on kernel objects.
-However, note that theorems are also constructed in an analogous way to terms.
-The HOL symmetry rule is exposed as:
+Space constraints prevent us from describing the entire Supervisionary ABI for working with, and on, kernel objects.
+However, note that theorems are also constructed in a similar way to terms, inductively building derivation trees.
+For example, the HOL symmetry rule is exposed as:
 \begin{displaymath}
@@
 \end{displaymath}
-Here, $\mathtt{pre}$ points-to a theorem $\Gamma \vdash r = s$ and after succeeding, $\mathtt{out}$ contains a handle that points-to a newly-allocated theorem $\Gamma \vdash s = r$.
+Here, $\mathtt{pre}$ points-to a theorem $\Gamma \vdash r = s$ and after succeeding, passing obvious checks, $\mathtt{out}$ contains a handle that points-to a newly-allocated theorem $\Gamma \vdash s = r$.
 
@@
 This can be approached by examining Supervisionary both as a programming platform, and as a proof-assistant kernel.
-What follows is speculative, and a work-in-progress.
+What follows is largely speculative, and a work-in-progress.
 
@@
 Nothing in principle stops us from extending Supervisionary's ABI with facilities for sockets, querying a hierarchical filesystem, and other similar system tasks.
-In particular, given our use of Wasm, we could implement the Wasm System Interface, a POSIX-like interface for Wasm.
-This transforms Supervisionary from a simple programmable proof-checker into a general-purpose virtual machine, capable of supporting a wide-range of programs.
+In particular, given our use of Wasm, we could implement the Wasm System Interface---a POSIX-like interface for Wasm---transforming Supervisionary from a programmable proof-checker into a general-purpose virtual machine, capable of supporting a wide-range of programs.
 
 \emph{What happens if we blur the lines between Supervisionary's ABIs for system access and proof-checking?}
-In particular, Supervisionary could force untrusted code to prove some theorem each time it wished to open a file on the filesystem, write to a socket, or otherwise perform some side-effect on the system.
-These theorems could be arbitrary correctness or security-related theorems, corresponding to a prevailing \deffont{policy} in force.
-Moreover, as Supervisionary is supervisory software, and capable of capturing the runtime state of untrusted code, the statements of these theorems can be arbitrary HOL predicates and functions of the reified runtime state of both untrusted code, the kernel itself, and the arguments, and name of, the invoked system call.
+Untrusted code could be challenged to prove some theorem each time it wished to open a file on the filesystem, or otherwise perform some side-effect.
+These theorems could be correctness or security-related theorems, corresponding to a prevailing \deffont{policy} in force.
+Moreover, as Supervisionary is capable of capturing the runtime state of untrusted code, these challenges can be HOL predicates---functions of the reified runtime state of untrusted code, the kernel, and the arguments, and name of, the invoked system call.
 
-Obvious predicates of interest include $\lam{k}\lam{u}\lam{s}\top$ and $\lam{k}\lam{u}\lam{s}\bot$.
-Here, $\top$ and $\bot$ are truth and falsity, respectively, and $k$, $u$, and $s$ the kernel and untrusted code states, and packed system call meta-data, reified as HOL types, at the point the system call was invoked.
-One can always prove $\vdash \top$, trivially, and therefore $\lam{k}\lam{u}\lam{s}\top$ represents no restriction, whereas $\vdash \bot$ is never provable in a consistent context, and $\lam{k}\lam{u}\lam{s}\bot$ therefore represents a ``closing off'' of a system call.
-Note that, by making the property a function of the system call name and arguments, this ``closing off'' can be very specific: a process could be banned from calling a particular system call, or a system call with a particular set of arguments, for example.
-This would allow us to mimic existing security mechanisms, like \texttt{seccomp} from Linux.
+Two predicates of interest are $\lam{k}\lam{u}\lam{s}\top$ and $\lam{k}\lam{u}\lam{s}\bot$.
+Here, $\top$ and $\bot$ are truth and falsity, and $k$, $u$, and $s$ the kernel and untrusted code states, and packed system call metadata, reified as HOL data, at the point of invocation of the system call.
+One can always prove $\{\} \vdash \top$ and therefore $\lam{k}\lam{u}\lam{s}\top$ represents no restriction, whereas $\{ \} \vdash \bot$ is never provable, absent axioms, with $\lam{k}\lam{u}\lam{s}\bot$ representing a closing off of a system call.
+By making the challenge a function of the system call name and arguments, this closing off can be specific, banning a process from calling a particular system call, or a system call with a particular set of arguments, allowing us to mimic mechanisms like \texttt{seccomp} from Linux.
 
 We can go further.
-Supervisionary could maintain meta-data about the behaviour of a running process, under its supervision: for example, a trace of the system calls invoked by a process, thus far.
-This meta-data could be used in forming security or correctness challenges, for example, by forcing untrusted code to prove that only a certain number of files have been written, or writes to a socket only ever happen after a read.
-Correctness or information flow properties could also be captured: by requiring a demonstration that there exists a pure HOL function that, when applied to input from a particular file, produces an output written to another file, we are able to rule out information leakage from any other file into the result.
-In short, HOL becomes a \emph{lingua franca} between kernel and untrusted code through which arbitrarily complex security and correctness policies may be expressed.
+Metadata about the behaviour of a running process could be maintained: for example, a trace of the system calls invoked by a process, thus far, and used in forming security or correctness challenges, for example by forcing untrusted code to prove that writes to a socket only ever happen after a read.
+Information flow policies could also be captured, for example, by forcing untrusted code to commit to a pure function, $f$, to be computed before reading any files, and later showing that a result corresponds to $f$ applied to the content of a particular input file, thereby independent of the content of any other file.
+In short, HOL acts as \emph{lingua franca} between kernel and untrusted code, through which arbitrarily complex policies may be expressed.
 
-\deffont{Jailing}, through which a process voluntarily sheds capabilities, is common.
-A similar pattern can be captured by allowing a change in the prevailing policy, $O$, to a new policy $N$, as long as the process is able to prove that $N$ is a refinment of $O$: $\vdash \fall{k}\fall{u}\fall{s} O\ k\ u\ s \longrightarrow N\ k\ u\ s$.
-Note that this essentially states that the machine states described by $N$ are a subset of those described by $O$.
+\deffont{Jailing}, wherein a process voluntarily sheds capabilities, is common.
+This can be captured by allowing a change in the prevailing policy, $\phi$, to a new policy $\psi$, as long as one can prove that $\psi$ is a refinment of $\phi$: $\{\} \vdash \fall{k}\fall{u}\fall{s} \phi\ k\ u\ s \longrightarrow \psi\ k\ u\ s$.
+Note that this expresses that the states described by $\psi$ are a subset of those described by $\phi$.
 
-Note that \deffont{proof carrying code} modified the operating system loader, and endowed binaries with skeleton proofs of adherence to some policy which was reconstructed by the loader, prior to execution.
-The idea sketched above represents something different, where proofs are generated dynamically, as the program executes, and is perhaps more aptly described as \deffont{proof generating code}.
+\deffont{Proof carrying code} modified the operating system loader, and endowed binaries with skeleton proofs of adherence to some policy, later reconstructed by the loader prior to execution.
+The ideas sketched above are different: proofs are generated dynamically, as the program executes, and perhaps more aptly called \deffont{proof generating code}.
 
-\paragraph{Specifications}
+\paragraph{Reinventing the proof-assistant}
+
+Switching tack, proof-assistants in the LCF lineage are also architected around a \deffont{kernel}, with ostensibly little in common with the kernels discussed so far.
+A closer examination shows that both are tasked with preserving important invariants in the face of unbounded interaction with untrusted code.
+Moreover, whilst operating system kernels enjoy a monopoly on direct hardware access, proof-assistant kernels enjoy a monopoly on the \deffont{authentication} of theorems.
+In short, both act as the roots of all trust within their domains.
+Yet, the two are implemented in different ways, with LCF kernels implemented as libraries in a \deffont{meta-language}, usually a functional language.
+Kernel interaction is via an \emph{API}, rather than an ABI.
+
+An immediate consequence of our design is that the meta-language is non-existent.
+Rather untrusted code ``driving'' the kernel can be written in any language capable of producing Wasm binary-compatible with our ABI.
+Arguably, the \emph{natural} choice for this is now not even functional, given the inherently stateful and pointer-oriented nature of the kernel.
+
+LCF-style interaction is centred around the REPL of the meta-language, or some thin layer built atop, itself started from an operating system shell---a kind of shell within a shell.
+Our design allows us to dispense with this middle-man, and develop proofs interactively directly in a shell implemented on top of our kernel.
+Moreover our kernel inherently internalises kernel objects, and could therefore make them available to browse as special files within a hierarchical filesystem, as discussed previously, in analogy to how Linux exposes its internals as special files---for example, the contents of \texttt{/proc/} and \texttt{/dev/}.
+In this way, hierarchical theories of related material could appear grouped under directories, and searched using Unix-like command-line utilities.
+
+\paragraph{Lastly, specifications}
 
@@
 Write $\mathtt{out} \mapsto \mathtt{b}$ to assert that $\mathtt{out}$ points-to the Boolean value $\mathtt{b}$.
+Write $P \star Q$ for the \deffont{separating conjunction}.
 Finally, write $\{ P \} C \{ Q \}$ to assert that $C$ executing in a state satisfying $P$ succeeds and produces a state satisfying $Q$.
@@
 \begin{gather*}
-\{ h \mapsto_{\mathtt{trm}} \mathtt{Application(l, r)} \} \\
+\{ h \mapsto_{\mathtt{trm}} \mathtt{Application(l, r)} \star \mathtt{out} \mapsto b \} \\
 \mathtt{Term.Handle.IsApplication(h, out)} \\
-\{ \mathtt{out} \mapsto \mathtt{True} \}
+\{ h \mapsto_{\mathtt{trm}} \mathtt{Application(l, r)} \star \mathtt{out} \mapsto \mathtt{True} \}
 \end{gather*}
-Note that the structural \deffont{frame rule} captures the fact that any Supervisionary system call only manipulates or queries a particular limited ``footprint'' of the kernel state.
+Note that the structural \deffont{frame rule} captures the fact that any system call only manipulates a defined footprint of kernel and untrusted code's memory.
 
-\paragraph{Reinventing the proof-assistant}
-
-Proof-assistants in the LCF lineage are also architected around a \deffont{kernel}.
-Ostensibly these have little in common with the kernels discussed so far, but a closer examination shows that both are tasked with preserving important system invariants in the face of unbounded interaction with untrusted code.
-Moreover, whilst operating system kernels enjoy a monopoly on direct access to the system, proof-assistant kernels likewise have a monopoly on the \deffont{authentication} of theorems.
-In short, both act as the roots of all trust in their respective domains.
-Yet the two types of kernel are implemented in very different ways, with LCF kernels implemented as libraries in a \deffont{meta-language}---usually some functional language, a descendent of Milner's ML.
-Untrusted code interacts with the kernel via an \emph{API}, rather than an ABI.
-
-One immediate consequence of Supervisionary's design is that the meta-language is non-existent.
-Rather, as noted above, untrusted code ``driving'' the kernel can be written in any language capable of producing Wasm that is binary-compatible with our ABI.
-Arguably, the \emph{natural} choice of language for this is now not even functional, given the inherently stateful and pointer-oriented nature of the kernel.
-
-Observe also that interaction with an LCF-style proof-assistant is centred around the meta-language's REPL, or some thin GUI layer atop, itself started from an operating system shell.
-Supervisionary's design allows us to potentially dispense with this middle-man, and develop proofs interactively directly in a shell implemented on top of our kernel.
-Moreover, and in contrast to LCF, our kernel inherently internalises kernel objects and could make them available to browse as special files within a hierarchical filesystem, made available to untrusted code as discussed previously, in analogy to how Linux exposes its internals as special files (e.g.
-the \texttt{/proc/} file system).
-In this way, hierarchical theories of related material appear grouped under directories, and can be searched using Unix-like command-line utilities.
 \end{document}

### 2021-10-08 paper/prisc-2022.tex
prisc-2022: some minor word rearranging, and squashing

@@
 The kernel also offers an abstract machine interface to user-space software, smoothing over the vulgarities of programming against any particular peripheral, say.
-Useful abstractions, such as sockets and a file-system, are also commonly offered.
+Useful abstractions, such as sockets and a filesystem, are also commonly offered.
 This interface is largely exposed as \deffont{system calls}, defining an \deffont{application binary interface}, or ABI henceforth, which user-space and kernel must understand to interact.
@@
 \end{displaymath}
-Here, $\mathtt{pre}$ points-to a theorem $\Gamma \vdash r = s$ and after succeeding, passing obvious checks, $\mathtt{out}$ contains a handle that points-to a newly-allocated theorem $\Gamma \vdash s = r$.
+Here, $\mathtt{pre}$ points-to an existing theorem object $\Gamma \vdash r = s$ and after succeeding, passing obvious checks, $\mathtt{out}$ contains a handle that points-to a newly-allocated theorem $\Gamma \vdash s = r$.
 
@@
 \deffont{Jailing}, wherein a process voluntarily sheds capabilities, is common.
-This can be captured by allowing a change in the prevailing policy, $\phi$, to a new policy $\psi$, as long as one can prove that $\psi$ is a refinment of $\phi$: $\{\} \vdash \fall{k}\fall{u}\fall{s} \phi\ k\ u\ s \longrightarrow \psi\ k\ u\ s$.
+This can be captured by allowing a change in the prevailing policy, $\phi$, to a new policy $\psi$, as long as one can prove that $\psi$ is a refinment of $\phi$: $\{\} \vdash \fall{k}\fall{u}\fall{s} \psi\ k\ u\ s \longrightarrow \phi\ k\ u\ s$.
 Note that this expresses that the states described by $\psi$ are a subset of those described by $\phi$.
@@
 \deffont{Proof carrying code} modified the operating system loader, and endowed binaries with skeleton proofs of adherence to some policy, later reconstructed by the loader prior to execution.
-The ideas sketched above are different: proofs are generated dynamically, as the program executes, and perhaps more aptly called \deffont{proof generating code}.
+The ideas sketched above are different: proofs are generated dynamically, as the program executes, and more aptly called \deffont{proof generating code}.
+Moreover, theorems could be generated statically, offline, and used dynamically:
 
@@
 
-Switching tack, proof-assistants in the LCF lineage are also architected around a \deffont{kernel}, with ostensibly little in common with the kernels discussed so far.
-A closer examination shows that both are tasked with preserving important invariants in the face of unbounded interaction with untrusted code.
-Moreover, whilst operating system kernels enjoy a monopoly on direct hardware access, proof-assistant kernels enjoy a monopoly on the \deffont{authentication} of theorems.
-In short, both act as the roots of all trust within their domains.
+Proof-assistants in the LCF lineage are also architected around a \deffont{kernel}, with ostensibly little in common with operating system kernels.
+Examining, one sees that both are tasked with preserving invariants in the face of unbounded interaction with untrusted code.
+Moreover, whilst operating system kernels enjoy a monopoly on hardware access, proof-assistant kernels enjoy a monopoly on theorem \deffont{authentication}.
+In short, both act as roots of all trust within their domains.
 Yet, the two are implemented in different ways, with LCF kernels implemented as libraries in a \deffont{meta-language}, usually a functional language.
@@
 Finally, write $\{ P \} C \{ Q \}$ to assert that $C$ executing in a state satisfying $P$ succeeds and produces a state satisfying $Q$.
-We then have, for example:
+We then have the following example \deffont{local axiom}:
 \begin{gather*}

### 2021-10-08 paper/prisc-2022.tex
prisc-2022: some minor word rearranging, and squashing

@@
 \title{The Supervisionary proof-checking kernel}
-\subtitle{Or: toward proof-generating code}
+\subtitle{Or: a work-in-progress toward proof-generating code}
 
@@
 These theorems could be correctness or security-related theorems, corresponding to a prevailing \deffont{policy} in force.
-Moreover, as Supervisionary is capable of capturing the runtime state of untrusted code, these challenges can be HOL predicates---functions of the reified runtime state of untrusted code, the kernel, and the arguments, and name of, the invoked system call.
+Moreover, as Supervisionary is capable of capturing the runtime state of untrusted code, these challenges can be HOL predicates that are functions of the reified runtime state of untrusted code, the kernel, and the arguments, and name of, the invoked system call.
 
@@
 Moreover our kernel inherently internalises kernel objects, and could therefore make them available to browse as special files within a hierarchical filesystem, as discussed previously, in analogy to how Linux exposes its internals as special files---for example, the contents of \texttt{/proc/} and \texttt{/dev/}.
-In this way, hierarchical theories of related material could appear grouped under directories, and searched using Unix-like command-line utilities.
+In this way, related material could appear grouped in directories, searched using command-line utilities, or opened by a process for use at runtime.
 

### 2021-10-08 paper/prisc-2022.tex
prisc-2022: moved proof-asssitant section up

@@
 We have described the \emph{what} of Supervisionary and now turn our attention to the \emph{why}.
-This can be approached by examining Supervisionary both as a programming platform, and as a proof-assistant kernel.
+This can be approached by examining Supervisionary both as a proof-assistant kernel, and as a programming platform.
 What follows is largely speculative, and a work-in-progress.
+
+\paragraph{Reinventing the proof-assistant}
+
+Proof-assistants in the LCF lineage are also architected around a \deffont{kernel}, with ostensibly little in common with operating system kernels.
+Examining, one sees that both are tasked with preserving invariants in the face of unbounded interaction with untrusted code.
+Moreover, whilst operating system kernels enjoy a monopoly on hardware access, proof-assistant kernels enjoy a monopoly on theorem \deffont{authentication}.
+In short, both act as roots of all trust within their domains.
+Yet, the two are implemented in different ways, with LCF kernels implemented as libraries in a \deffont{meta-language}, usually a functional language.
+Kernel interaction is via an \emph{API}, rather than an ABI.
+
+An immediate consequence of our design is that the meta-language is non-existent.
+Rather untrusted code ``driving'' the kernel can be written in any language capable of producing Wasm binary-compatible with our ABI.
+Arguably, the \emph{natural} choice for this is now not even functional, given the inherently stateful and pointer-oriented nature of the kernel.
+
+LCF-style interaction is centred around the REPL of the meta-language, or some thin layer built atop, itself started from an operating system shell---a kind of shell within a shell.
+Our design allows us to dispense with this middle-man, and develop proofs interactively directly in a shell implemented on top of our kernel.
+Moreover our kernel inherently internalises kernel objects, and could therefore make them available to browse as special files within a hierarchical filesystem, as discussed below, in analogy to how Linux exposes its internals as special files---for example, the contents of \texttt{/proc/} and \texttt{/dev/}.
+In this way, related material could appear grouped in directories, searched using command-line utilities, or opened by a process for use at runtime.
 
@@
 Moreover, theorems could be generated statically, offline, and used dynamically:
-
-\paragraph{Reinventing the proof-assistant}
-
-Proof-assistants in the LCF lineage are also architected around a \deffont{kernel}, with ostensibly little in common with operating system kernels.
-Examining, one sees that both are tasked with preserving invariants in the face of unbounded interaction with untrusted code.
-Moreover, whilst operating system kernels enjoy a monopoly on hardware access, proof-assistant kernels enjoy a monopoly on theorem \deffont{authentication}.
-In short, both act as roots of all trust within their domains.
-Yet, the two are implemented in different ways, with LCF kernels implemented as libraries in a \deffont{meta-language}, usually a functional language.
-Kernel interaction is via an \emph{API}, rather than an ABI.
-
-An immediate consequence of our design is that the meta-language is non-existent.
-Rather untrusted code ``driving'' the kernel can be written in any language capable of producing Wasm binary-compatible with our ABI.
-Arguably, the \emph{natural} choice for this is now not even functional, given the inherently stateful and pointer-oriented nature of the kernel.
-
-LCF-style interaction is centred around the REPL of the meta-language, or some thin layer built atop, itself started from an operating system shell---a kind of shell within a shell.
-Our design allows us to dispense with this middle-man, and develop proofs interactively directly in a shell implemented on top of our kernel.
-Moreover our kernel inherently internalises kernel objects, and could therefore make them available to browse as special files within a hierarchical filesystem, as discussed previously, in analogy to how Linux exposes its internals as special files---for example, the contents of \texttt{/proc/} and \texttt{/dev/}.
-In this way, related material could appear grouped in directories, searched using command-line utilities, or opened by a process for use at runtime.
 

### 2021-10-11 paper/prisc-2022.tex
prisc-2022: fix some minor typos

@@
 Hypervisors, the Secure World of Arm TrustZone, and virtual machines (in the Wasm sense) are structured in this way.
-Call software designed in this pattern \deffont{supervisory software}.
+We call software designed in this pattern \deffont{supervisory software}.
 
@@
 Nothing in principle stops us from extending Supervisionary's ABI with facilities for sockets, querying a hierarchical filesystem, and other similar system tasks.
-In particular, given our use of Wasm, we could implement the Wasm System Interface---a POSIX-like interface for Wasm---transforming Supervisionary from a programmable proof-checker into a general-purpose virtual machine, capable of supporting a wide-range of programs.
+In particular, given our use of Wasm, we could implement the Wasm System Interface---a POSIX-like interface for Wasm---transforming Supervisionary from a programmable proof-checker into a general-purpose virtual machine, capable of supporting a wide range of programs.
 

### 2021-10-11 paper/prisc-2022.tex
prisc-2022: some minor rewording

@@
 
-\deffont{Jailing}, wherein a process voluntarily sheds capabilities, is common.
-This can be captured by allowing a change in the prevailing policy, $\phi$, to a new policy $\psi$, as long as one can prove that $\psi$ is a refinment of $\phi$: $\{\} \vdash \fall{k}\fall{u}\fall{s} \psi\ k\ u\ s \longrightarrow \phi\ k\ u\ s$.
+\deffont{Jailing}, wherein a process voluntarily sheds capabilities, is common in existing operating systems.
+This can be captured in Supervisionary by allowing a process to replace the prevailing policy, $\phi$, with a new policy $\psi$, after proving that $\psi$ is a refinment of $\phi$: $\{\} \vdash \fall{k}\fall{u}\fall{s} \psi\ k\ u\ s \longrightarrow \phi\ k\ u\ s$.
 Note that this expresses that the states described by $\psi$ are a subset of those described by $\phi$.

### 2021-10-23 paper/prisc-2022.tex
prisc-2022: from the ashes a fire shall be woken

@@
 \maketitle
+
+\paragraph{Some scene setting}
+
+Interactive theorem proving software is typically designed around a proof-checking \emph{kernel}.
+Proofs are constructed by submitting theorem derivations to the kernel for authentication, and then composing them with other authentic theorems.
+A proof may be large and complex, as may be the software components, both interactive and automated, used to construct it.
+However, the small and simple kernel is soley responsible for validating it.
+This implementation strategy---known as the ``LCF-style'', after Milner's system that first implemented it---is a reliability mechanism, aiming to ensure that any purported theorem produced by the system is indeed a theorem.
+
+In an operating system, the \emph{kernel} is the most privileged component, responsible for mediating interaction by less privileged components with the hardware and with each other.
+An unpriviledged process interacts with the rest of the system by issuing \emph{system calls} accross a hardware privilege boundary to the operating system kernel, which may contain requests for hardware resources, accesses to shared system services such as the filesystem, or messages addressed to other processes.
+In this way, the operating system kernel \emph{supervises} less priviledged processes.
+
+In a typical proof assistant, the boundary between a proof-checking kernel and the rest of the system is a programming-language level module boundary.
+In this work, we explore the implications of a proof-checking kernel which acts as a supervisor to the rest of theorem proving system, across a more concrete privilege boundary.
+
+\paragraph{The Supervisionary system interface}
+
+\emph{Supervisionary} is the name of our prototype implementation of a proof-checking system for Gordon's HOL structured as supervisory software, for expressing and experimenting with the ideas presented in this paper.
+In particular, our prototype specifies a \emph{system interface} to the proof-checking kernel which serve as the substance of communication across any of the concrete privilege boundries described in the body of this paper.
+
+% this bit is copied mostly verbatim from the previous version
+
+The kernel manages various private heaps within which \emph{kernel objects} are allocated, corresponding to the paraphenalia of any HOL implementation---type-formers, types, term constants, terms, and theorems---and never directly exposed to untrusted code.
+Kernel objects are allocated in response to system calls like:
+\begin{displaymath}
+\mathtt{Term.Handle.AllocateApplication(left, right, out)}
+\end{displaymath}
+Here, both $\mathtt{left}$ and $\mathtt{right}$ are \emph{kernel handles}, assumed to point-to allocated terms, whilst $\mathtt{out}$ points-to a buffer in untrusted code's memory.
+If neither $\mathtt{left}$ nor $\mathtt{right}$ \emph{dangle}, and the types of their referents match, a fresh handle is generated which points-to a new HOL term application object, with internal pointers to the functional- and argument-terms.
+This handle is returned to the caller via the $\mathtt{out}$ pointer.
+
+The manipulation and querying of kernel objects is performed defensively by the kernel itself on behalf of untrusted code.
+The kernel is careful to maintain invariants such as the \emph{inductivity} of its heaps, with nodes in the kernel object graph pointing-to allocated objects at all times.
+
+Space constraints prevent us from describing the entire Supervisionary ABI for working with, and on, kernel objects.
+However, note that theorems are also constructed in a similar way to terms, inductively building derivation trees.
+For example, the HOL symmetry rule is exposed as:
+\begin{displaymath}
+\mathtt{Theorem.Handle.AllocateSym(pre, out)}
+\end{displaymath}
+Here, $\mathtt{pre}$ points-to an existing theorem object $\Gamma \vdash r = s$ and after succeeding, passing obvious checks, $\mathtt{out}$ contains a handle that points-to a newly-allocated theorem $\Gamma \vdash s = r$.
+
+\paragraph{More flexible theorem proving tools}
+
+Immediately, we can re-imagine the implementation of the foundations of familiar theorem proving tools.
+For this case, the unpriviledged bulk of the theorem proving system runs inside of a WebAssembly (Wasm) sandbox, supervised by the proof-checking kernel which runs as native code in the process hosting the sandbox.
+In the case of a typical LCF-style proof assistant such as HOL4, the entire system could be compiled to Wasm, and the module implementing the kernel could be modified to act as a proxy in front of the Supervisionary kernel outside of the sandbox.
+This modification has immediate implications for the reliability and auditability of the system, because the only critical component now isolated from the bulk of the system by a priviledged boundary.
+
+% versus just type system
+% even memory corruption vulns in bulk aren't a problem
+% next step: polyglot theorem proving software
+% another benefit: strong isolation boundaries within your proof: isolate important spec from the large proof
+
+\paragraph{Generalized runtime verification}
+
+% I can draft this section on Saturday.
+
+\paragraph{More}
+
+TODO
+
+% VM/PA
+
+% more about our prototype?
+
+\iffalse
 
@@
 
+\fi
+
 \end{document}

### 2021-10-24 paper/prisc-2022.tex
prisc-2022: large changes to draft paper
- Tightened up introductory paragraphs,
- Added in section on Separation Logic, carried over from previous draft,
- Added back in section on runtime monitoring,
- Made some changes to our discussion of proof-carrying code.

--- 2021-10-11 paper/prisc-2022.tex
@@
 \maketitle
+
+\paragraph{Some scene setting}
+
+Interactive theorem proving software is typically designed around a trusted proof-checking \emph{kernel}.
+The kernel is the sole system component capable of authenticating theorems.
+Untrusted automation procedures reside outside of the kernel, and drive it to deduce new theorems via a defined API.
+This strategy---first introduced by Milner in his LCF proof assistant---is a reliability mechanism, aiming to ensure that any purported theorem produced by the system is indeed a theorem of the logic.
+
+Switching focus, operating systems are also typically designed around a trusted \emph{kernel}, a privileged component responsible for---amongst other things---mediating interaction betwixt user-space software and hardware.
+Untrusted processes interact with the system by issuing kernel \emph{system calls} across a hardware privilege boundary.
+In this way, the operating system kernel \emph{supervises} user-space processes.
+
+In this abstract, we explore designing proof-checking kernels as \emph{supervisory software}, where separation between kernel and untrusted code is enforced by \emph{privilege}, not programming language type abstraction and modularity.
+
+\paragraph{The Supervisionary system interface}
+
+\emph{Supervisionary} is a proof-checking system for Gordon's HOL, structured as supervisory software.
+Supervisionary provides a \emph{system interface} to untrusted code similar to an operating system's system call interface.
+We have implemented Supervisionary as a Wasm host, which allows us to rapidly explore the most important ideas.
+Moreover, we use Rust as the system's implementation language---an unusual choice, albeit one that entails no risk of system unsoundness, providing the kernel's system call interface is carefully designed.
+
+In particular, the kernel manages various private heaps within which \emph{kernel objects} are allocated, corresponding to the paraphenalia of any HOL implementation---type-formers, types, term constants, terms, and theorems---and never directly exposed to untrusted code.
+Kernel objects are allocated in response to system calls like:
+\begin{displaymath}
+\mathtt{Term.Handle.AllocateApplication(left, right, out)}
+\end{displaymath}
+Here, both $\mathtt{left}$ and $\mathtt{right}$ are \emph{kernel handles}, assumed to point-to allocated terms, whilst $\mathtt{out}$ points-to a buffer in untrusted code's memory.
+If neither $\mathtt{left}$ nor $\mathtt{right}$ \emph{dangle}, and the types of their referents match, a fresh handle is generated which points-to a new HOL term application object, with internal pointers to the functional- and argument-terms.
+This handle is returned to the caller via the $\mathtt{out}$ pointer.
+
+The manipulation and querying of kernel objects is performed defensively by the kernel itself on behalf of untrusted code.
+The kernel is careful to maintain invariants such as the \emph{inductivity} of its heaps, with nodes in the kernel object graph pointing-to allocated objects at all times.
+
+Space constraints prevent us from describing the entire Supervisionary system interface for working with, and on, kernel objects.
+However, note that theorems are also constructed in a similar way to terms, inductively building derivation trees.
+For example, the HOL symmetry rule is exposed as:
+\begin{displaymath}
+\mathtt{Theorem.Handle.AllocateSym(pre, out)}
+\end{displaymath}
+Here, $\mathtt{pre}$ points-to an existing theorem object $\Gamma \vdash r = s$ and after succeeding, passing obvious checks, $\mathtt{out}$ contains a handle that points-to a newly-allocated theorem $\Gamma \vdash s = r$.
+
+Note that one consequence of this style of implementation is the ability to provide concise \emph{specifications} of Supervisionary ABI functions.
+Essentially, the Supervisionary kernel is a grand exercise in pointer manipulation, and as such our ABI specifications can be expressed as Hoare Triples, using Separation Logic as our assertion language.
+Writing $\mathtt{h} \mapsto_{\mathtt{trm}} \mathtt{Application(l, r)}$ to assert that the handle $\mathtt{h}$ points-to a term application (of the term pointed-to by $l$ to the term pointed-to by $r$), and writing $\mathtt{out} \mapsto \mathtt{b}$ to assert that $\mathtt{out}$ points-to the Boolean value $\mathtt{b}$, we have, for example:
+\begin{gather*}
+\{ h \mapsto_{\mathtt{trm}} \mathtt{Application(l, r)} \} \\
+\mathtt{Term.Handle.IsApplication(h, out)} \\
+\{ \mathtt{out} \mapsto \mathtt{True} \}
+\end{gather*}
+(Here, the triple $\{ P \} C \{ Q \}$ asserts that if the command $C$ executes in a state concordant with $P$ then the command succeeds and produces a state concordant with $Q$.)
+
+We now turn our attention to describing some potential uses of the Supervisionary kernel.
+Exploring what follows is still a work-in-progress.
+
+\paragraph{More flexible theorem proving tools}
+
+Immediately, we can re-imagine the implementation of the foundations of familiar theorem proving tools.
+For this case, the unpriviledged bulk of the theorem proving system runs inside of a WebAssembly (Wasm) sandbox, supervised by the proof-checking kernel which runs as native code in the process hosting the sandbox.
+In the case of a typical LCF-style proof assistant such as HOL4, the entire system could be compiled to Wasm, and the module implementing the kernel could be modified to act as a proxy in front of the Supervisionary kernel outside of the sandbox.
+This modification has immediate implications for the reliability and auditability of the system, because the only critical component now isolated from the bulk of the system by a priviledged boundary.
+
+% versus just type system
+% even memory corruption vulns in bulk aren't a problem
+% next step: polyglot theorem proving software
+% another benefit: strong isolation boundaries within your proof: isolate important spec from the large proof
+
+\paragraph{Runtime verification}
+
+Nothing in principle stops us from extending Supervisionary's system interface with facilities for sockets, querying a hierarchical filesystem, and other similar tasks tied to abstractions typically provided by an operating system.
+In particular, given our use of Wasm, we could implement the Wasm System Interface---a POSIX-like interface for Wasm---transforming Supervisionary from a programmable proof-checker into a general-purpose virtual machine, capable of supporting a wide range of programs.
+
+\emph{What happens if we blur the lines between Supervisionary's interfaces for system access and proof-checking?}
+
+Untrusted code executing under Supervisionary's supervision could be challenged to prove some theorem each time it wished to open a file on the filesystem, or otherwise perform some side-effect.
+These theorems could be correctness or security-related theorems, corresponding to a prevailing \deffont{policy} in force.
+Moreover, as Supervisionary is capable of capturing the runtime state of untrusted code, these challenges can be HOL predicates that are functions of the reified runtime state of untrusted code, the kernel, and the arguments, and name of, the invoked system call.
+
+Two predicates of interest are $\lam{k}\lam{u}\lam{s}\top$ and $\lam{k}\lam{u}\lam{s}\bot$.
+Here, $\top$ and $\bot$ are truth and falsity, and $k$, $u$, and $s$ the kernel and untrusted code states, and packed system call metadata, reified as HOL data, at the point of invocation of the system call.
+One can always prove $\{\} \vdash \top$ and therefore $\lam{k}\lam{u}\lam{s}\top$ represents no restriction, whereas $\{ \} \vdash \bot$ is never provable, absent axioms, with $\lam{k}\lam{u}\lam{s}\bot$ representing a ``closing off'' of a system call.
+By making the challenge a function of the system call name and arguments, this closing off can be specific, banning a process from calling a particular system call, or a system call with a particular set of arguments, allowing us to mimic mechanisms like \texttt{seccomp} from Linux.
+
+We can go further: metadata about the behaviour of a running process could be maintained---for example, a record of the system calls invoked by a process, thus far.
+This record could be used in forming security or correctness challenges, for example by forcing untrusted code to prove that writes to a socket only ever happen after a read, or reads and writes on sockets satisfy some protocol.
+In short, HOL acts as \emph{lingua franca} between kernel and untrusted code, through which arbitrarily complex policies may be expressed.
+
+\deffont{Jailing}, wherein a process voluntarily sheds capabilities, is common in existing operating systems.
+This can be captured in Supervisionary by allowing a process to dynamically replace the prevailing policy, $\phi$, with a new policy $\psi$, after proving that $\psi$ is a refinement of $\phi$: $\{\} \vdash \fall{k}\fall{u}\fall{s} \psi\ k\ u\ s \longrightarrow \phi\ k\ u\ s$.
+Note that this expresses that the states described by $\psi$ are a subset of those described by $\phi$.
+
+Note that this idea shares some similarities with \deffont{proof carrying code}, wherein binaries are accompanied with (skeleton) proofs of their adherence to some policy, and these proofs checked by the operating system prior to execution.
+However, the ideas sketched above generalise this: proofs can be generated dynamically, as the program executes, and could more aptly be called \deffont{proof generating code}.
+However, this potential use of Supervisionary also blurs the boundary between static and runtime verification, and is capable of expressing use-cases typical of proof-carrying code.
+For example, the operational semantics and instruction decoding functionality of Wasm is clearly embeddable in HOL.
+Using this, properties of a program $P$ to be executed under Supervisionary's supervision could be established statically, and registered with the kernel, perhaps interactively by a user or by another program executing before $P$ executes.
+This theorem can then be used by $P$ in closing challenges from Supervisionary, issued at runtime.
+
+\end{document}
+
+\iffalse
 

### 2021-10-25 paper/prisc-2022.tex
prisc-2022: improve theorem proving tools section
- Rename from "More flexible theorem proving tools" to "Robust and flexible theorem proving tools"
- Improve language of existing content
- Add paragraph on possibilites enabled by isolation boundaries within the rich portion of a theorem proving system

@@
 Supervisionary provides a \emph{system interface} to untrusted code similar to an operating system's system call interface.
-We have implemented Supervisionary as a Wasm host, which allows us to rapidly explore the most important ideas.
+We have implemented Supervisionary as a WebAssembly (Wasm henceforth) host, which allows us to rapidly explore the most important ideas.
 Moreover, we use Rust as the system's implementation language---an unusual choice, albeit one that entails no risk of system unsoundness, providing the kernel's system call interface is carefully designed.
@@
 
-\paragraph{More flexible theorem proving tools}
+\paragraph{Robust and flexible theorem proving tools}
 
-Immediately, we can re-imagine the implementation of the foundations of familiar theorem proving tools.
-For this case, the unpriviledged bulk of the theorem proving system runs inside of a WebAssembly (Wasm) sandbox, supervised by the proof-checking kernel which runs as native code in the process hosting the sandbox.
-In the case of a typical LCF-style proof assistant such as HOL4, the entire system could be compiled to Wasm, and the module implementing the kernel could be modified to act as a proxy in front of the Supervisionary kernel outside of the sandbox.
-This modification has immediate implications for the reliability and auditability of the system, because the only critical component now isolated from the bulk of the system by a priviledged boundary.
+Immediately, we can improve the robustness and flexibility of familiar theorem proving software by restructuring it around the Supervisionary kernel.
+For example, we can refactor a typical LCF-style proof assistant so that the source-level module implementing its proof-checking kernel acts as frontend for the Supervisionary kernel, proxying derivations across an isolation boundary such as a process boundary or Wasm sandbox boundary.
+Before, we relied on the ML type system to protect the soundness of our system from errors outside of the kernel.
+Now, we rely on a stronger system-level boundary, meaning that even memory corruption in the rich portion of our system does not pose a threat to its soundness.
 
-% versus just type system
-% even memory corruption vulns in bulk aren't a problem
-% next step: polyglot theorem proving software
-% another benefit: strong isolation boundaries within your proof: isolate important spec from the large proof
+Stepping further, we can design less familiar theorem proving systems where the rich portion itself contains internal system-level isolation boundaries.
+For example, under a single instance of the Supervisionary kernel, a group of distinct theorem proving clients could construct a proof in concert.
+Furthermore, in certain cases, isolating specification from proof might allow for more confidence in the correctness of a specification's expression.
+
+% note: explicitly mention polyglot theorem proving software?
 

### 2021-10-25 paper/prisc-2022.tex
prisc-2022: more changes

--- 2021-10-24 paper/prisc-2022.tex
@@
 
-Interactive theorem proving software is typically designed around a trusted proof-checking \emph{kernel}.
-The kernel is the sole system component capable of authenticating theorems.
-Untrusted automation procedures reside outside of the kernel, and drive it to deduce new theorems via a defined API.
-This strategy---first introduced by Milner in his LCF proof assistant---is a reliability mechanism, aiming to ensure that any purported theorem produced by the system is indeed a theorem of the logic.
+Interactive theorem proving software is typically designed around a trusted proof-checking \emph{kernel}, the sole system component capable of authenticating theorems.
+Untrusted automation procedures reside outside of the kernel, and drives it to deduce new theorems via an API.
+Kernel and untrusted automation are typically implemented in the same programming language---the ``meta-language''---usually some functional programming language in the ML family.
+This strategy---introduced by Milner in his LCF proof assistant---is a reliability mechanism, aiming to ensure that any purported theorem produced by the system is indeed a legitimate entailment from the theory within the logic.
 
-Switching focus, operating systems are also typically designed around a trusted \emph{kernel}, a privileged component responsible for---amongst other things---mediating interaction betwixt user-space software and hardware.
+Switching tack, operating systems are also typically designed around a trusted \emph{kernel}, a privileged component responsible for---amongst other things---mediating interaction betwixt user-space software and hardware.
 Untrusted processes interact with the system by issuing kernel \emph{system calls} across a hardware privilege boundary.
 In this way, the operating system kernel \emph{supervises} user-space processes.
+Usually, the source executables of untrusted processes need not have been written in the same language as the kernel.
 
-In this abstract, we explore designing proof-checking kernels as \emph{supervisory software}, where separation between kernel and untrusted code is enforced by \emph{privilege}, not programming language type abstraction and modularity.
+Though ostensibly very different, squinting, we see the two kinds of kernel are tasked with solving the same task: enforcing system invariants in the face of interaction with untrusted code.
+Yet, the two solutions to solving this problem, employed by the respective kinds of kernel, are very different.
+In this abstract, we explore designing proof-checking kernels as \emph{supervisory software}, where separation between kernel and untrusted code is enforced by \emph{privilege}.
 
-\paragraph{The Supervisionary system interface}
+\paragraph{System interface}
 
-\emph{Supervisionary} is a proof-checking system for Gordon's HOL, structured as supervisory software.
-Supervisionary provides a \emph{system interface} to untrusted code similar to an operating system's system call interface.
-We have implemented Supervisionary as a Wasm host, which allows us to rapidly explore the most important ideas.
-Moreover, we use Rust as the system's implementation language---an unusual choice, albeit one that entails no risk of system unsoundness, providing the kernel's system call interface is carefully designed.
+\emph{Supervisionary} is a proof-checking system for Gordon's HOL, structured as supervisory software, and provides a \emph{system interface} to untrusted code similar to an operating system's system call interface.
+Supervisionary is implemented as a WebAssembly (Wasm henceforth) host, allowing us to rapidly prototype.
 
+We use Rust as our implementation language, rather than a functional programming language.
+This entails no risk of unsoundness providing our system interface is carefully designed.
 In particular, the kernel manages various private heaps within which \emph{kernel objects} are allocated, corresponding to the paraphenalia of any HOL implementation---type-formers, types, term constants, terms, and theorems---and never directly exposed to untrusted code.
@@
 
-Note that one consequence of this style of implementation is the ability to provide concise \emph{specifications} of Supervisionary ABI functions.
+Note that one interesting consequence of this style of implementation is the ability to provide concise \emph{specifications} of Supervisionary's system interface functions.
 Essentially, the Supervisionary kernel is a grand exercise in pointer manipulation, and as such our ABI specifications can be expressed as Hoare Triples, using Separation Logic as our assertion language.
-Writing $\mathtt{h} \mapsto_{\mathtt{trm}} \mathtt{Application(l, r)}$ to assert that the handle $\mathtt{h}$ points-to a term application (of the term pointed-to by $l$ to the term pointed-to by $r$), and writing $\mathtt{out} \mapsto \mathtt{b}$ to assert that $\mathtt{out}$ points-to the Boolean value $\mathtt{b}$, we have, for example:
+Writing $\mathtt{h} \mapsto_{\mathtt{trm}} \mathtt{Application(l, r)}$ to assert that the handle $\mathtt{h}$ points-to a term application (of the term pointed-to by $l$ to the term pointed-to by $r$), and writing $\mathtt{out} \mapsto \mathtt{b}$ to assert that $\mathtt{out}$ points-to the Boolean value $\mathtt{b}$, we have:
 \begin{gather*}
@@
 
-We now turn our attention to describing some potential uses of the Supervisionary kernel.
+Note that Supervisionary lacks any analogue of the traditional LCF meta-language.
+Kernel and automation are now decoupled, and code written in \emph{any} language can ``drive'' the kernel, providing it produces binary-compatible executables.
+
+We now turn to speculation around potential uses of Supervisionary.
 Exploring what follows is still a work-in-progress.
 
-\paragraph{More flexible theorem proving tools}
-
-Immediately, we can re-imagine the implementation of the foundations of familiar theorem proving tools.
-For this case, the unpriviledged bulk of the theorem proving system runs inside of a WebAssembly (Wasm) sandbox, supervised by the proof-checking kernel which runs as native code in the process hosting the sandbox.
-In the case of a typical LCF-style proof assistant such as HOL4, the entire system could be compiled to Wasm, and the module implementing the kernel could be modified to act as a proxy in front of the Supervisionary kernel outside of the sandbox.
-This modification has immediate implications for the reliability and auditability of the system, because the only critical component now isolated from the bulk of the system by a priviledged boundary.
-
-% versus just type system
-% even memory corruption vulns in bulk aren't a problem
-% next step: polyglot theorem proving software
-% another benefit: strong isolation boundaries within your proof: isolate important spec from the large proof
+% note: explicitly mention polyglot theorem proving software?
 
@@
 
-Nothing in principle stops us from extending Supervisionary's system interface with facilities for sockets, querying a hierarchical filesystem, and other similar tasks tied to abstractions typically provided by an operating system.
-In particular, given our use of Wasm, we could implement the Wasm System Interface---a POSIX-like interface for Wasm---transforming Supervisionary from a programmable proof-checker into a general-purpose virtual machine, capable of supporting a wide range of programs.
+Given our use of Wasm, we could extend our system interface by also implementing the Wasm System Interface---a POSIX-like interface for Wasm.
+This would transform Supervisionary from a mere programmable proof-checker into a general-purpose sandbox, capable of executing arbitrary programs, albeit with an unusual interface for constructing proofs.
 
-\emph{What happens if we blur the lines between Supervisionary's interfaces for system access and proof-checking?}
+But: \emph{what happens if we blur the lines between Supervisionary's interfaces for system access and proof-checking?}
 
@@
 However, the ideas sketched above generalise this: proofs can be generated dynamically, as the program executes, and could more aptly be called \deffont{proof generating code}.
-However, this potential use of Supervisionary also blurs the boundary between static and runtime verification, and is capable of expressing use-cases typical of proof-carrying code.
+
+\paragraph{Restructuring proof-checking tools}
+
+Existing tools in the HOL familiy could be refactored around Supervisionary by changing their kernels to act as a shim between untrusted automation and the Supervisionary kernel.
+Arguably, this increases the robustness of existing tools, enforcing separation by isolation, rather than module boundaries.
+More interestingly, this also means that Supervisionary acts as a mechanism for ``transporting'' definitions and theorems between systems within the wider HOL family: systems are caapble of referring to, and manipulating, kernel objects produced by other systems through Supervisionary kernel handles.
+With this, we can also bootstrap a full theorem-proving environment, with associated libraries of content, on top of Supervisionary without writing it from scratch.
+
+Our prototype acts as both programmable proof-checker and general-purpose sandbox.
+The previous section highlighted that this idea can be used to enforce runtime properties of programs.
+But, we can also use this blurring to import ideas from the operating systems community into the design of proof-checking software itself.
+For example, libraries of mathematical theorems and definitions could be presented to Supervisionary's users via a hierarchical or tag-based file-system, and explored with command-line tools in an interactive shell atop the Supervisionary kernel.
+
+\paragraph{Lastly}
+
+Supervisionary's dual interpretation as sandbox and proof-checker blurs the boundary between static and runtime verification, and between proof-checker and sandbox.
+Whilst proofs can be generated by executables themselves at runtime, they could also be generated interactively by users, prior to the execution of a program, or even generated for use by a program by other programs.
+
 For example, the operational semantics and instruction decoding functionality of Wasm is clearly embeddable in HOL.
-Using this, properties of a program $P$ to be executed under Supervisionary's supervision could be established statically, and registered with the kernel, perhaps interactively by a user or by another program executing before $P$ executes.
-This theorem can then be used by $P$ in closing challenges from Supervisionary, issued at runtime.
+Using this, properties of a program $P$ to be executed under Supervisionary could be established statically, and registered with the kernel, perhaps interactively by a user or by another program executing before $P$ executes.
+This theorem can then be used by $P$ in closing challenges from Supervisionary, issued at runtime, tieing the correctness of $P$ to its execution.
 

### 2021-10-25 paper/prisc-2022.tex
prisc-2022: minor grammatical changes

@@
 
-Though ostensibly very different, squinting, we see the two kinds of kernel are tasked with solving the same task: enforcing system invariants in the face of interaction with untrusted code.
+Though ostensibly very different, squinting, we see that the two kinds of kernel are tasked with solving the same task: enforcing system invariants in the face of interaction with untrusted code.
 Yet, the two solutions to solving this problem, employed by the respective kinds of kernel, are very different.
@@
 \emph{Supervisionary} is a proof-checking system for Gordon's HOL, structured as supervisory software, and provides a \emph{system interface} to untrusted code similar to an operating system's system call interface.
-Supervisionary is implemented as a WebAssembly (Wasm henceforth) host, allowing us to rapidly prototype.
+Supervisionary is implemented as a WebAssembly (Wasm henceforth) host, allowing us to prototype rapidly.
 
@@
 Arguably, this increases the robustness of existing tools, enforcing separation by isolation, rather than module boundaries.
-More interestingly, this also means that Supervisionary acts as a mechanism for ``transporting'' definitions and theorems between systems within the wider HOL family: systems are caapble of referring to, and manipulating, kernel objects produced by other systems through Supervisionary kernel handles.
+More interestingly, this also means that Supervisionary acts as a mechanism for ``transporting'' definitions and theorems between systems within the wider HOL family: systems are capable of referring to, and manipulating, kernel objects produced by other systems through Supervisionary kernel handles.
 With this, we can also bootstrap a full theorem-proving environment, with associated libraries of content, on top of Supervisionary without writing it from scratch.
@@
 Using this, properties of a program $P$ to be executed under Supervisionary could be established statically, and registered with the kernel, perhaps interactively by a user or by another program executing before $P$ executes.
-This theorem can then be used by $P$ in closing challenges from Supervisionary, issued at runtime, tieing the correctness of $P$ to its execution.
+This theorem can then be used by $P$ in closing challenges from Supervisionary, issued at runtime, tying the correctness of $P$ to its execution.
 

### 2021-10-25 paper/prisc-2022.tex
prisc-2020: minor rewording according to slack discussion

@@
 
-Switching tack, operating systems are also typically designed around a trusted \emph{kernel}, a privileged component responsible for---amongst other things---mediating interaction betwixt user-space software and hardware.
+Changing tack, operating systems are also typically designed around a trusted \emph{kernel}, a privileged component responsible for---amongst other things---mediating interaction betwixt user-space software and hardware.
 Untrusted processes interact with the system by issuing kernel \emph{system calls} across a hardware privilege boundary.
 In this way, the operating system kernel \emph{supervises} user-space processes.
-Usually, the source executables of untrusted processes need not have been written in the same language as the kernel.
 
@@
 
-Existing tools in the HOL familiy could be refactored around Supervisionary by changing their kernels to act as a shim between untrusted automation and the Supervisionary kernel.
+Existing tools in the HOL family could be refactored around Supervisionary by changing their kernels to act as frontends to the Supervisionary kernel for untrusted automation routines.
 Arguably, this increases the robustness of existing tools, enforcing separation by isolation, rather than module boundaries.

### 2021-10-25 paper/prisc-2022.tex
prisc-2022: switch emphasis style in some cases for the sake of uniformity

@@
 
-Interactive theorem proving software is typically designed around a trusted proof-checking \emph{kernel}, the sole system component capable of authenticating theorems.
+Interactive theorem proving software is typically designed around a trusted proof-checking \deffont{kernel}, the sole system component capable of authenticating theorems.
 Untrusted automation procedures reside outside of the kernel, and drives it to deduce new theorems via an API.
@@
 
-Changing tack, operating systems are also typically designed around a trusted \emph{kernel}, a privileged component responsible for---amongst other things---mediating interaction betwixt user-space software and hardware.
-Untrusted processes interact with the system by issuing kernel \emph{system calls} across a hardware privilege boundary.
-In this way, the operating system kernel \emph{supervises} user-space processes.
+Changing tack, operating systems are also typically designed around a trusted kernel, a privileged component responsible for---amongst other things---mediating interaction betwixt user-space software and hardware.
+Untrusted processes interact with the system by issuing kernel \deffont{system calls} across a hardware privilege boundary.
+In this way, the operating system kernel \deffont{supervises} user-space processes.
 
@@
 Yet, the two solutions to solving this problem, employed by the respective kinds of kernel, are very different.
-In this abstract, we explore designing proof-checking kernels as \emph{supervisory software}, where separation between kernel and untrusted code is enforced by \emph{privilege}.
+In this abstract, we explore designing proof-checking kernels as \deffont{supervisory software}, where separation between kernel and untrusted code is enforced by \emph{privilege}.
 
@@
 
-\emph{Supervisionary} is a proof-checking system for Gordon's HOL, structured as supervisory software, and provides a \emph{system interface} to untrusted code similar to an operating system's system call interface.
+\emph{Supervisionary} is a proof-checking system for Gordon's HOL, structured as supervisory software, and provides a \deffont{system interface} to untrusted code similar to an operating system's system call interface.
 Supervisionary is implemented as a WebAssembly (Wasm henceforth) host, allowing us to prototype rapidly.
@@
 This entails no risk of unsoundness providing our system interface is carefully designed.
-In particular, the kernel manages various private heaps within which \emph{kernel objects} are allocated, corresponding to the paraphenalia of any HOL implementation---type-formers, types, term constants, terms, and theorems---and never directly exposed to untrusted code.
+In particular, the kernel manages various private heaps within which \deffont{kernel objects} are allocated, corresponding to the paraphenalia of any HOL implementation---type-formers, types, term constants, terms, and theorems---and never directly exposed to untrusted code.
 Kernel objects are allocated in response to system calls like:
@@
 \end{displaymath}
-Here, both $\mathtt{left}$ and $\mathtt{right}$ are \emph{kernel handles}, assumed to point-to allocated terms, whilst $\mathtt{out}$ points-to a buffer in untrusted code's memory.
-If neither $\mathtt{left}$ nor $\mathtt{right}$ \emph{dangle}, and the types of their referents match, a fresh handle is generated which points-to a new HOL term application object, with internal pointers to the functional- and argument-terms.
+Here, both $\mathtt{left}$ and $\mathtt{right}$ are \deffont{kernel handles}, assumed to point-to allocated terms, whilst $\mathtt{out}$ points-to a buffer in untrusted code's memory.
+If neither $\mathtt{left}$ nor $\mathtt{right}$ dangle, and the types of their referents match, a fresh handle is generated which points-to a new HOL term application object, with internal pointers to the functional- and argument-terms.
 This handle is returned to the caller via the $\mathtt{out}$ pointer.
@@
 The manipulation and querying of kernel objects is performed defensively by the kernel itself on behalf of untrusted code.
-The kernel is careful to maintain invariants such as the \emph{inductivity} of its heaps, with nodes in the kernel object graph pointing-to allocated objects at all times.
+The kernel is careful to maintain invariants such as the inductivity of its heaps, with nodes in the kernel object graph pointing-to allocated objects at all times.
 
@@
 
-Note that one interesting consequence of this style of implementation is the ability to provide concise \emph{specifications} of Supervisionary's system interface functions.
+Note that one interesting consequence of this style of implementation is the ability to provide concise specifications of Supervisionary's system interface functions.
 Essentially, the Supervisionary kernel is a grand exercise in pointer manipulation, and as such our ABI specifications can be expressed as Hoare Triples, using Separation Logic as our assertion language.

### 2021-10-26 paper/prisc-2022.tex
prisc-2022: start bibliography

@@
 \settopmatter{printfolios=true, printccs=false, printacmref=false}
+
+\acmConference[PriSC'22]{6th Workshop on Principles of Secure Compilation}{January 22, 2022}{Philadelphia, PA, USA}
+\acmYear{2022}
+\acmISBN{}
+\acmDOI{}
+
+\setcopyright{none}
+
+\bibliographystyle{ACM-Reference-Format}
 
@@
 \newcommand{\rulefont}[1]{(\ensuremath{\mathbf{#1}})}
-
-\bibliographystyle{ACM-Reference-Format}
-
-\setcopyright{none}
 
@@
 Kernel and untrusted automation are typically implemented in the same programming language---the ``meta-language''---usually some functional programming language in the ML family.
-This strategy---introduced by Milner in his LCF proof assistant---is a reliability mechanism, aiming to ensure that any purported theorem produced by the system is indeed a legitimate entailment from the theory within the logic.
+This strategy---introduced by Milner in his LCF proof assistant~\cite{Milner1972LogicFC}---is a reliability mechanism, aiming to ensure that any purported theorem produced by the system is indeed a legitimate entailment from the theory within the logic.
 
@@
 \emph{Supervisionary} is a proof-checking system for Gordon's HOL, structured as supervisory software, and provides a \deffont{system interface} to untrusted code similar to an operating system's system call interface.
-Supervisionary is implemented as a WebAssembly (Wasm henceforth) host, allowing us to prototype rapidly.
+Supervisionary is implemented as a WebAssembly~\cite{DBLP:conf/pldi/HaasRSTHGWZB17} (Wasm henceforth) host, allowing us to prototype rapidly.
 
@@
 
-Given our use of Wasm, we could extend our system interface by also implementing the Wasm System Interface---a POSIX-like interface for Wasm.
+Given our use of Wasm, we could extend our system interface by also implementing the Wasm System Interface~\cite{Wasi}---a POSIX-like interface for Wasm.
 This would transform Supervisionary from a mere programmable proof-checker into a general-purpose sandbox, capable of executing arbitrary programs, albeit with an unusual interface for constructing proofs.
@@
 This theorem can then be used by $P$ in closing challenges from Supervisionary, issued at runtime, tying the correctness of $P$ to its execution.
+
+\newpage
+
+\bibliography{prisc-2022}
 

### 2021-10-26 paper/prisc-2022.tex
prisc-2022: more work on bibliography, a few more language tweaks

@@
 Kernel and untrusted automation are typically implemented in the same programming language---the ``meta-language''---usually some functional programming language in the ML family.
-This strategy---introduced by Milner in his LCF proof assistant~\cite{Milner1972LogicFC}---is a reliability mechanism, aiming to ensure that any purported theorem produced by the system is indeed a legitimate entailment from the theory within the logic.
+This strategy---introduced by Milner in his LCF proof assistant~\cite{Milner1972LogicFC}---is a reliability mechanism, aiming to ensure that any purported theorem produced by the system is indeed entailed by the theory within the logic.
 
@@
 Note that one interesting consequence of this style of implementation is the ability to provide concise specifications of Supervisionary's system interface functions.
-Essentially, the Supervisionary kernel is a grand exercise in pointer manipulation, and as such our ABI specifications can be expressed as Hoare Triples, using Separation Logic as our assertion language.
+Essentially, the Supervisionary kernel is a grand exercise in pointer manipulation, and as such our system call specifications can be expressed as Hoare Triples, using Separation Logic~\cite{DBLP:conf/lics/Reynolds02} as our assertion language.
 Writing $\mathtt{h} \mapsto_{\mathtt{trm}} \mathtt{Application(l, r)}$ to assert that the handle $\mathtt{h}$ points-to a term application (of the term pointed-to by $l$ to the term pointed-to by $r$), and writing $\mathtt{out} \mapsto \mathtt{b}$ to assert that $\mathtt{out}$ points-to the Boolean value $\mathtt{b}$, we have:
@@
 
+\vspace{\baselineskip}
 We now turn to speculation around potential uses of Supervisionary.
 Exploring what follows is still a work-in-progress.
-
-% note: explicitly mention polyglot theorem proving software?
-
 \paragraph{Runtime verification}
@@
 
+\vspace{\baselineskip}
 But: \emph{what happens if we blur the lines between Supervisionary's interfaces for system access and proof-checking?}
+\vspace{\baselineskip}
 
@@
 
-Note that this idea shares some similarities with \deffont{proof carrying code}, wherein binaries are accompanied with (skeleton) proofs of their adherence to some policy, and these proofs checked by the operating system prior to execution.
+Note that this idea shares some similarities with \deffont{proof carrying code}, wherein binaries are accompanied with (skeleton) proofs of their adherence to some policy, and these proofs checked by the operating system prior to execution~\cite{DBLP:conf/popl/Necula97}.
 However, the ideas sketched above generalise this: proofs can be generated dynamically, as the program executes, and could more aptly be called \deffont{proof generating code}.
+Processes and the Supervisionary kernel work together to come to a ``mutual understanding'' that the behaviour of the process is indeed concordant with the prevailing policy.
 
@@
 
-For example, the operational semantics and instruction decoding functionality of Wasm is clearly embeddable in HOL.
+For example, the operational semantics and instruction decoding functionality of Wasm is clearly embeddable in HOL~\cite{DBLP:conf/cpp/Watt18}.
 Using this, properties of a program $P$ to be executed under Supervisionary could be established statically, and registered with the kernel, perhaps interactively by a user or by another program executing before $P$ executes.

### 2022-05-11 paper/prisc-2022.tex
kernel+paper: final non-review copy of abstract, small kernel changes

@@
-\documentclass[sigplan, review]{acmart}
+\documentclass[sigplan]{acmart}
 \settopmatter{printfolios=true, printccs=false, printacmref=false}
@@
 
-\bibliography{prisc-2022}
+\input{prisc-2022.bbl}
 
 \end{document}
-
-\iffalse
-
-\paragraph{Some scene setting}
-
-Modern operating systems are used for a wide array of tasks.
-We focus on just two: providing a relatively portable abstract machine interface to user-space software, and enforcing system-wide security policies.
-
-Most operating systems are architected around a \deffont{kernel}, the root of all trust within a system, which engages in a grand conspiracy with its host microprocessor to take a monopoly on direct access to the system's hardware and sensitive system configuration options.
-%Relative to user-space code, the kernel executes at a high level of \emph{privilege}, using hardware \deffont{exception levels}.
-%Hardware ensures that sensitive system configuration features, such as system registers and page tables, may only be modified by software executing at a high level of privilege---in effect, only the kernel itself.
-Using this monopoly, the kernel protects its own memory from inadvertent or malicious modification, using paged memory protection, and acts as a ``choke point'' at which security policies are enforced.
-
-%Users-space code wishing to interact with a system resource must do so by interacting with the kernel.
-
-The kernel also offers an abstract machine interface to user-space software, smoothing over the vulgarities of programming against any particular peripheral, say.
-Useful abstractions, such as sockets and a filesystem, are also commonly offered.
-This interface is largely exposed as \deffont{system calls}, defining an \deffont{application binary interface}, or ABI henceforth, which user-space and kernel must understand to interact.
-%Executing a system call forces a \deffont{context switch}, within which the thread of execution enters the kernel.
-The result of a system call may be written directly into the memory of the calling code, via a pointer argument, for example.
-Note the asymmetry: in theory, the kernel can read and modify all of a user-space program's state.
-
-Code written in any language can make use of services via the kernel ABI.
-However, programs and runtimes typically make extensive use of low-level libraries---\texttt{libc} being an exemplar---that provide further abstraction over the raw ABI.
-
-This pattern of interaction between trusted and untrusted code via an ABI, wherein trusted code can read and modify the state of untrusted code, is common.
-Hypervisors, the Secure World of Arm TrustZone, and virtual machines (in the Wasm sense) are structured in this way.
-We call software designed in this pattern \deffont{supervisory software}.
-
-\paragraph{Supervisionary} is a prototype proof-checking system for Gordon's HOL structured as supervisory software, specifically as a Wasm \deffont{host environment}, which allows us to explore the consequences of the design, without becoming entangled in e.g.
-a real machine's boot ceremony.
-The ideas discussed herein could be applied to a real operating system.
-
-Supervisionary is implemented in Rust which does not pose any risk to the logical soundness of the system providing we expose a well-designed ABI.
-The kernel manages various private heaps within which \deffont{kernel objects} are allocated, corresponding to the paraphenalia of any HOL implementation---type-formers, types, term constants, terms, and theorems---and never directly exposed to untrusted code.
-Kernel objects are allocated in response to system calls like:
-\begin{displaymath}
-\mathtt{Term.Handle.AllocateApplication(left, right, out)}
-\end{displaymath}
-Here, both $\mathtt{left}$ and $\mathtt{right}$ are \deffont{kernel handles}, assumed to point-to allocated terms, whilst $\mathtt{out}$ points-to a buffer in untrusted code's memory.
-If neither $\mathtt{left}$ nor $\mathtt{right}$ \deffont{dangle}, and the types of their referents match, a fresh handle is generated which points-to a new HOL term application object, with internal pointers to the functional- and argument-terms.
-This handle is returned to the caller via the $\mathtt{out}$ pointer.
-
-Modulo bugs, system calls do not panic, but rather return \deffont{error codes} to signal failure.
-If $\mathtt{left}$ or $\mathtt{right}$ above dangle, for example, then a defined code is returned.
-These checks---which all system calls perform---ensure that the kernel heaps remain \deffont{inductive}, with nodes in the kernel object graph pointing-to allocated objects at all times, a key invariant.
-Moreover, kernel objects are only ever allocated, and \emph{never} deallocated or otherwise modified once created, thus ensuring the meaning of an object remains immutable (though it is possible that some \deffont{garbage collection} could be used).
-
-%To check if a handle points-to a $\lambda$-abstraction, for example, one may use:
-%\begin{displaymath}
-%\mathtt{Term.Handle.IsLambda(handle, out)}
-%\end{displaymath}
-%This again returns a defined error code if $\mathtt{handle}$ dangles and otherwise writes $\mathtt{true}$ or $\mathtt{false}$ to $\mathtt{out}$, as appropriate.
-
-The manipulation and querying of kernel objects is performed by the kernel itself on behalf of untrusted code.
-Space constraints prevent us from describing the entire Supervisionary ABI for working with, and on, kernel objects.
-However, note that theorems are also constructed in a similar way to terms, inductively building derivation trees.
-For example, the HOL symmetry rule is exposed as:
-\begin{displaymath}
-\mathtt{Theorem.Handle.AllocateSym(pre, out)}
-\end{displaymath}
-Here, $\mathtt{pre}$ points-to an existing theorem object $\Gamma \vdash r = s$ and after succeeding, passing obvious checks, $\mathtt{out}$ contains a handle that points-to a newly-allocated theorem $\Gamma \vdash s = r$.
-
-We have described the \emph{what} of Supervisionary and now turn our attention to the \emph{why}.
-This can be approached by examining Supervisionary both as a proof-assistant kernel, and as a programming platform.
-What follows is largely speculative, and a work-in-progress.
-
-\paragraph{Reinventing the proof-assistant}
-
-Proof-assistants in the LCF lineage are also architected around a \deffont{kernel}, with ostensibly little in common with operating system kernels.
-Examining, one sees that both are tasked with preserving invariants in the face of unbounded interaction with untrusted code.
-Moreover, whilst operating system kernels enjoy a monopoly on hardware access, proof-assistant kernels enjoy a monopoly on theorem \deffont{authentication}.
-In short, both act as roots of all trust within their domains.
-Yet, the two are implemented in different ways, with LCF kernels implemented as libraries in a \deffont{meta-language}, usually a functional language.
-Kernel interaction is via an \emph{API}, rather than an ABI.
-
-An immediate consequence of our design is that the meta-language is non-existent.
-Rather untrusted code ``driving'' the kernel can be written in any language capable of producing Wasm binary-compatible with our ABI.
-Arguably, the \emph{natural} choice for this is now not even functional, given the inherently stateful and pointer-oriented nature of the kernel.
-
-LCF-style interaction is centred around the REPL of the meta-language, or some thin layer built atop, itself started from an operating system shell---a kind of shell within a shell.
-Our design allows us to dispense with this middle-man, and develop proofs interactively directly in a shell implemented on top of our kernel.
-Moreover our kernel inherently internalises kernel objects, and could therefore make them available to browse as special files within a hierarchical filesystem, as discussed below, in analogy to how Linux exposes its internals as special files---for example, the contents of \texttt{/proc/} and \texttt{/dev/}.
-In this way, related material could appear grouped in directories, searched using command-line utilities, or opened by a process for use at runtime.
-
-\paragraph{Runtime monitoring}
-
-Nothing in principle stops us from extending Supervisionary's ABI with facilities for sockets, querying a hierarchical filesystem, and other similar system tasks.
-In particular, given our use of Wasm, we could implement the Wasm System Interface---a POSIX-like interface for Wasm---transforming Supervisionary from a programmable proof-checker into a general-purpose virtual machine, capable of supporting a wide range of programs.
-
-\emph{What happens if we blur the lines between Supervisionary's ABIs for system access and proof-checking?}
-Untrusted code could be challenged to prove some theorem each time it wished to open a file on the filesystem, or otherwise perform some side-effect.
-These theorems could be correctness or security-related theorems, corresponding to a prevailing \deffont{policy} in force.
-Moreover, as Supervisionary is capable of capturing the runtime state of untrusted code, these challenges can be HOL predicates that are functions of the reified runtime state of untrusted code, the kernel, and the arguments, and name of, the invoked system call.
-
-Two predicates of interest are $\lam{k}\lam{u}\lam{s}\top$ and $\lam{k}\lam{u}\lam{s}\bot$.
-Here, $\top$ and $\bot$ are truth and falsity, and $k$, $u$, and $s$ the kernel and untrusted code states, and packed system call metadata, reified as HOL data, at the point of invocation of the system call.
-One can always prove $\{\} \vdash \top$ and therefore $\lam{k}\lam{u}\lam{s}\top$ represents no restriction, whereas $\{ \} \vdash \bot$ is never provable, absent axioms, with $\lam{k}\lam{u}\lam{s}\bot$ representing a closing off of a system call.
-By making the challenge a function of the system call name and arguments, this closing off can be specific, banning a process from calling a particular system call, or a system call with a particular set of arguments, allowing us to mimic mechanisms like \texttt{seccomp} from Linux.
-
-We can go further.
-Metadata about the behaviour of a running process could be maintained: for example, a trace of the system calls invoked by a process, thus far, and used in forming security or correctness challenges, for example by forcing untrusted code to prove that writes to a socket only ever happen after a read.
-Information flow policies could also be captured, for example, by forcing untrusted code to commit to a pure function, $f$, to be computed before reading any files, and later showing that a result corresponds to $f$ applied to the content of a particular input file, thereby independent of the content of any other file.
-In short, HOL acts as \emph{lingua franca} between kernel and untrusted code, through which arbitrarily complex policies may be expressed.
-
-\deffont{Jailing}, wherein a process voluntarily sheds capabilities, is common in existing operating systems.
-This can be captured in Supervisionary by allowing a process to replace the prevailing policy, $\phi$, with a new policy $\psi$, after proving that $\psi$ is a refinment of $\phi$: $\{\} \vdash \fall{k}\fall{u}\fall{s} \psi\ k\ u\ s \longrightarrow \phi\ k\ u\ s$.
-Note that this expresses that the states described by $\psi$ are a subset of those described by $\phi$.
-
-\deffont{Proof carrying code} modified the operating system loader, and endowed binaries with skeleton proofs of adherence to some policy, later reconstructed by the loader prior to execution.
-The ideas sketched above are different: proofs are generated dynamically, as the program executes, and more aptly called \deffont{proof generating code}.
-Moreover, theorems could be generated statically, offline, and used dynamically:
-
-\paragraph{Lastly, specifications}
-
-Supervisionary is an extended exercise in pointer manipulation.
-The semantics of our system calls can therefore be expressed as Hoare Triples, using formulae of \deffont{Separation Logic} as an assertion language.
-Write $\mathtt{h} \mapsto_{\mathtt{trm}} \mathtt{Application(l, r)}$ to assert that the handle $\mathtt{h}$ points-to a term application (of the term pointed-to by $l$ to the term pointed-to by $r$).
-Write $\mathtt{out} \mapsto \mathtt{b}$ to assert that $\mathtt{out}$ points-to the Boolean value $\mathtt{b}$.
-Write $P \star Q$ for the \deffont{separating conjunction}.
-Finally, write $\{ P \} C \{ Q \}$ to assert that $C$ executing in a state satisfying $P$ succeeds and produces a state satisfying $Q$.
-We then have the following example \deffont{local axiom}:
-\begin{gather*}
-\{ h \mapsto_{\mathtt{trm}} \mathtt{Application(l, r)} \star \mathtt{out} \mapsto b \} \\
-\mathtt{Term.Handle.IsApplication(h, out)} \\
-\{ h \mapsto_{\mathtt{trm}} \mathtt{Application(l, r)} \star \mathtt{out} \mapsto \mathtt{True} \}
-\end{gather*}
-Note that the structural \deffont{frame rule} captures the fact that any system call only manipulates a defined footprint of kernel and untrusted code's memory.
-
-\fi
-
-\end{document}
-
-To enforce system-wide security policies, most operating systems implement a self-contained \emph{kernel} which serves as the root of all trust within the system.
-The kernel is the sole component able to directly interact with hardware, and provides services to executing user-space software, for example the abstraction of a hierarchical filesystem.
-Untrusted user-space code resides outside of the kernel and interacts with it through a defined \emph{ABI}.
-
-To enforce logical soundness, many proof-checking systems implement a self-contained \emph{kernel} which serves as the root of all trust within the system.
-The kernel is the sole system component able to authenticate a proof as legitimate.
-Untrusted code resides outside of the kernel and interacts with it through a defined \emph{API}.
-
-Though ostensibly wildly different, in \emph{some} respects the goals of operating system and proof-checking kernels are the same: the maintenance of important system invariants in the face of unbounded interaction with untrusted code.
-The mechanisms employed to protect these invariants, by each type of system, are different however.
-For operating systems, machine-oriented notions of separation are employed, with the kernel executing at a higher prilege level than untrusted code, and cosseted by paged memory protection---hardware and privileged software work together to provide an abstract machine interface to untrusted code.
-On the other hand, LCF-style proof-checking kernels are typically implemented as libraries within a \emph{meta-language}---usually an ML dialect---with invariants maintained through type abstraction and code encapsulation behind module boundaries.
-Logical soundness follows from type-soundness of the meta-language.
-
-Yet, what happens if we structure proof-checking software more like supervisory system software---such as operating systems, hypervisors, and programming language runtimes?
-Now, the proof-checking kernel is not implemented as a library, but rather becomes a programming platform, exposing an abstract machine to untrusted code, in much the same way that the Linux kernel, for example, exposes an abstract machine to user-space code through its system call interface.
-We are currently developing a prototype system, \emph{Supervisionary}, implemented as a WebAssembly host, to explore the consequences of this design.
-
-One immediate consequence of Supervisionary's design is the abolition of the LCF-style meta-language.
-Rather, Supervisionary's kernel is implemented in Rust---a notably \emph{unsafe} systems language.
-Moreover, untrusted code which ``drives'' the Supervisionary kernel to produce an authenticated theorem need not necessarily be written in the same language that the kernel is written in: rather any language that can be compiled to code that is binary-compatible with the Supervisionary abstract machine can be used, even a mixture of them, in much the same way that Linux user-space software can be written in Haskell, Ada, or any other language that can be linked together, despite the kernel being written in C.
-
-Note that the use of Rust to implement the Supervisionary kernel does not pose any risk to the logical soundness of the system, providing we expose a carefully designed abstract machine through the Supervisionary ABI.
-The Supervisionary kernel manages various private memory spaces, or heaps, within which \emph{kernel objects} are managed.
-These kernel objects correspond to the parephenalia of a typical HOL implementation---type-formers, types, term constants, terms, and theorems---but are never directly exposed to untrusted code.
-Rather, kernel objects are allocated by the kernel in one of its private heaps, through ABI functions like the following:
-\begin{displaymath}
-\mathtt{Term.Handle.AllocateApplication(left, right, out)}
-\end{displaymath}
-Above, both $\mathtt{left}$ and $\mathtt{right}$ passed to the function are kernel handles, assumed to point-to allocated terms.
-On the other hand, the $\mathtt{out}$ pointer points-into the heap of the code calling the function (note that the kernel can freely read from, and write to, the memory of untrusted code, given it is supervisory software).
-Providing both $\mathtt{left}$ and $\mathtt{right}$ indeed point-to allocated terms, and their types match, a new kernel object corresponding to a term application is allocated, with internal pointers to the functional- and argument-terms, in the kernel's heap.
-This new object is pointed-to by a freshly-generated handle, returned back to the caller via the $\mathtt{out}$ pointer.
-
-Note that Supervisionary ABI functions do not panic, or otherwise abort, modulo bugs in the implementation of the kernel itself.
-Rather, errors are signalled to calling code via a defined \emph{error code} which can be used to deduce the cause of failure: if either $\mathtt{left}$ or $\mathtt{right}$ above dangle, then the $\mathtt{DanglingHandle}$ code is returned, for example.
-Note that these checks for dangling handles---which all Supervisionary ABI functions perform---ensures that the kernel heaps remain \emph{inductive}, or ``downward closed'', with nodes in the kernel object graph always pointing to previously-allocated objects at all times---a basic kernel invariant.
-
-Post allocation, all manipulation and querying of kernel objects is performed by the kernel itself.
-For example, to check if a handle points-to a $\lambda$-abstraction, untrusted code can call the following function:
-\begin{displaymath}
-\mathtt{Term.Handle.IsLambda(handle, out)}
-\end{displaymath}
-This again returns a defined error code if $\mathtt{handle}$ dangles, leaving $\mathtt{out}$ unchanged, or otherwise writing $\mathtt{true}$ or $\mathtt{false}$ to the $\mathtt{out}$ pointer as appropriate.
-
-Unfortunately, space constraints here prevent us from providing a pr\`{e}cis of all ABI functions for allocating, querying, destructuring, and otherwise manipulating type-formers, types, term constants, and terms.
-However, note that theorems are also gradually constructed in an analogous way to how terms are constructed.
-For example, the HOL symmetry inference rule is exposed as:
-\begin{displaymath}
-\mathtt{Theorem.Handle.AllocateSym(pre, out)}
-\end{displaymath}
-Here, $\mathtt{pre}$ points-to an allocated theorem $\Gamma \vdash r = s$ and after successfully completing, $\mathtt{out}$ contains a handle that points-to a newly-allocated theorem $\Gamma \vdash s = r$.
-Note that the heap of theorem objects must only ever grow monotonically, and a theorem object must always remain immutable once allocated by the kernel.
-This immutable aspect also applies to any kernel object mentioned by any allocated theorem---terms, types, and so on: it would be decidedly \emph{inconvenient} if the meaning of the HOL truth constant were redefined, for example.
-On the other hand, it seems possible that some kernel heaps could shrink, for example allocated type-formers, types, term constants, and terms, that are not referenced transitively by any allocated theorem could be garbage collected.
-This is not yet implemented in Supervisionary.
-
-Note that one consequence of this style of implementation is the ability to provide concise \emph{specifications} of Supervisionary ABI functions.
-Essentially, the Supervisionary kernel is a grand exercise in pointer manipulation, and as such our ABI specifications can be expressed as Hoare Triples, using Separation Logic as our assertion language.
-Writing $\mathtt{h} \mapsto_{\mathtt{trm}} \mathtt{Application(l, r)}$ to assert that the handle $\mathtt{h}$ points-to a term application (of the term pointed-to by $l$ to the term pointed-to by $r$), and writing $\mathtt{out} \mapsto \mathtt{b}$ to assert that $\mathtt{out}$ points-to the Boolean value $\mathtt{b}$, we have, for example:
-\begin{gather*}
-\{ h \mapsto_{\mathtt{trm}} \mathtt{Application(l, r)} \} \\
-\mathtt{Term.Handle.IsApplication(h, out)} \\
-\{ \mathtt{out} \mapsto \mathtt{True} \}
-\end{gather*}
-(Here, the triple $\{ P \} C \{ Q \}$ asserts that if the command $C$ executes in a state concordant with $P$ then the command succeeds and produces a state concordant with $Q$.)
-
-The material presented so far is implemented in Supervisionary.
-We now speculate on opportunities for future work.
-
-First, observe that Supervisionary is not merely a rehashing of LCF, perhaps with some minor tweaks.
-In particular, Supervisionary ``internalises'' theorems as kernel objects in a way that LCF does not: the latter authenticates theorems by producing an opaque object, of a particular abstract type, which in a sense``passes out'' of the kernel.
-Moreover, Supervisionary behaves like a typical operating system, or programming language runtime, in that it takes a monopoly on the system's resources from the point of view of untrusted code.
-To access the filesystem, or a device for example, untrusted code generally has no choice but to interact with the operating system, which gates access based on a prevailing security policy. 
-Adding ABI functions to Supervisionary for filesystem access allows WebAssembly programs executing under Supervisionary to access the system's filesystem, at the cost of passing through the Supervisionary kernel.
-What, then, if we force code wishing to interact with the filesystem (or any other system resource) to provide a particular \emph{theorem} to the kernel before that access is granted?

### 2022-05-11 kernel/src/runtime_state.rs

@@ -3540,6 +3540,7 @@ impl RuntimeState {
         let conclusion = self
             .substitution(body, vec![((name, typ), trm)])
             .expect(DANGLING_HANDLE_ERROR);
+        let premisses = thm.premisses().clone();
 
         Ok(self.admit_theorem(Theorem::new(premisses, conclusion)))
     }
